### 2022-08-04 latex/summary.tex
add LaTeX files

\documentclass[a4paper]{report}

\usepackage{needed}

\title{PHYS213}
\author{Alfaifi, Ammar}
\date{\today}

\begin{document}
    \maketitle
    \chapter{The Quantum Theory of Light}
        \section{Heartz's Experiments; Light as EM Wave}

            \paragraph{$\bullet$} Between 1865 and 1873, James Clerk Maxwell, and according to his theory, 
            an alternating current would set up fluctuating electric and magnetic fields in the 
            region surrounding the original disturbance.
            \paragraph{$\bullet$} These waves were predicted to have a frequency equal to the frequency of the current oscillations.
            \paragraph{$\bullet$} And more importnatly, they would behave in every way like \textit{light}.
            \paragraph{$\bullet$} Naturally this led to the unifying and simplifying postulate that light was also a 
            type of Maxwell wave or electromagnetic disturbance, created by extremely high frequency electric oscillators in matter.
            \paragraph{$\bullet$} it was apparent that this model of light emission was incapable of direct experimental 
            verification, because the highest electrical frequencies then attainable were about $10^9Hz$.

            \begin{wrapfigure}{r}{0.25\linewidth}
                \includegraphics[width=0.9\linewidth]{figures/Hertz_transmitter.png}
                \caption{Heartz's Transmitter.}
                \label{fig:heartz's transmitter}
            \end{wrapfigure}

            \paragraph{$\bullet$} Heinrich Hertz showed that Maxwell’s theory was correct and that an oscillating electric 
            current does indeed radiate electromagnetic waves that possess every characteristic of light 
            except the same wavelength as light.

            \paragraph{$\bullet$} By using a simple spark gap oscillator in Fig.\eqref{fig:heartz's transmitter}.

            \paragraph{$\bullet$} Hertz next proceeded to show that these electromagnetic waves could be reflected, 
            refracted, focused, polarized, and made to interfere—- in short, he convinced physicists 
            of the period that Hertzian waves and light waves were one and the same.

        \section{Blackbody Radiation}
            The problem is to predict the radiation intensity at a given wavelength emitted by 
            a hot glowing solid at a specific temperature.
            \paragraph{$\bullet$} In 1792, Thomas Wedgwood observed that all the objects in his ovens, regard- less 
            of their chemical nature, size, or shape, became red at the same tempera- ture.
            \paragraph{$\bullet$} In 1859, Gustav Kirchhoff proved a theorem when he showed by arguments based 
            on thermodynamics that for any body in thermal equilibrium with radiation the emitted power is 
            proportional to the power absorbed. More specifically,
            \coloredeq{Kirchhoff emitted power theorem}{e_f=J(f, T)A_f}
            {\tiny \begin{itemize}
                \item $e_f$ is the power emitted per unit area per unit frequenct by a heated object.
                \item $A_f$ is the absorption power (fraction of the incident power absorbed per 
                unit area per unit frequency by the heated object).
                \item $J(f, T)$ is a universal function (the same for all bodies) that depends only on $f$.
            \end{itemize}}

            \paragraph{$\bullet$} A \textbf{blackbody} is defined as an object that absorbs all the electromagnetic 
            radiation falling on it and consequently appears black. It has $a_f=1$ for all frequencies 
            and so Kirchhoff’s theorem for a blackbody becomes,
            \begin{align} \label{kirchhoff theorm blackbody}
                e_f=J(f, T)
            \end{align}

            \paragraph{$\bullet$} Eq.\eqref{Kirchhoff emitted power theorem} shows that the power emitted only
            per unit area per unit frequency by a blackbody depends only on the light frequency and the 
            temperature.

            \paragraph{$\bullet$} Because absorption and emission are connected by Kirchhoff’s theorem, 
            we see that a \textit{blackbody} or perfect absorber is also an ideal \textit{radiator}.\\
            In practice, a small opening in any heated \textit{cavity}, such as a port in an oven, behaves 
            like a blackbody because such an opening traps all incident radiation.

            \paragraph{$\bullet$} In 1879, Josef Stefan found experimentally that the total power per unit 
            area emitted at all frequencies by a hot solid, was proportional to the fourth 
            power of its absolute temperature. Therefore, \textit{Stefan’s law} may be written as
            \coloredeq{eq:stefan's law}{e_{total}=\int_0^\infty e_f \, df = \sigma T^4}
            {\tiny \begin{itemize}
                \item $e_{total}$ is the power per unit area emitted at the surface of the blackbody at all frequencies.
                \item $e_f$ is the power per unit area per unit frequency emitted by the blackbody.
                \item $T$ is the absolute temperature of the body.
                \item $\sigma$ is the Stefan – Boltzmann constant $\sigma= \SI[per-mode=symbol]{3e5}{\W\per\m\squared\per\K\tothe{4}}$.
            \end{itemize}}
            A body that is not an ideal radiator will obey the same general law but with a 
            coefficient, $a$, less than 1: \begin{align}\label{emitted power from non-ideal body}e_{total}=a \sigma T^4\end{align}

            \begin{wrapfigure}{r}{0.25\linewidth}
                \includegraphics[width=0.9\linewidth]{figures/blackbody emission.png}
                \caption{Emission from a glowing solid.}
                \label{fig:blackbody emission}
            \end{wrapfigure}

            \paragraph{$\bullet$} Notice that in Fig.\eqref{fig:blackbody emission} $\lambda_{max}$ shifts
            to shorter wavelengths as the blackbody gets hotter.\\
            In 1893, Wilhelm Wien proposed a general form for the blackbody distribution law 
            $J(f, T)$ that gave the correct experimental behavior of $\lambda_{max}$ with temperature. This 
            law is called \textbf{Wien’s displacement law} and may be written
            \begin{align} \label{Wien's displacement law}
                \lambda_{max} T = \SI{2.898e-3}{\m\K}
            \end{align}
            {\tiny \begin{itemize}
                \item $\lambda_{max}$ is the the blackbody’s maximum intensity.
                \item $T$ is the absolute temperature of the surface.
            \end{itemize}}

            \paragraph{$\bullet$} Let's consider the energy per unit volume per unit frequency of 
            the radiation within the blackbody cavity, $u( f, T )$. For light in equilibrium with 
            the walls, the power emitted per square centimeter of opening is simply proportional to 
            the energy density of the light in the cavity.

            \paragraph{$\bullet$} An important guess as to the form of the universal function $u(f, T)$ was
            made in 1893 by Wien and had the form 
            \begin{align} \label{Wien's exponential law}
                u(f,T) = A f^3 e^{- \beta f/T}
            \end{align}
            {\tiny \begin{itemize}
                \item $A$ and $\beta$ are constants
            \end{itemize}}

            \paragraph{$\bullet$} Within a year Friedrich Paschen, had confirmed Wien’s guess 
            by working in the then difficult infrared range of 1 to \SI{4}{\mu\m} and at temperatures 
            of 400 to \SI{1600}{\K}.Paschen had made most of his 
            measurements in the maximum energy region of a body heated to 1500K and had found good 
            agreement with Wien’s exponential law. In 1900, however, Lummer and Pringsheim extended 
            the measurements to \SI{18}{\mu\m}, and Rubens and Kurlbaum went even farther$-$ to \SI{60}{\mu\m}.
            Both teams concluded that Wien’s law failed in this region.

            \paragraph{\color{c3}Enter Planck}
            On a Sunday evening early in October of 1900, Max Planck discovered the famous blackbody formula.

            Planck knew that Wien’s law agreed well with the data at high frequency and indeed had been 
            working hard for several years to derive Wien’s exponential law from the principles of statistical 
            mechanics and Maxwell’s laws. See Eq.\eqref{eq:Plank distribution law}

            \paragraph{\color{c3}The Quantum Energy}
            Planck was convinced that blackbody radiation was produced by vibrating submicroscopic 
            electric charges, which he called \textit{resonators}. He assumed that the walls of a glowing 
            cavity were composed of literally billions of these \textit{resonators} (whose exact nature was 
            unknown at the time), all vibrating at different frequencies. Hence, according to Maxwell, 
            each \textit{oscillator} should emit radiation with a frequency corresponding to its vibration 
            frequency. \textbf{Also according to \textit{classical Maxwellian theory}, an oscillator of frequency $f$ 
            could have \textbf{any} value of energy and could change its amplitude continuously as it 
            radiated any fraction of its energy.}

            \paragraph{$\bullet$} This is where Planck made his revolutionary proposal. 
            To secure agreement with experiment, Planck had to assume that the total energy of a 
            resonator with mechanical frequency $f$ could only be an integral multiple of $hf$ or
            \coloredeq{eq:Planck resonator energy}{E_{resonator}=nhf\quad n=1, 2, 3, \dots}
            he concluded that emission of radiation of frequency f occurred when a resonator dropped 
            to the next lowest energy state.\\
            Thus the resonator can change its energy only by the difference $\Delta{E}$ according to
            \begin{align} \label{eq:energy change of hf}
                \Delta{E} = hf
            \end{align}
            \textit{That is, it cannot lose just any amount of its total energy, but only a finite amount, $hf$, 
            the so-called \textbf{quantum} of energy.}

            \paragraph{$\bullet$} Planck explained the continuous spectrum of the blackbody by assuming 
            that \textit{the heated walls contained resonators vibrating at many different frequencies}, each 
            emitting light at the same frequency as its vibration frequency.

            \paragraph{$\bullet$}  considering the conditions leading to equilibrium between the \textit{wall resonators}
            and the \textit{radiation} in the blackbody cavity, he found that the 
            spectral energy density $u(f, T)$ could be expressed as
            \coloredeq{eq:spectral energy desity}{u(f, T)\, df= \bar{E}N(f)\, df}
            {\tiny \begin{itemize}
                \item $N(f)$ number of oscillators having frequency between $f$ and $f + df$.
                \item $\bar{E}$ average energy emitted per oscillator.
            \end{itemize}}

            Furthermore, Planck showed that the number of oscillators with frequency
            between $f$ and $f + df$ was proportional to $f^2$ or 
            \begin{align}
                \label{eq:number of oscillator}
                N(f)\, df = \frac{8 \pi f^2}{c^3} \, df
            \end{align}

        \section{The Rayleigh-Jeans Law and Planck's Law}
            Both Planck’s law and the Rayleigh–Jeans law\footnote{the classical theory of blackbody radiation formulated by 
            Lord Rayleigh, John William Strutt, 1842–1919, English physicist, and James Jeans, 1887–1946, English 
            astronomer and physicist} may be derived using the idea that the blackbody radiation energy per unit volume can expressed
            as in Eq.\eqref{eq:spectral energy desity}. We will consider \textbf{Rayleigh–Jeans} and \textbf{Planck} calculations 
            to see the effect on $u(f, T)$ of calculating $\bar{E}$ from a \textit{continuous} distribution of classical oscillator energies 
            (Rayleigh–Jeans) as opposed to a \textit{discrete} set of quantum oscillator energies (Planck).

            \paragraph{\color{c3}Rayleigh-Jeans Law} % (fold)
            \label{par:Rayleigh-Jeans Law}     
            \begin{wrapfigure}{r}{0.2\linewidth}
                \includegraphics[width=0.9\linewidth]{figures/standing wave.png}
                \caption{A one-dimensional harmonic oscillator.}
                \label{fig:A one-dimensional harmonic oscillator}
                \includegraphics[width=0.9\linewidth]{figures/spectral energy density.png}
                \caption{The failure of the classical Rayleigh–Jeans law Eq.\eqref{eq:Rayleigh-Jeans energy density} 
                to fit the observed spectrum of a blackbody heated to 1000 K.}
                \label{fig:spectral energy density}
            \end{wrapfigure}

            \textit{While Planck concentrated on the thermal equilibrium of cavity radiation with oscillating electric charges in 
            the cavity walls, Rayleigh concentrated directly on the electromagnetic waves in the cavity}. 
            \paragraph{\color{c3}$\bullet$}Rayleigh and Jeans 
            reasoned that the \textit{standing electromagnetic waves} in the cavity could be considered to have a \texttt{temperature} $T$, 
            because they constantly exchanged energy with the walls and caused a thermometer within the cavity to reach 
            the same temperature as the walls. Further, they considered a \textit{standing polarized electromagnetic wave} to be 
            equivalent to a one-dimensional oscillator \eqref{fig:A one-dimensional harmonic oscillator}.

            \paragraph{\color{c3}$\bullet$}Using the same general idea as Planck, they expressed 
            the energy density as a product of the number of standing waves (oscillators) and the average energy per 
            oscillator. They found the average oscillator energy $\bar{E}$ to be independent of frequency and equal to $k_B T$ from 
            the \textit{Maxwell-Boltzmann distribution law} (see Chapter 10). According to this distribution law, 
            \begin{align}
                \label{eq:probability of finding a system}
                P(E) = P_0\, e^{-(E-E_0)/k_B T}
            \end{align}
            {\tiny \begin{itemize}
                \item the probability $P$ of finding an individual system (such as a molecule or an atomic oscillator) 
                \item with energy $E$ above some  
                \item minimum energy($E_0$) in a large group of systems at 
                \item temperature $T$
                \item where P0 is the probability that a system has the minimum energy.
            \end{itemize}}

            \paragraph{\color{c3}$\bullet$} In the case of a discrete set of allowed energies, the average energy($\bar{E}$) is given by
            \begin{align}
                \label{eq:average energy}
                \bar{E} = \frac{\Sigma E\, P(E)}{\Sigma P(E)}
            \end{align}
            where division by the sum in the denominator serves to normalize the total probability to 1.

            \paragraph{\color{c3}$\bullet$} \textit{In the classical case considered by Rayleigh}, an oscillator can have any energy $E$ in
            a continuous range from 0 to $\infty$. Thus the Eq.\eqref{eq:average energy} muse be replaced with integrals, so 
            \begin{align}
                \label{eq:average energy integral}
                \bar{E} = \frac{\int_0^\infty Ee^{-E/k_B T}\, dE}{\int_0^\infty e^{-E/k_B T}\, dE} = k_B T
            \end{align}
            The derivation of the density of modes, $N( f )\, df$, gives
            \begin{align}
                \label{eq:density of modes}
                N(f)\, df = \frac{8 \pi f^3}{c^3} \, df \qquad \text{or in wavelength} \qquad 
                N(\lambda) \, d\lambda = \frac{8 \pi }{\lambda^4}\, d\lambda  
            \end{align}
            The spectral energy density is simply the \textit{density of modes} multiplied by $k_B T$, or
            \begin{align}
                \label{eq:Rayleigh-Jeans energy density}
                u(f, T)\, df = \frac{8 \pi f^2}{c^3} k_B T\, df \qquad \text{or in wavelength} \qquad 
                u(\lambda, T)\, d\lambda = \frac{8 \pi}{\lambda^4} k_B T\, d\lambda
            \end{align}

            \paragraph{\color{c3}$\bullet$} However, as in Figure~\ref{fig:spectral energy density}, this classical expression, 
            known as the \textit{Rayleigh–Jeans law}, does not agree with the experimental results in the short wavelength region.
            Eq.\eqref{eq:Rayleigh-Jeans energy density} diverges as $\lambda \to 0$, predicting unlimited energy emission in the 
            ultraviolet region, which was dubbed the “ultraviolet catastrophe”. To conclude that the classical theory fails 
            to explain blackbody radiation.
            % paragraph Rayleigh-Jeans Law (end)

            \paragraph{\color{c3}Planck Law} % (fold)
            \label{par:Planck Law}
            Planck concentrated on the energy \textit{states of resonators} in the cavity walls and used the condition that the 
            \texttt{resonators and cavity radiation were inequilibrium} to determine the spectral quality of the radiation. 
            By thermodynamic reasoning (and apparently unaware of Rayleigh’s derivation), he arrived at the same expression 
            for $N( f )$ as Rayleigh. However, Planck arrived at a different form for $E$ by allowing only \textit{discrete} values of 
            energy for his resonators. He found, using the Maxwell-Boltzmann distribution law,
            \begin{align}
                \label{eq:plank average energy}
                \bar{E} = \frac{hf}{e^{hf/k_b T}-1}
            \end{align}
            Multiplying $\bar{E}$ by $N(f)$ gives the Planck distribution law,
            \coloredeq{eq:Plank distribution law}{
                \begin{aligned}
                    u(f, T)\, df &= \frac{8\pi f^2}{c^3} \left(\frac{hf}{e^{hf/k_B T}-1}\right)\, df \\
                    \text{or in wavelength} \qquad \qquad 
                    u(\lambda, T)\, d\lambda &= \frac{8\pi h c}{\lambda^5(e^{hc/\lambda k_B T}-1)}\, d\lambda
                \end{aligned}
            }
            {\tiny \begin{itemize}
                \item $h$ Planck's constant \SI{6.626e-34}{\J\s}
                \item $k_B$ Boltzmann's constant \SI{1.380e-23}{\J\per\K}
            \end{itemize}}
            \noindent
            In summary, Planck arrived at his blackbody formula by making two startling assumptions: 
            \begin{enumerate}
                \item The energy of a charged oscillator of frequency $f$ is limited to discrete values $nhf$.
                \item During emission or absorption of light, the change in energy of an oscillator is $hf$.
            \end{enumerate}

            \paragraph{\color{c3}$\bigstar$} By integrating Eq.\eqref{eq:Plank distribution law} for $0 \to \infty$ we will get 
            Stefan’s law in Eq.\eqref{eq:stefan's law}.
            % paragraph Planck Law (end)

        \section{Light Quantization and The Photoelectric Effect}
            Einstein recognized an inconsistency between Planck’s quantization of oscillators in the walls 
            of the blackbody and Planck’s insistence that the cavity radiation consisted of classical electromagnetic waves.

            \paragraph{\color{c3}$\bigstar$} By showing that the change in entropy of blackbody radiation was like 
            the change in entropy of an ideal gas consisting of independent particles, Einstein reached the conclusion 
            that light itself is composed of “grains,” irreducible finite amounts, or quanta of energy.

            \begin{wrapfigure}{r}{0.2\linewidth}
                \includegraphics[width=0.9\linewidth]{figures/photocurrent versus applied voltage.png}
                \caption{Photocurrent versus applied voltage}
                \label{fig:Photocurrent versus applied voltage}
                \includegraphics[width=0.9\linewidth]{figures/Kmax.png}
                \caption{$K_{max}$ versus frequency}
                \label{fig:Kmax versus frequency}
            \end{wrapfigure}

            \paragraph{$\bullet$} Hertz first established that clean metal surfaces emit charges when exposed to 
            ultraviolet light. In 1888 Hallwachs discovered that the emitted charges were negative, 
            and in 1899 J. J. Thomson showed that the emitted charges were electrons, now called photoelectrons.

            \paragraph{$\bullet$} In 1902,  Philip Lenard was studying the photoelectric effect with 
            intense carbon arc light sources. He found that electrons are emitted from the metal with a range of 
            velocities and that the maximum kinetic energy of photoelectrons, $K_{max}$, does not depend on the 
            \textit{intensity} of the exciting light. He lso indicated that $K_{max}$ increases with light \textit{frequency}.

            \paragraph{$\bullet$} In the photoelectric setup, $K_{max}$ is easily measured by applying a retarding voltage and gradually 
            increasing it until the most energetic electrons are stopped and the photocurrent becomes zero. At this point,
            \begin{align}
                \label{eq:Kmax of a photoelectron}
                K_{max} = \frac{1}{2} m_e v^2_{max} = eV_s
            \end{align}

            As shown in Fig.\eqref{fig:Photocurrent versus applied voltage} Lenard found that 
            $K_{max}$ or $V_s$ is independent of light intensity $I$.\\
            The classical electromagnetic theory has major difficulties explaining:
            \begin{enumerate}
                \item The linear dependence of  $K_{max}$ on light frequency, shown in Fig.\eqref{fig:Kmax versus frequency}.
                \item There is no \textit{time lag} between the start of illumination and the start of the photocurrent.
            \end{enumerate}

            \paragraph{$\bigstar$} Einstein’s theory of the photoelectric effect:
            \textbf{\color{c3}
            He maintained that the energy of light is not distributed evenly over the classical wavefront, 
            but is concentrated in discrete regions (or in “bundles”), called quanta, each containing energy, $hf$.}

            \paragraph{$\bullet$} Therefore, according to Einstein, the maximum kinetic energy for emitted electrons is
            \coloredeq{eq:Einstein's Kmax}{K_{max} = hf - \phi}
            where $\phi$ is the \textbf{work function} of the metal, which corresponds to the minimum energy 
            with which an electron is bound in the metal.
            Eq.\eqref{eq:Einstein's Kmax} explained the independence of $K_{max}$ and intensity found by Lenard. 
            For a fixed light frequency $f$, an increase in light intensity means more photons and more photoelectrons 
            per second, although $K_{max}$ remains unchanged according to this equation.

            \paragraph{$\bullet$} By setting $K_{max}=0$ we get the characteristic threshold frequency,
            \begin{align}
                \label{eq:threshold frequency}
                f_0 = \frac{\phi}{h}
            \end{align}

        \section{X-Rays}
            X-rays were discovered in 1895 by the German physicist Wilhelm Roentgen. 
            He found that a beam of high-speed electrons striking a metal target produced a new and extremely penetrating type of radiation.
            Max von Laue in Germany and William Henry Bragg and William Lawrence Bragg (a father and son team) in 
            England suggested using single crystals such as calcite as natural three-dimensional gratings, the 
            periodic atomic arrangement in the crystals constituting the \textit{grating rulings}.

            \begin{wrapfigure}{r}{0.2\linewidth}
                \includegraphics[width=0.9\linewidth]{figures/x-ray acattering.png}
                \caption{Bragg scattering of x-rays from successive planes of atoms.}
                \label{fig:Bragg scattering}
            \end{wrapfigure}

            \paragraph{$\bullet$} A particularly simple method of analyzing the scattering of x-rays from parallel 
            crystal planes was proposed by W. L. Bragg in 1912. Consider two successive planes of atoms 
            as shown in Fig.\eqref{fig:Bragg scattering}.

            \paragraph{$\bullet$} The adjacent atoms in a single plane, $A$, will scatter constructively if the angle of 
            incidence, $\theta_i$, equals angel of reflaction, $\theta_r$.\\
            Atoms in \textit{successive planes} (A \& B), will scatter constructively if the the path length difference for rays (1) \& (2)
            is a ehole number of wavelength, thus $n\theta$, so 
            \begin{align}
                \label{The condition of scattering}
                \overline{AB} + \overline{BC} = n\theta, \qquad n=1, 2, 3, \dots
            \end{align}
            Thus Bragg equation is 
            \coloredeq{eq:Bragg equation}{\Delta{L} = 2d\sin{\theta} = n\lambda}

\end{document}

### 2022-08-04 latex/MATH208/chapters/chapter1.tex
add LaTeX files

\chapter{First-Order Differential Equations}
    \section{Differential Equations \& Mathmatical Models}
        \begin{itemize}
            \item An equation relating an unknown function and one or more of its derivatives 
            is called a differential equation, such $\frac{dx}{dt}=x^2+t^2$.
            \item The order of a differential equation is the order of the highest derivative that appears in it.
            \item The most general form of an \textbf{$n$th-order} differential equation with 
            independent variable 5 and unknown function or dependent variable $y=y(x)$ is
            \begin{align} \label{nth-order}
                F(x, y^{\prime}, y^{\prime \prime}, \dots, y^{(n)})=0
            \end{align}
            \item The continuous function $u=u(x)$ is a solution of the differential equation in \eqref{nth-order} 
            on the interval $I$ provided that the derivatives $u^\prime, u^{\prime \prime}, \dots, u^{(n)}$ 
            exist on $I$ and 
            \begin{align}
                F(x, u^{\prime}, u^{\prime \prime}, \dots, u^{(n)})=0
            \end{align}
            for all $x$ in $I$.
            \item Then $u=u(x)$ \textbf{satisfies} the differential equation in \eqref{nth-order} on $I$.
            \item An $n$th-order differential equation ordinarily has an $n$-parameter family of solutions.
            \item  An \textbf{ordinary} \textit{differential equations} where the unknown function 
                (dependent variable) depends on only a single independent variable.
            \item We concentrate on first-order differential equations with initial condition to \textbf{solve}
                \begin{align} \label{ODE with initial cond.}
                    \frac{dy}{dx} = f(x, y), \quad y(x_0)=y_0
                \end{align}
                meameans to find a differentiable function $y=y(x)$ that satisfies both conditions in Eq.\eqref{ODE with initial cond.} 
                on some interval containing $x_0$.
        \end{itemize}

    \section{Integrals as General \& Particular Solutions}
        \begin{itemize}
            \item As an especial case, if $f$ in Eq.\eqref{ODE with initial cond.} 
                does not actually involve the dependent variable $y$, so
                \coloredeq{ODE without y}{\frac{dy}{dx} = f(x)}
                then we need only to take the integral of Eq.\eqref{ODE without y}, so 
                \coloredeq{ODE-integral solution}{y(x) = \int f(x)\, dx + C}
            \item Eq.\eqref{ODE-integral solution} is a \textbf{general solution} of Eq.\eqref{ODE without y}.
            \item When applying an initial condition ($y_0=y(x_0)$), we get a \textbf{particular solution}.
            \item This idea extends to second-order differential equations of the special form:
                \coloredeq{secon-order De}{\frac{d^2y}{dx^2} = g(x)}
                in which the function g on the right-hand side involves neither the dependent
                variable $y$ nor its derivative $dy/dx$.                
        \end{itemize}

    \section{Separable Equations}
        \begin{itemize}
            \item In general, a first-order differential equation is called \textit{separable} 
            provided that $f(x,y)$ can be written as the product of a function of $x$ and a function of $y$:
            \coloredeq{general separable ODE}{\frac{dy}{dx} = f(x,y)=g(x)k(y)=\frac{g(x)}{h(y)}}
            where $h(y)=1/k(y)$
            \item So we can write informally
            \begin{align} \label{separable ODE:1}
                h(y)dy = g(x)dx
            \end{align}
            \item Which we can integrate both side as
            \coloredeq{separable ODE solution}{\int h(y)\,dy = \int g(x)\,dx}
            provided that both antiderivatives of $h(y)$ and $g(x)$ can be found.
        \end{itemize}

### 2022-08-04 latex/MATH208/chapters/chapter2.tex
add LaTeX files

\chapter{Linear Systems and Matrices} % (fold)
\label{cha:Linear Systems and Matrices}
    \section{Introduction to Linear Systems}
        Consider the following equation 
        \[ ax + by  +cz = d \qquad \text{or}\, ax + by = c \]
        each one is called a \textbf{linear equation}; because the \textbf{variables} are power of one.

        \bulletpar A \textbf{system} of linear equations (or \textbf{linear system}) is a finite collections of linear equations involving ceratin variables 
        (\textit{unknowns}). for examole, \[ a_1 x + b_1 y = c_1 \\ a_2 x + b_2 y = c_2\] which is a linear system in two unkowns.
        A \textbf{solution} is a pair of $(x, y)$ that satisfies \textit{both} equations \textit{simultaneously}.

        \begin{itemize}
            \item A \textbf{consistent} systems have at least one solution.
            \item A \textbf{inconsistent} systems have no solution.
        \end{itemize}

        \noindent
        For a linear system, we havr three possiblities:
        \begin{enumerate}
            \item exactly on solution.
            \item no solution.
            \item infinitely many solutions.
        \end{enumerate}

        \paragraph{The Method of Eliminations}
            \begin{enumerate}
                \item Multiply one equation with nonzero number.
                \item Interchange two equations.
                \item Add a constant multiple of one equation to another equation.
            \end{enumerate}

        \paragraph{A Differential Equation}
            Consider this equation 
            \begin{align}
                \label{eq:differential application:1}
                y(x) = Ae^{nx} + Be^{-nx}
            \end{align} 
            after differentiation we get $$ y'(x) = nAe^{nx} - n Be^{-nx} $$
            also
            $$ y''(x) = n^2 A e^{nx} + n^2 B e^{-nx} = n^2 y(x)$$
            Thus Eq.\eqref{eq:differential application:1} is a solution to the differential equation
            \begin{align}
                y'' - n^2y = 0 
            \end{align}
            to find the constant $A$ and $B$ we need to solve a linear systme after applying the initial condition.

    \section{Matrices and Gaussian Elimination}
        A general system of $m$ linear equation in $n$ variables $x_1, x_2, \dots, x_n$ written in form 
        \begin{align}
            \begin{aligned}
                \label{eq:general system}
                a_{11}x_1 + a_{12}x_2 + , \dots + a_{1n}x_n &= b_1 \\
                a_{21}x_1 + a_{22}x_2 + , \dots + a_{2n}x_n &= b_2 \\
                                                            &\vdots \\
                a_{m1}x_1 + a_{m2}x_2 + , \dots + a_{mn}x_n &= b_m \\
            \end{aligned}
        \end{align}
        The \textbf{coefficient matrix} of this linea system is written in $m \times n$ matrix 
        \begin{align*}
            A = \begin{bmatrix}
                a_{11} & a_{12} & \dots & a_{1n} \\
                a_{21} & a_{22} & \dots & a_{2n} \\
                \vdots & \vdots & \ddots & \vdots \\
                a_{m1} & a_{m2} & \dots & a_{mn}
            \end{bmatrix}
        \end{align*}
        For large number of entries we can denote $a_{ij}$ as $a(i, j)$. The \textbf{constant matrix} is 
        \begin{align*}
            b = \begin{bmatrix}
                b_1 \\ b_2 \\ \vdots \\ b_m
            \end{bmatrix}
        \end{align*}
        This is called a \textbf{vector} (\textbf{column}) matrix. Combining both matrices we get a $m \times (n+1)$ \textbf{augmented matrix}
        \begin{align*}
            [A\, b] = \begin{bmatrix}
                a_{11} & a_{12} & \dots & a_{1n} & b_1 \\
                a_{21} & a_{22} & \dots & a_{2n} & b_2 \\
                \vdots & \vdots & \ddots & \vdots & \vdots \\
                a_{m1} & a_{m2} & \dots & a_{mn} & b_m
            \end{bmatrix}
        \end{align*}

        \paragraph{Elemnetary Row Operations} % (fold)
        \label{par:Elemnetary Row Operations}
        As we did before for solving a linear sustem, we can generalize that for augmented matrices as 
        \coloredeq{elemetary row operations}{\parbox[t]{0.8\linewidth}{
            \textbf{Elemnetary Row Operations}
            \begin{enumerate}
                \item Multiply any row by a nonzero number.
                \item Interchange two rows.
                \item Add a constant multiple of one row to another row.
            \end{enumerate}
        }}

        \coloredeq{Row-Equivalent Matrices}{\parbox[t]{0.7\linewidth}{
            \textbf{\color{g2}Row-Equivalent Matrices}\\
            Two matrices are called \textbf{tow eauivalent} if one can be obtained from the pther by a finite sequence of elementary row opearations
        }}

        \coloredeq{theorem1}{\parbox[t]{0.7\linewidth}{
            \textbf{\color{g2}THEOREM1:  Equivalent Systems and Equivalent Matrices}\\ \newline
            If the augmneted coefficient matrices of two systems are row equivalent, then the two systems have the same solution set.
        }}
        % paragraph Elemnetary Row Operations (end)

        \paragraph{Echelon Matrices} % (fold)
        \label{par:Echelon Matrices}
        Let us see how sould the augmented matrix look after some elementry row operations,

        \coloredeq{theorem1}{\parbox[t]{0.8\linewidth}{
            \textbf{\color{g2}DEFINITION:  Echelon Matrix}\\ \newline
            The matrix $E$ is called an \textbf{echelon matrix} provided it has the following two properties:
            \begin{enumerate}
                \item Every row that consists of entirely zeros (if any) liew \textit{beneath} every row that contains a nonzero element.
                \item In each  row contains nonzero element, the \textit{first} nonzero element lies strictly to the right of the first nonzero element in the preceding 
                row (if there is a preseding row).
            \end{enumerate}
        }}
        % paragraph Echelon Matrices (end)

% chapter Linear Systems and Matrices (end)

### 2022-08-04 latex/PHYS210/chapters/chapter2.tex
add LaTeX files

\chapter{Complex Numbers} \label{ch:complex-numbers}
        \section{Introduction}
            The \textit{imaginary} number is written as $i=\sqrt{-1}$. We use the term \textit{complex numbe}r to 
            mean any one of the whole set of numbers, \textit{real}, \textit{imaginary}, or combinations of the two
            \begin{itemize}
                \item The complex number can araise from the quadratic equation \coloredeq{eq:quad}{az^2+bz+c=0}
                \item Where $z$ is a unknown variable and its solution is 
                \coloredeq{eq:quad-sol}{z=\frac{-b\pm\sqrt{b^2-4ac}}{2a}}
                if the \textit{discriminant} $d=b^2-4ac<0$, $z$ will be a complex number.
                \item Notice the pattern if $i$ to power of some numbers \coloredeq{eq:i-short}{i^2=-1,\, i^3=-i,\, i^{4n}=1}
                \item A \textit{complex} number such $3+5i$ has two parts a \textit{real part} (here, $3$) 
                and an \textit{imaginary part} $(5)$.
            \end{itemize}
        \section{The Complex Plane}
            \begin{itemize}
                \item The rectangular coordinates representation for a complex number in the form 
                \colorbox{c1}{$x+yi$} is $(x, y)$.
                \item In the polar coordinates recall that \coloredeq{eq:polar}{x&=r\cos{\theta},\\ y&=r\sin{\theta}}
                \item Then we have (by euler formula) assigned to $z$. All $\theta$ are in \textit{radian}.
                \coloredeq{eq:polar euler z}{z=x+yi=r(\cos{\theta}+i\sin{\theta})=re^{i\theta}}
                \item The \textit{modulus} or \textit{absolute value} of $z$ is
                \coloredeq{eq:abs-z}{|z|=r=\sqrt{x^2+y^2}=\sqrt{z\bar{z}}}
            \end{itemize}
        \section{Complex Algebra}
            \begin{itemize}
                \item Note that in Eq:\eqref{eq:abs-z} $\ge0$, always real.
                \item The \textit{conjugate} of $z$ is found by substituting $\theta=-\theta$ in Eq:\eqref{eq:polar euler z}
                \coloredeq{eq:conj-z}{\bar{z}=r(\cos{\theta}-i\sin{\theta})=re^{-i\theta}}
                \item Inotherwords,anyequation involving complex numbers is really two equations involving real numbers.
            \end{itemize}
        \paragraph{Complex Equations}
        In other words, any equation involving complex numbers is really two equations involving real numbers.

        \section{Complex Infinite Series}
            \begin{itemize}
                \item The partial sums of a series of complex numbers will be complex numbers.
                \item It can be written as \colorbox{c1}{$S_n=X_n+Y_ni$} where $X_n$ and $Y_n$ are real.
                \item \textit{Convergence} is defined just as real series: if $S_n$ approaches a limit
                ($S=X+Yi$) as $n\to\infty$, we call the series convergent and call $S$ its sum.
                \item This means that $X_n \to X$ and $Y_n \to Y$.
                \item It can be proved that an \textit{absolutely convergent series} converges, 
                recall that Eq:\eqref{eq:abs-z} is positive term.
                \item $\sum_{n=0}^\infty \, z^n$ ia a geomtric series, with ratio $=z$ 
                and convergent when $|z|<1$.
            \end{itemize}
            \paragraph{$\bigstar$} Thus any of the tests given in Chapter:\ref{ch:series} for convergence of 
            series of positive terms may be used here to test a \textit{complex series} for \textit{absolute convergence}.

        \section{Complex Power Series}
            They are in form of \coloredeq{eq:general z power series}{\sum \, a_nz^n} where $z=x+yi$ 
            and $a_n$ are complex numbers.
            \begin{itemize}
                \item Note that Eq:\eqref{eq:general z power series} includes the \textit{real series} 
                as a special case when $y=0$.
            \end{itemize}

        \section{Euler's Formula}
            To derive Euler's Formula, for real $\theta$ we have from Chapter:\ref{ch:series}
            \coloredeq{eq:sin theta}{\sin{\theta}&=\dots\\
            \cos{\theta}&=\dots}
            thus we have \coloredeq{eq:euler formula}{e^{i\theta}=\cos{\theta}+i\sin{\theta}}

        \section{Powers and Roots of Complex Numbers}
            \begin{itemize}
                \item The $n$th power (and $n$th root) of $z$ is given by 
                \coloredeq{eq:nth power of z}{z^n=(re^{i\theta})^n=r^ne^{ni\theta}}
                \item When $r=1$ Eq:\eqref{eq:nth power of z} becomes \textbf{DeMoivre’s theorem}
                \coloredeq{eq:DeMoivere}{(e^{i\theta})^n=(\cos{\theta}+i\sin{\theta})^n=\cos{n\theta}+i\sin{n\theta}}
            \end{itemize}

        \section{The Exponential and The Trigonometric}
            \begin{itemize}
                \item Although we have already defined $e^z$ by a power series $e^z=\sum_{n=0}^\infty \, \frac{z^n}{n!}$, 
                it is worth while to write it in another form.
                \coloredeq{eq:e^z power series}{e^z=e^{x+yi}=e^xe^{yi}=e^x(\cos{y}+i\sin{y})}
                \item We have already seen that there is a close relationship [Euler’s formula \eqref{eq:euler formula}]
                between complex exponentials and trigonometric functions of real angles. It is useful 
                to write this relation in another form.
                \item from Eq\eqref{eq:euler formula} we can rearrange it to
                \coloredeq{eq:euler trig of theta}{\sin{\theta}&=\frac{e^{i\theta}-e^{-i\theta}}{2i}\\
                \cos{\theta}&=\frac{e^{i\theta}+e^{-i\theta}}{2}}
                \item It can be shown that in Eq\eqref{eq:euler trig of theta} $\theta$ can be $z$ a complex number.
                \item The rest of the trigonometric functions of $z$ are defined in the usual way in terms of these.

                $\bigstar$ If $z$ is a complex number, $\sin{z}$ and $\cos{z}$ can have \textit{any} value we like.
            \end{itemize}

        \section{Hyperbolic Functions}
            \begin{itemize}
                \item Let us look at $\sin{z}$ and $\cos{z}$ for pure imaginary $z$ in Eq\eqref{eq:euler trig of theta}, that is, $z = iy$:
                \begin{align} \label{eq:euler trig of yi}
                    \sin{yi}&=\frac{e^{-y} - e^{y}}{2i} = i \frac{e^y-e^{-y}}{2}\\
                    \cos{yi}&=\frac{e^{-y} + e^{y}}{2} = \frac{e^y + e^{-y}}{2}
                \end{align}
                \item The real functions on the right have special names because these particular 
                combinations of exponentials arise frequently in problems. They are called the hyperbolic. 
                Their definitions for all $z$:
                \coloredeq{eq:z hyperbolic sin&cos} 
                {\sinh{z}&=\frac{e^{z} - e^{-z}}{2}\\
                \cosh{z}&=\frac{e^{z} + e^{-z}}{2}}
                \item The other hyperbolic functions are definded as in normal trig. functions.
                \item Thus from Eq\eqref{eq:euler trig of yi} we have:
                \coloredeq{eq:trig-hyper equivlent}{\sin{yi} &=i \sinh{y}\\ \cos{yi} &= \cosh{y}}
            \end{itemize}

### 2022-08-04 latex/PHYS210/chapters/chapter3.tex
add LaTeX files

\chapter{Linear Algebra} \label{ch:linear algebra}
        \section{Introduction}
        \begin{itemize}
            \item problems in various fields of science and mathematics involve the solution of sets of linear equations.
            Suppose you have solved two simultaneous linear equations and have found $x=2$ and $y=-3$. We can think of
            $x=2$, $y=-3$ as the point $(2,-3)$ in the $(x,y)$ plane. Since two linear equations represent two straight lines,
            the \emph{solution} is then the \textit{point} of intersection of the lines.

            \item The language of vectors is very useful in studying sets of simultaneous equations. 
            \textit{Quantities} such as the velocity of an object, which have both \emph{magnitude} and \emph{direction}. 
            Such quantities are called \textbf{vectors}; contrast them with such quantities as mass,
            which have \textit{magnitude} only and are called \textbf{scalars}.
            \item Vector formulas are independent of the choice of coordinate system.
            \item A vector equation in two dimensions is equivalent to two component equations.
        \end{itemize}

        \section{Matrices: Row Reduction}
            \begin{itemize}
                \item A matrix (plural: matrices) is just a rectangular array of quantities, such as
                \begin{align} \label{matrix example}
                    A= \begin{bmatrix} 
                        1 & 5 & -2 \\
                        -3 & 0 & 6 
                    \end{bmatrix}
                \end{align}
                \item The $A$ letter does not have a numerical value; it simply stands for the array.
                \item To indicate a number in the array, we will write $A_{ij}$ where $i$ is the 
                \textit{row} number and $j$ is the \textit{column} number.
                \item We will call a matrix with $m$ rows and $n$ columns as \textit{$m$ by $n$ matrix}.
                \item \textbf{Transpose of a Matrix} We write:
                \begin{align} \label{transpose example}
                    A= \begin{bmatrix} 
                    1 & -3 \\ 
                    5 & 0 \\
                    -2 & 6
                    \end{bmatrix}
                \end{align}
                and call $A^T$ the transpose of the matrix A in Matrix\eqref{matrix example}.
                \item To transpose a matrix, we simply write the rows as columns.
                \item Not that, $(A^T)_{ij} = A_{ji}$
                \item Consider the set of equations of:
                \begin{align}\label{set of equations}
                    \begin{cases}
                        2x - z =2\\
                        6x + 5y + 3z = 7\\
                        2x - y = 4
                    \end{cases}
                \end{align}
                \item Let's agree in organizing them in \textit{standard form}, each column be for a specific variable.
                \item There are three markable matricies to extract from \eqref{set of equations}; 
                first one is the \textit{matrix of the coefficients}
                \begin{align} \label{matrix of the coefficiants}
                    M= \begin{bmatrix}
                        2 & 0 & -1 \\
                        6 & 5 & 3 \\
                        2 & -1 & 0
                    \end{bmatrix}
                \end{align}
                Also, there are $3\times1$ matrices, $r$ and $k$.
                \begin{align}
                    r= \begin{bmatrix}
                        x \\ y \\ z
                    \end{bmatrix}, \quad
                    k= \begin{bmatrix}
                        2 \\ 7 \\ 4
                    \end{bmatrix}
                \end{align}
                \item The Eqs\eqref{set of equations} can be written in matrix as $Mr=k$.
                \item Now we want to write Eqs\eqref{set of equations} in an \textit{augmented matrix}
                \begin{align} \label{augmented matrix}
                    A= \begin{bmatrix}
                        2 & 0 & -1 & 2 \\
                        6 & 5 & 3 & 7 \\
                        2 & -1 & 0 & 4
                    \end{bmatrix}
                \end{align}
                and then solve it by the \textit{row reduction} with the following \textit{elementary row operations}:
                \begin{align}  \label{elementry row operations}
                    \coloredpar{
                    \begin{enumerate}
                        \item Interchange two rows.
                        \item Multiply (or divide) a row by a (nonzero) constant.
                        \item Add a multiple of one row to another; this includes 
                        subtracting, that is, using a negative multiple.
                    \end{enumerate}}
                \end{align}
                \item If the last row in a rediced augmented matrix is like $0 \times z=5$, which cannot be 
                true for any finite number on $z$, then it is called \textit{inconsistent}.
                \item \textbf{Rank of a Matrix} The number of nonzero rows remaining when a matrix has been 
                row reduced is called the rank of the matrix.
                \begin{align}
                    \coloredpar{
                        \begin{enumerate}
                            \item If $(\rank M)<(\rank A)$, the equations are inconsistent and there is no solution.
                            \item If $(\rank M)=(\rank A)=n$(\# of unknowns), there is one solution.
                            \item If $(\rank M)=(\rank A)=R<n$, then $R$ unknowns can be found in terms of the remaining $n-R$ unknowns.
                        \end{enumerate}}
                \end{align}
            \end{itemize}

        \section{Determinants; Carmer's Rule}
            For a square matrix, however, there is a useful number called the determinant of the matrix.
            \begin{itemize}
                \item For a $2\times2$ Matrix:
                \begin{align} \label{det of 2by2 matrix}
                    A= \begin{bmatrix}
                        a & b \\
                        c & d
                    \end{bmatrix}, \quad
                    det\, A = \begin{vmatrix}
                        a & b \\
                        c & d
                    \end{vmatrix} =ad-bc
                \end{align}
                \item If we remove one row and one column from a determinant of order $n$, 
                    we have a determinant of order $n-1$.
                \item When removing the \textit{row} and \textit{column} containing the element $a_{ij}$ and call the 
                    remaining determinant $M_{ij}$, which is called the \textbf{minor} of $a_{ij}$.
                \item The \textbf{cofactor} of $a_{ij}$ is: \coloredeq{eq:cofactor of a_ij}{C_{ij}=(-1)^{i+j}M_{ij}}
                \item In general, for $n \times n$ matrix we have,
                \begin{align}
                    \label{eq:general det of nXn matrix}
                    \det{A} = \sum_{j=1}^n\, a_{ij}C_{ij}, \quad \text{where $j$ is fixed}
                \end{align}
                \item The signs goes like $\begin{vmatrix}
                    + & - &  \\
                    - & + &  \\
                    &   &  \ddots
                \end{vmatrix}$.
            \end{itemize}

            \paragraph{$\bigstar$} The value of a \textit{determinant}: Multiply each element of one row (or one column) 
                by its cofactor and add the results.

            \textbf{Useful Fact About Determinants}:
            \begin{align}
                \coloredpar{\begin{enumerate}
                    \item If each element of one row (or one column) of a determinant is multiplied by a number $k$, 
                    the value of the determinant is multiplied by $k$.
                    \item The value of a determinant is zero if
                    \begin{enumerate}
                        \item all elements of one row (or column) are zero; or if
                        \item two rows (or two columns) are identical; or if
                        \item two rows (or two columns) are proportional.
                    \end{enumerate}
                    \item If two rows (or two columns) of a determinant are interchanged, 
                    the value of the determinant changes sign.
                    \item The value of a determinant is unchanged if
                    \begin{enumerate}
                        \item rows are written as columns and columns as rows; or if
                        \item we add to each element of one row, $k$ times the corresponding element of another row, 
                        where k is any number (and a similar statement for columns).
                    \end{enumerate}
                \end{enumerate}}
            \end{align}

            \paragraph{Carmer's Rule} This is a formula in terms of determinants for the solution of $n$ linear equations 
            in $n$ unknowns when there is exactly one solution.\\
            Let's start with the following equations:
            \begin{align} \label{eq:equations set example}
                \begin{cases}
                    a_1x + b_1y = c_1 \\
                    a_2x + b_2y = c_2
                \end{cases}
            \end{align}
            If we multiply the first equation by $b_2$, the second by $b_1$, and then subtract the
            results and solve for x, we get if $(a_1b_2 - a_2b_1 \neq 0)$
            \begin{align} \label{eq:equations set example:1}
                x = \frac{c_1b_2 - c_2b_1}{a_1b_2 - a_2b_1}, \quad 
                y = \frac{a_1c_2 - a_2c_1}{a_1b_2 - a_2b_1}
            \end{align}
            Using the definition \eqref{det of 2by2 matrix} we can write Eq.\eqref{eq:equations set example:1}
            \begin{align}
                x = 
                \frac{\begin{vmatrix}
                    c_1 & b_1 \\
                    c_2 & b_2
                \end{vmatrix}}
                {\begin{vmatrix}
                    a_1 & b_1 \\
                    a_2 & b_2
                \end{vmatrix}} = \frac{1}{D}
                \begin{vmatrix}
                    c_1 & b_1 \\
                    c_2 & b_2
                \end{vmatrix}, \quad\quad
                y = 
                \frac{\begin{vmatrix}
                    a_1 & c_1 \\
                    a_2 & c_2
                \end{vmatrix}}
                {\begin{vmatrix}
                    a_1 & b_1 \\
                    a_2 & b_2
                \end{vmatrix}} = \frac{1}{D}
                \begin{vmatrix}
                    a_1 & c_1 \\
                    a_2 & c_2
                \end{vmatrix}, \quad D \neq 0
            \end{align}
            \paragraph{$\bigstar$} To remember it, the denominator for both $x$ and $y$ is the 
            \textit{determinant of the coefficients},for the numerator for $x$ replace its column with 
            the $k$ matrix (constants column), and do so for $y$.

            \begin{align} \label{Carmer's Rule for solving n eqs}
                \coloredpar{
                    This method of solution of a set of linear equations is called Cramer’s rule. 
                    It may be used to solve $n$ equations in $n$ unknowns if $D \neq 0$; the solution then consists of 
                    one value for each unknown. The denominator determinant D is the $n$ by $n$ determinant of the 
                    \textit{coefficients} when the equations are arranged in standard form. The numerator determinant for 
                    each unknown is the determinant obtained by replacing the column of coefficients of that 
                    unknown in D by the constant terms from the right-hand sides of the equations. Then to find 
                    the unknowns, we must evaluate each of the determinants and divide.}
            \end{align}

            \paragraph{Rank of Matrix} Here is another way to find the rank of a matrix.
            A submatrix means a matrix remaining if we remove some rows and/or remove some 
            columns from the original matrix. To find the rank of a matrix, we look at all 
            the square submatrices and find their determinants. The \textit{order of the largest
            nonzero determinant} is the rank of the matrix.

        \section{Vectors}
            \paragraph{}{Notation} We shall indicate a vector by a boldface letter (for example, \textbf{A}) and a 
            component of a vector by a subscript (for example $A_x$ is the $4$ component of \textbf{A}). 
            For handwriting you should write a vector with an arrow (like, $\vec{A}$).

            \paragraph{Magnitude of a Vector} The \textit{length} of the arrow representing a vector $\textbf{A}$
            is called the \textit{length} or the \textit{magnitude} of \textbf{A} (written $|\textbf{A}|$ or $A$) or
            the \textit{norm} of \textbf{A} (written $||\textbf{A}||$). Note the use of A to mean the magnitude of \textbf{A}.

            \paragraph{$\bullet$} By the Pythagorean theorem, we find
            \coloredeq{eq:Pythagorean for vectors}{A = |\vec{A}| = \sqrt{A_x^2 + A_y^2 + A_z^2}}

            \paragraph{Addition of Vectors} There two ways to add vectors: 
            Addition of Vectors
            by the parallelogram law: To find $\vec{A} + \vec{B}$, place the tail of 
            $\vec{B}$ at the head of $\vec{A}$ and draw the vector from the tail of $\vec{A}$ to the head of $\vec{B}$.
            Or by adding their components togather, like $A_x + B_x$ and $A_y + B_y$. They follow:
            \begin{align} \label{vectors properties}
                \vec{A} + \vec{B} = \vec{B} + \vec{A}, \quad \text{(commutative law for addition)};\\
                (\vec{A} + \vec{B}) + \vec{C} = \vec{A} + (\vec{B} + \vec{C}), \quad \text{(associative law for addition)}.
            \end{align}
            \paragraph{$\bullet$} In other words, vectors may be added together by the usual laws of algebra.
            They can be multiplied by a number. Ane we can define vectors substraction so:
            $$ \vec{A} - \vec{B} = \vec{A} + (-\vec{B}) $$

            \paragraph{$\bullet$} The \textbf{zero vector} is a vector of zero magnitude; its 
            components are all zero and it does not have a direction.\\
            A vector of length or magnitude 1 is called a \textbf{unit vector}. 
            Then for any $\vec{A} \neq 0$, the vector $\vec{A}/|\vec{A}|$ is a unit vector.

            \begin{wrapfigure}{r}{0.2\linewidth}
                \includegraphics[width=0.9\linewidth]{figures/basis vectors.png}
                \caption{The unit basis vectors in a rectangular system.}
                \label{fig:basis vectors}
            \end{wrapfigure}

            \paragraph{Vectors in Terms of Components} We consider a set of rectangular axes as in Fig.\eqref{fig:basis vectors}
            let $\ihat $ be a unit vector in the positive $x$ direction; $\jhat $ and $\khat $
            in $y$ and $z$ positive direction. For $\vec{A}$
            \begin{align}
                \vec{A} = A_x\ihat  + A_y\jhat  + A_z\khat 
            \end{align} 
            The vectors $\ihat , \jhat , \khat $ are called \textit{unit basis vectors}.

            \paragraph{Multiplication of Vectors} There are two kinds of product of two vectors. 
            One, called the \text{scalar} product (or \textbf{dot product} or \textbf{inner product}), 
            gives a result which is a scalar; the other, called the \textit{vector product} 
            (or \textbf{cross product}), gives a vector answer.

            \paragraph{Scalar Product} For vectors $\vec{A}$, and $\vec{B}$, and the angle $\theta$($\le 180^\circ$) between them,
            we have:
            $$ \vec{A} \cdot \vec{B} = (\ihat A_x +\jhat A_y +\khat A_z) \cdot (\ihat B_x +\jhat B_y +\khat B_z) $$
            \coloredeq{eq:dot product}{\vec{A} \cdot \vec{B} = |\vec{A}||\vec{B}|\cos{\theta} = A_xB_x + A_yB_y + A_zB_z}
            The dot product in Eq.\eqref{eq:dot product} holds the commutative law.\\
            And the distributive law.

            \paragraph{$\bullet$} A vector dot itself gives its magnitude:
            $$ \vec{A} \cdot \vec{A} = |\vec{A}|^2\cos{0} = |A|^2 = A^2 $$

            \paragraph{Perpendicular and Parallel Vectors}  % (fold)
            \label{par:Perpendicular and Parallel Vectors}
            If two vectors are perpendicular, then $\cos{\theta} = 0$; thus
            \coloredeq{eq:Perp. vectors with dot product}{
                \begin{aligned}
                    \vec{A} \cdot \vec{B} &= A_xB_x + A_yB_y + A_zB_z = 0, && \text{if $\vec{A}$ and $\vec{B}$
                    are perpendicular vectors.} \\
                    \frac{A_x}{B_x} &= \frac{A_y}{B_y} = \frac{A_z}{B_z}, && 
                    \text{if $\vec{A}$ and $\vec{B}$ are perpendicular vectors.}
                \end{aligned}
            }
            % paragraph Perpendicular and Parallel Vectors (end)
            \paragraph{Vector Product} % (fold)
            \label{par:Vector Product}
            The vector or cross product of $\vec{A}$ and $\vec{B}$ is written $\vec{A} \times \vec{B}$. By
            definition, $\vec{A} \times \vec{B}$ is a vector whose magnitude and direction are given as follows:

            \begin{align} \label{eq:cross product magintude}\coloredpar{
                $$ |\vec{A} \times \vec{B}| = |\vec{A}||\vec{B}|\sin{\theta}, $$
                The direction of $\vec{A} \times \vec{B}$ is perpendicular to the 
                plane of $\vec{A}$ and $\vec{B}$, using \textit{right-hand rule}}
            \end{align}

            \paragraph{Intresting properties:}
            \coloredeq{eq:cross product properties}
            {\begin{aligned}
                \vec{A} \times \vec{B} &= - \vec{B} \times \vec{A}  && \text{not commutative,}\\
                \vec{A} \times \vec{B} &= 0 \quad && \parbox[c]{0.4\linewidth}{if $\vec{A}$ and $\vec{B}$ are parallel or antiparallel,}\\
                \vec{A} \times \vec{A} &= 0 \quad && \text{for any $\vec{A}$}
            \end{aligned}}
            % paragraph Vector Product (end)

            \paragraph{$\bigstar$} A good way to remember the basis-unit vectors cross-product is 
            to write them cyclically. Reading around the circle counterclockwise (positive $\theta$ direction), 
            we get the positive products (for example, $\ihat  \times \jhat  = \khat $); reading 
            the other way we get the negative products (for example, $\ihat  \times \khat  = -\jhat $). This works with 
            the \textit{right-haned systems}.

            To write A × B in component form we need the distributive law, namely
            \begin{align} \label{eq:cross product distributive law}
                \vec{A} \times (\vec{B} + \vec{C}) = \vec{A} \times \vec{B} + \vec{A} \times \vec{c}
            \end{align}
            Thus we have:
            \coloredeq{eq:cross product vector}{
                \begin{aligned}
                    \vec{A} \times \vec{B} &= 
                    (\ihat A_x +\jhat A_y +\khat A_z) \times (\ihat B_x +\jhat B_y +\khat B_z)\\
                    &=\ihat (A_yB_z - A_zB_y) + \jhat (A_zB_x - A_xB_z) + \khat (A_xB_y - A_yB_x)\\
                    &= \begin{vmatrix}
                        \ihat  & \jhat  & \khat  \\
                        A_x & A_y & A_z \\
                        B_x & B_y & B_z
                    \end{vmatrix}
                \end{aligned}
            }

        \section{Lines and Planes}
            In analytic geometry a point is a set of three coordinates $(x,y,z)$; we shall think of 
            this point as the \textit{head} of a vector $\vec{r} = \ihat x + \jhat y + \khat z$ with 
            \textit{tail} at the \texttt{origin}.
            In two dimensions, we write the equation of a \textit{straight line} through $(x_0,y_0)$ with slope $m$ as
            \begin{equation}
                \label{eq:straight line}
                \frac{y-y_0}{x-x_0} = m
            \end{equation}

            \paragraph{$\bullet$} Suppose, instead of the slope, we are given a vector in the direction 
            of the line, say $\vec{A} = \ihat a+\jhat b$. Then the line through $(x_0,y_0)$ and parallel to
            $\vec{A}$ we can write its equation. If we have two points on the line from $(x_0,y_0)$
            to any point $(x,y)$, then the vector $\vec{r} - \vec{r_0}$ with components $x-x_0 $ and $y-y_0$:
            \begin{align}
                \label{eq:vector bewtween two points}
                \vec{r} - \vec{r_0} = (x-x_0)\ihat  + (y-y_0)\jhat 
            \end{align}
            Since this vector is parallel to $\vec{A}$, then their components are proportional, so (for $a, b\neq 0$):
            \begin{align}
                \label{eq:line equation:2}
                \frac{x-x_0}{a} = \frac{y-y_0}{b} && or\quad \frac{y-y_0}{x-x_0} = \frac{b}{a}
            \end{align}
            The equation is for a given \textbf{line}, see tht Eq.\eqref{eq:straight line} and 
            Eq.\eqref{eq:line equation:2} are identical.

            \paragraph{$\bullet$} Since $\vec{r} -\vec{r_0}$ and $\vec{A}$ are parallel, thus they are only differen 
            by a factor of $t$, so
            \begin{align}
                \label{eq:r and A equivlent}
                \vec{r} -\vec{r_0} = \vec{A} t \quad or \quad \vec{r} = \vec{r_0} + \vec{A} t
            \end{align}
            Then theier \textit{components form}:
            \begin{align}
                \begin{aligned}
                    x - x_0 = at \\
                    y - y_0 = bt
                \end{aligned} \quad or \quad
                \begin{aligned}
                    x = x_0 + at \\
                    y = y_0 + bt
                \end{aligned}
            \end{align}
            Eliminating $t$ yields the equation of the line in Eq.\eqref{eq:line equation:2}

            \paragraph{$\bullet$} In three dimensions, we want the equations of
            a straight line through a given point $(x_0,y_0,z_0)$ and parallel to a given vector 
            $\vec{A} = a\ihat  + b\jhat  + c\khat $. If $(x,y,z)$ is any point on the line, the 
            vector joining $(x_0,y_0,z_0)$ and 
            $(x,y,z$) is parallel to $\vec{A}$. Then,
            \coloredeq{eq:symmetric line in 3D}{
                \begin{aligned}
                    \frac{x-x_0}{a} &= \frac{y-y_0}{b} = \frac{z-z_0}{c} && \text{$a, b, c \neq 0$}\\
                    \frac{x-x_0}{a} &= \frac{y-y_0}{b} , z=z_0 && \text{if say $c=0$}
                \end{aligned} 
            }
            \begin{wrapfigure}{r}{0.2\linewidth}
                \label{fig:parametric line}
                \includegraphics[width=0.9\linewidth]{figures/parametric line.png}
                \caption{}
            \end{wrapfigure}
            The paramtric equations for a line, see Fig.\eqref{fig:parametric line}, from Eq.\eqref{eq:symmetric line in 3D}
            \coloredeq{eq:paramtric equations}{
                \vec{r} = \vec{r_0} + \vec{A}t \quad or 
                \begin{aligned}
                    \begin{cases}
                        x = x_0 + at \\
                        y= y_0 + bt \\
                        z = z_0 +ct
                    \end{cases}
                \end{aligned} 
            }

            \paragraph{$\bullet$} Suppose we want the equation of a 
            straight line $L$ through the point $(x_0, y_0)$ and \texttt{perpendicular} to a given vector $\vec{N} = a\ihat +b\jhat $.
            Let the vector in Eq.\eqref{eq:vector bewtween two points} lies along the line but to be perpendicular
            to $\vec{N}$. \textit{Recall} from \eqref{eq:Perp. vectors with dot product}, we have 
            $(\vec{r}-\vec{r_0}) \cdot \vec{N} = 0$, thus their components
            \begin{align}
                \label{eq:Perp. line equation}
                a(x-x_0) + b(y-y_0) = 0 \quad or \quad \frac{y-y_0}{x-x_0} =- \frac{a}{b}
            \end{align}
            This is the equation for a line that is perpendicular to $\vec{N}$

            \paragraph{$\bullet$} In 3D, we can use this method to assign an equation for a plne. 
            Suppose that two points on the plane, $(x_0, _0, z_0)$ and $(x, y, z)$ is any point and represented by the
            vector in \eqref{eq:vector bewtween two points}.\\
            If $\vec{N} = a\ihat +b\jhat +c\khat $ is a normal vector to the plane, we have $(\vec{r}-\vec{r_0}) \cdot \vec{N} = 0$, so 
            \coloredeq{eq:3D plane equation}{
                \begin{aligned}
                    a(x - x_0) + b(y - y_0) + c(z - z_0) &= 0,\\
                    ax + by + cz &= d \quad \textit{where} \quad d=ax_0 + by_0 + cz_0
                \end{aligned}
            }

        \section{Matrix Operations}
            In Section 2 we used matrices simply as arrays of numbers. Now we want to go farther into the subject 
            and discuss the meaning and use of multiplying a matrix by a number and of combining matrices by addition, 
            subtraction, multiplication, and even (in a sense) division.

            \paragraph{Matrix Equations} % (fold)
            \label{par:Matrix Equations}
            Two matrices are only equal if they are identical, for example,
            \begin{align}
                \begin{bmatrix}
                    w & m \\
                    r & k
                \end{bmatrix} = 
                \begin{bmatrix}
                    4 & 5i \\
                    5 & 0
                \end{bmatrix}
            \end{align}
            then $w=4, \quad m=5i, \quad r=5, \quad k=0$.\\ Remeber this consept in the equation 
            $z = x + iy = 2 - 3i$ is equivalent to the two real equations $x = 2$, $y = -3$; \\
            a vector equation in three dimensions is equivalent to three component equations.
            % paragraph Matrix Equations (end)
            \paragraph{Multiplication of a Matrix by a Number} % (fold)
            \label{par:Multiplication of a Matrix by a Number}
            We can write a vector $\vec{A} = a\ihat  + b\jhat  + c\khat $ in matrix-like way,
            \begin{equation}
                \begin{aligned}
                \label{eq:vector in matrix}
                A = \begin{bmatrix}
                    a\\
                    b\\
                    c
                \end{bmatrix}\, \text{a column matrix or column vector,} \\
                A^T= \begin{bmatrix}
                    a & b & c
                \end{bmatrix}\, \text{called a row matrix or row vector.}
                \end{aligned}
            \end{equation}
            The row matrix $A^T$ is the transpose of the column matrix $A$
            % paragraph Multiplication of a Matrix by a Number (end)

            \paragraph{$\bullet$} Suppose we multiply the vector 
            $\vec{A}$ by a constant $C$, them $C\vec{A} = aC\ihat  + bC\jhat  + cC\khat $ and its matrix becomes
            \begin{align}
                A = \begin{bmatrix}
                    aC & bC & cC
                \end{bmatrix}
            \end{align}
            \paragraph{$\bigstar$}
            Thus when a matrix is multiplied by a number each element is multiplied by it.

            \coloredpar{Multiplying a matrix 
            by a number $k$ means multiplying every element by $k$,\\ But multiplying just one \textit{row} of a determinant
            by $k$ multiplies the determinant by $k$. Thus $\det{kA} = k^2 \det{A}$ for a 2 by 2 matrix.}
            \paragraph{Addition of Matrices} % (fold)
            \label{par:Addition of Matrices}
            When we add vectors algebraically, we add them by components. Matrices are added in the
            same way, but by adding \textit{corresponding elements}. If the two matrices have different $m$ by $n$,
            we say that the sum is undefined or meaningless.
            % paragraph Addition of Matrices (end)

            \paragraph{Multiplication of Matrices} % (fold)
            \label{par:Multiplication of Matrices}
            Let us start by defining the product of two matrices,
            \begin{align}
                \label{eq:multiply two 2X2 matrices}
                AB = \begin{bmatrix}
                    a & b \\
                    c & d
                \end{bmatrix}
                \begin{bmatrix}
                    e & f \\
                    g & h
                \end{bmatrix} =
                \begin{bmatrix}
                    ae+bg & af+bh\\
                    ce+dg & cf+dh
                \end{bmatrix} = C
            \end{align}
            Each row for each column. :“row times column”.
            \coloredeq{eq:matrix multiplication in indeces}{
                \parbox{0.6\linewidth}{The element in row $i$ and column $j$ of the product matrix $AB$ 
                is equal to row $i$ of $A$ times column $j$ of $B$. In index notation}
                 &&
                \begin{aligned}
                    (AB)_{ij} = \sum_k\, A_{ik}B_{kj}
                \end{aligned}
            }
            \coloredpar{The product $AB$ (in that order) can be found if and only if the number of elements 
            in a \textit{row} of $A$ equals the number of elements in a \textit{column} of $B$; the matrices $A$, $B$ in that 
            order are then called \textbf{conformable}. \\(Observe that the number of rows in \textit{A} and of columns 
            in $B$ have nothing to do with the question of whether we can find $AB$ or not.)}

            \coloredeq{eq:commutator}{[A,B]=AB-BA= \text{\textit{commutator} of $A$ and $B$.}}
            % paragraph Multiplication of Matrices (end)

            \paragraph{Zero Matrix} % (fold)
            \label{par:Zero Matrix}
            The \textbf{zero} or \textbf{null} matrix means one with all its elements equal to zero. 
            It is often abbreviated by 0, but we must be careful about this
            % paragraph Zero Matrix (end)

            \paragraph{Identity Matrix or Unit Matrix} % (fold)
            \label{par:Identity Matrix or Unit Matrix}
            This is a square matrix with every element of the main diagonal
            equal to 1 and all other elements equal to zero. For example
            \begin{align}
                \label{identity matrix}
                I = \begin{bmatrix}
                    1 & 0 & 0 \\
                    0 & 1 & 0\\
                    0 & 0 & 1
                \end{bmatrix}
            \end{align}
            In multiplication, a unit matrix acts like the number 1, 
            that is, if $A$ is any matrix and $I$ is the \textit{unit} matrix \textit{conformable} with $A$ in the order 
            in which we multiply, then $IA = AI = A$.
            % paragraph Identity Matrix or Unit Matrix (end)

            \paragraph{Operations with Determinants} % (fold)
            \label{par:Operations with Determinants}
            we multiply determinants the same way we multiply matrices.
            \coloredeq{eq:multiplication of determinants}{\det{AB} = \det{BA} = \det{A} \cdot \det{B}}
            % paragraph Operations with Determinants (end)

            \paragraph{Applications of Matrix Multiplication} % (fold)
            \label{par:Applications of Matrix Multiplication}
            We can now write sets of simultaneous linear equations in a very simple 
            form using matrices. Consider the matrix equation
            \begin{align}
                \label{matrices manipulation:1}
                \begin{bmatrix}
                    2 & 6 & -3 \\
                    7 & 4 & 4 \\
                    2 & 1 & 7
                \end{bmatrix}
                \begin{bmatrix}
                    x\\ y\\ z
                \end{bmatrix} = 
                \begin{bmatrix}
                    5 \\ 4 \\2
                \end{bmatrix}
            \end{align}
            By matrices multiplication,
            \begin{align}
                \label{matrices manipulation:2}
                \begin{bmatrix}
                    2x+6y-3z \\
                    7x+4y+4z \\
                    2x+y+8z
                \end{bmatrix} = 
                \begin{bmatrix}
                    5 \\ 4 \\2
                \end{bmatrix}
            \end{align}
            Recall from \eqref{par:Matrix Equations}, then we get
            \begin{align}
                \label{matrices manipulation:3}
                \begin{cases}
                    2x+6y-3z&=5,\\
                    7x+4y+4z&=4,\\
                    2x+y+8z&=2
                \end{cases}
            \end{align}
            Consequently \eqref{matrices manipulation:1} is the matrix form for the set 
            of equations in \eqref{matrices manipulation:3}, if
            \begin{align}
                \label{matrices manipulation:4}
                M=\begin{bmatrix}
                    2x+6y-3z \\
                    7x+4y+4z \\
                    2x+y+8z
                \end{bmatrix}, \quad
                r=\begin{bmatrix}
                    x\\ y\\ z
                \end{bmatrix}, \quad
                k=\begin{bmatrix}
                    5 \\ 4 \\2
                \end{bmatrix}
            \end{align}
            then we can write the \eqref{matrices manipulation:1} as $Mr=k$ or $\sum_j\, M_{ij}r_i=k_i$.
            % paragraph Applications of Matrix Multiplication (end)

            \paragraph{Inverse of a Matrix} % (fold)
            \label{par:Inverse of a Matrix}
            \textit{Inverse of a Matrix} The reciprocal or inverse of a number $x$ is $x^{-1}$ such that the product 
            $xx^{-1} = 1$. We define the inverse of a matrix $M$ (if it has one) as the matrix $M^{-1}$ such that 
            $MM^{-1}$ and $M^{-1}M$ are both equal to a unit matrix $I$.
            \paragraph{$\bullet$} Note that only square matrices can have inverses 
            (otherwise we could not multiply both $MM^{-1}$ and $M^{-1}M$). Actually, some square matrices do 
            not have inverses either. \\You can see from \eqref{eq:multiplication of determinants} 
            that if $M^{-1}M = I$, then 
            $(\det{M^{-1}})(\det{M}) = \det{I} = 1$. If two numbers have product $= 1$, then neither of them is 
            zero; thus $\det{M} \neq 0$ is a requirement for $M$ to have an inverse.

            \paragraph{$\bullet$} If a matrix has an inverse we say that it is \textbf{invertible}; 
            if it doesn’t have an inverse, it is called \textbf{singular}.
            We can find the inverse of a matrix as
            \coloredeq{eq:inverse of matrix}{M^{-1} = \frac{1}{\det{M}}C^T}
            % paragraph Inverse of a Matrix (end)

        \section{Linear Combinations, Functions, and Opreators}
            \paragraph{$\bullet$}Given two vectors $\vec{A}$ and $\vec{B}$, the vector $3\vec{A} - 2\vec{B}$ is called a \textbf{linear combination} 
            of $\vec{A}$ and $\vec{B}$. In general, a linear combination of $\vec{A}$ and $\vec{B}$ means $a\vec{A} + b\vec{B}$ where $a$ 
            and $b$ are \textit{scalars}. Geometrically, if $\vec{A}$ and $\vec{B}$ have the same \textit{tail} and do not lie 
            along a line, then they determine a plane. All linear combinations of $\vec{A}$ and $\vec{B}$ lie in the plane.

            \paragraph{$\bullet$}It is also true that every vector in the plane can be written as a linear combination of 
            $\vec{A}$ and $\vec{B}$. The vector $\vec{r} = x\hat{i} + y\hat{j} + z\hat{k}$ with tail at the origin 
            (which we used in writing equations of lines and planes) is a linear combination 
            of the unit basis vectors $\ihat$, $\jhat$, $\khat$.

            \coloredeq{eq:A function of a vector}
            {
                \begin{aligned}
                    \parbox{0.5\linewidth}{A function of a vector, $f(\vec{r})$ ia called linear if}
                    \qquad \qquad \qquad \qquad \\
                    f(\vec{r_1} + \vec{r_2}) = f(\vec{r_1}) + f(\vec{r_2}), \quad and \quad
                    f(a\vec{r}) = a f(\vec{r})
                    \\
                    \text{where $a$ is a scalar.}
                \end{aligned}
            }

            \begin{wrapfigure}{r}{0.2\linewidth}
                \includegraphics[width=0.9\linewidth]{figures/triangle vectors sum.png}
                \caption{}
                \label{fig:triangle vectors sum}
            \end{wrapfigure}

            \paragraph{$\bullet$} Note,  $f(\vec{r}) = |\vec{r}|$ is not a linear function, because the length 
            of the sum of two vectors is not in general the sum of their lengths. That is,
            $$ f(\vec{r_1} + \vec{r_2}) = |\vec{r_1} + \vec{r_2}| \neq |\vec{r_1}| + |\vec{r_2}| = 
            f(\vec{r_1}) + f(\vec{r_2})$$ as in Fig.\eqref{fig:triangle vectors sum}.
            Also noticw that we call $y=ms+b$ a linear a equation, however the function
            $f(x)=mx+b$ is not linear by the test \eqref{eq:A function of a vector} 
            (unless $b=0$).

            \paragraph{$\bullet$} Now consider \textbf{vector function} of a vector $\vec{r}$
            \coloredeq{eq:vector function}
            {
                \begin{aligned}
                    \text{$\vec{F}(\vec{r})$ is a linear vector function if}&
                    \\
                    \vec{F}(\vec{r_1}+\vec{r_2}) &= \vec{F}(\vec{r_1}) + \vec{F}(\vec{r_2})
                    \quad and \quad
                    \vec{F}(a\vec{r}) = a\vec{F}(\vec{r})
                    \\
                    \text{where $a$ is a scalar.}
                \end{aligned}
            }

            \paragraph{$\bullet$} Recall that from calculus:
            \begin{align}
                \label{eq:differentiation of sum}
                \begin{aligned}
                    \frac{d}{dx}\left[ f(x) + g(x) \right] &= \frac{d}{dx}f(x) + \frac{d}{dx}g(x) \quad and\\
                    \frac{s}{dx}\left[ kf(x) \right] &= k\frac{d}{dx}f(x)\\
                    \text{where $k$ is a constant}
                \end{aligned}
            \end{align}

            This is silimilar to \eqref{eq:A function of a vector}, then we call $d/dx$ a \textbf{linear opreator}.
            An \textbf{operator} or \textbf{operation} simply means a rule or some kind of instruction telling us what 
            to do with whatever follows it. In other words, a \textit{linear operator} is a \textit{linear function}, so

            \coloredeq{eq:A linear operator}{
                \begin{aligned}
                    \text{$O$ is a linear operator if}& \\
                    O(A+B) &= O(A) + O(B) \qquad and \qquad O(kA) = kO(A)\\
                    \text{where $k$ is a number}
                \end{aligned}
            }

            \paragraph{Matrix Operators, Linear Transformations} % (fold)
            \label{par:Matrix Operators, Linear Transformations}
            Consider the following equation set 
            \begin{align}
                \label{eq:transformation:1}
                \begin{cases}
                    X = ax + by \\
                    Y = cx + dy
                \end{cases}
                \qquad or \qquad 
                \begin{bmatrix}
                    X \\ Y
                \end{bmatrix} = 
                \begin{bmatrix}
                    a & b \\
                    c & d
                \end{bmatrix}
                \begin{bmatrix}
                    x \\ y
                \end{bmatrix}
                \qquad or \qquad 
                R = Mr
            \end{align}
            where $a$, $b$, $c$, and $d$ are constants.
            For every point $(x,y)$, these equations results a point $(X, Y)$.

            If we think of each point of the $(x, y)$ plane being moved to some other 
            point (the origin not being moved), we can call this 
            process a \textbf{mapping} or \textbf{transformation} of the plane into itself. All the information 
            about this \textit{transformation} is contained in the matrix $M$. We say that this matrix is 
            an \textit{operator} which maps the plane into itself.

            \begin{wrapfigure}{r}{0.2\linewidth}
                \includegraphics[width=0.9\linewidth]{figures/transformation1.png}
                \caption{fixed coordinates axes}
                \label{fig:transformation1}
                \includegraphics[width=0.9\linewidth]{figures/transformation2.png}
                \caption{A fixed vector}
                \label{fig:transformation2}
            \end{wrapfigure}

            \paragraph{$\bullet$}Any matrix can be thought of as an \textit{operator} 
            on (conformable) column matrices $r$. From \eqref{eq:A linear operator}, $M$ is a linear operator.

            \paragraph{$\bullet$}We can interpreted Eqs.\eqref{eq:transformation:1} geometrically in two ways. In Fig.\eqref{fig:transformation1}
            The vector $\vec{r}$ has been changed to the vector $\vec{R}$ by the transformation \eqref{eq:transformation:1}.

            \paragraph{$\bullet$}However, in Fis.\eqref{fig:transformation2} \texttt{two} sets of coordinates 
            axes $(x, y)$ and $(x', y')$, and \texttt{one} vector $\vec{r} = \vec{r^\prime}$. The
            transformation can ve written as 
            \begin{align}
                \label{eq:transformation:2}
                \begin{cases}
                    x' = ax + by \\
                    y' = cx + dy
                \end{cases}
                \qquad or \qquad 
                \begin{bmatrix}
                    z'\\ y'
                \end{bmatrix} = 
                \begin{bmatrix}
                    a & b \\
                    c & d
                \end{bmatrix}
                \begin{bmatrix}
                    x \\ y
                \end{bmatrix}
                \qquad or \qquad 
                r' = Mr
            \end{align}
            % paragraph Matrix Operators, Linear Transformations (end)

            \paragraph{Orthogonal Transformations} % (fold)
            \label{par:Orthogonal Transformations}
            This is a special case of a \textit{linear transformation} which preserves the length of a vector.
            In \eqref{eq:transformation:2} is an \textit{orthogonal transformation} if
            \begin{align}
                \label{eq:orthogonal condition}
                x'^2 + y'^2  = x^2 + y^2
            \end{align}
            then either it is reflected or rotated. The matrix \textit{M} of an orthogonal transformation is called an \textbf{orthogonal matrix}.
            The inverse of an orthogonal matrix equals its transpose. \coloredeq{eq:orthogonal matrix}{M^{-1} = M^T, \quad M \text{ is orthogonal}}
            From Eq.\eqref{eq:transformation:2} and \eqref{eq:orthogonal condition}
            \begin{align*}
                x'^2 + y'^2 &= (ax + by)^2 + (cx + dy)^2 \\
                            &=(a^2 +c^2)x^2 +2(ab+cd)xy+(b^2 +d^2)y^2 \equiv x^2 +y^2\\
                        \end{align*}
            Thus we must have $a^2 +c^2=b^2 +d^2=1,\quad ab+cd=0$
            \begin{align*}
                \begin{aligned}
                    MM^T &= \begin{bmatrix}
                        a & c \\ b & d
                    \end{bmatrix}
                    \begin{bmatrix}
                        a & b \\ c & d
                    \end{bmatrix} = 
                    \begin{bmatrix}
                        a^2 +c^2 & ab+cd \\
                        ab+cd & b^2 +d^2
                    \end{bmatrix}
                    \equiv \begin{bmatrix}
                        1 & 0 \\ 0 & 1
                    \end{bmatrix}
                \end{aligned}
            \end{align*}

            Let Eq.\eqref{eq:orthogonal matrix} be $MM^T=I$ and from Eq.\eqref{eq:multiplication of determinants} we have
            \begin{align*}
                \det{M^TM} &= \det{M^T} \det{M}\\
                and \quad det{M^T} &= \det{M}\\
                so\quad \det{M}^2 &= \det{M^TM} = \det{I} = 1
            \end{align*}
            \coloredeq{eq:identity of orthogonal}{\det{M} = \pm 1}

            \paragraph{$\bigstar$} $\det{M} = 1$ corresponds geometrically to a \textit{rotation}, 
            and $\det{M} = -1$ means that a \textit{reflection} is involved.
            % paragraph Orthogonal Transformations (end)

            \begin{wrapfigure}{r}{0.2\linewidth}
                \includegraphics[width=0.9\linewidth]{figures/vectors rotation.png}
                \caption{Vectors rotation with an angle $\theta$}
                \label{fig:vectors rotation}
                \includegraphics[width=0.9\linewidth]{figures/axes rotation.png}
                \caption{Axes rotation}
                \label{fig:axes rotation}
            \end{wrapfigure}

            \paragraph{Rotations in 2 Dimensions} % (fold)
            \label{par:Rotations in 2 Dimensions}
            In Fig.\eqref{fig:vectors rotation} we have two vectors $\vec{R}$  and $\vec{r}$ which is rotated by $\theta$.
            we can write in matrix their transformation as 
            \begin{align}
                \label{eq:vector rotation matrix}
                \begin{bmatrix}
                    X \\ Y
                \end{bmatrix} = 
                \begin{bmatrix}
                    \cos{\theta} & -\sin{\theta} \\
                    \sin{\theta} & \cos{\theta}
                \end{bmatrix} 
                \begin{bmatrix}
                    x \\ y
                \end{bmatrix}
            \end{align}

            \paragraph{$\bullet$} However for axes rotation as in Fig.\eqref{fig:axes rotation} we have 
            \begin{align}
                \label{eq:axes rotation matrix}
                \begin{bmatrix}
                    x' \\ y'
                \end{bmatrix} = 
                \begin{bmatrix}
                    \cos{\theta} & \sin{\theta} \\
                    -\sin{\theta} & \cos{\theta}
                \end{bmatrix} 
                \begin{bmatrix}
                    x \\ y
                \end{bmatrix}
            \end{align}

            \paragraph{$\bullet$} Both Eq.\eqref{eq:vector rotation matrix} and Eq.\eqref{eq:axes rotation matrix} are called \textbf{rotation equations}
            and the matrices that contain $\theta$ are called \textbf{rotation matrices}. Also notice that Eq.\eqref{eq:vector rotation matrix} 
            and Eq.\eqref{eq:axes rotation matrix} are inverses of each other.
            % paragraph Rotations in 2 Dimensions (end)

            \paragraph{Rotations and Reflections in 3 Dimensions} % (fold)
            \label{par:Rotations and Reflections in 3 Dimensions}
            For a vector $ \vec{r} = \langle x, y, z \rangle$. The the following matrix is a rotation about $z-axis$,
            \begin{align}
                \label{eq:rotation about z}
                \begin{bmatrix}
                    \cos{\theta} & -\sin{\theta} & 0 \\
                    \sin{\theta} & \cos{\theta} & 0 \\
                    0 & 0 & 1
                \end{bmatrix}
            \end{align}
            And the following is a rotation and reflection about $xy-plane$
            \begin{align}
                \label{eq:rotationa z and reflection in xy}
                \begin{bmatrix}
                    \cos{\theta} & -\sin{\theta} & 0 \\
                    \sin{\theta} & \cos{\theta} & 0 \\
                    0 & 0 & -1
                \end{bmatrix}
            \end{align}
            % paragraph Rotations and Reflections in 3 Dimensions (end)

        \section{Linear Dependence and Independence}

            In general, a set of vectors is linearly dependent, if  there are some linear combinations of them equals \textit{zero} (other than all the coefficients 
            equal zero). We can use row reduction to form linear combinations by \textit{elementary row operations}. These operations are reversible, se we can combine
            the remaining vectors to form the original vectors. Thus the remaining vectors are \textit{independent} and called as \textit{basis vectors}.

            \paragraph{Linear Independence of Functions} % (fold)
            \label{par:Linear Independence of Functions}
            Similar to vectors, the functiond $f_1(x),\, f_2(x),\, \dots , f_n(x)$ are linearly dependent if there are some linear combinations of them is identically
            \textit{zero}, meaning, if there are constants $k_1, k_2, \dots , k_n$ (not all zero), such that,
            \begin{align*}
                k_1 f_1(x) + k_2 f_2(x) + \dots +k_n f_n(x) \equiv 0
            \end{align*}

            For a given set of functiond we have the fillowing theorem
            \coloredeq{eq:set of functions dependence}{
                \begin{aligned}
                    \text{if $f_1(x),\, f_2(x),\, \dots , f_n(x)$ have derivatives of order $n-1$ and if the determinant} & \\
                    W = \begin{vmatrix}
                        f_1(x) & f_2(x) & \dots & f_n(x) \\
                        f'_1(x) & f'_2(x) & \dots & f'_n(x) \\
                        f''_1(x) & f''_2(x) & \dots & f''_n(x) \\
                        \vdots & \vdots & \ddots & \vdots \\
                        f_1^{n-1}(x) & f_2^{n-1}(x) & \dots & f_n^{n-1}(x) \\
                    \end{vmatrix} \neq 0 & \\
                    \parbox{0.85\linewidth}{then the functions are linearly independent. the determinant of $W$ is called \textit{Wronskian} of the functions.
                    If $W \equiv 0$, it is not necessarily imply "functions dependent".}
                \end{aligned}
            }
            % paragraph Linear Independence of Function (end)

            \paragraph{Homogeneous Equations} % (fold)
            \label{par:Homogeneous Equations}
            We consider a set of equations, but a special case raises when the all the right side of those equations are zero.

            \coloredeq{homogeneous equations:1}{\parbox[t][3\baselineskip][c]{0.9\linewidth}{
                Homogeneous equations are never \textit{inconsistent}; they always have the solution "all unkowns $=0$" (called \textit{trivial solution}). If the number of 
                the independent \textit{equations} (the rank of the matrix) is the same as the the number of the \textit{unknowns}, this is the only solution. If the 
                \textit{rank} is less than the \textit{unknowns}, there are infinitely many solutions.
            }}

            \bulletpar For $n$ homogeneous equations in $n$ unknowns, they have only a trivial solution, unless their rank is less than $n$. Meaning that, at least one row
            of the reduced $n$ by $n$ coefficients matrix is zero row. Then its determinant is zero. Thus we say
            \coloredeq{homogeneous equations:2}{\parbox[t][2\baselineskip][c]{0.9\linewidth}{
                A system of $n$ homogeneous equations in $n$ unknowns has solutions other than the trivial solution iff the determinant of the coefficients is zero
            }}
            % paragraph Homogeneous Equations (end)

        \section{Special Matrices and Formulas}
            There are some importants special matrices. Consider a matrix $A$, so

            \begin{center}
                \begin{tabular}{m{11em} m{10em} m{10em}}
                    \hline
                    Name of matrix & Notations & Procedure \\
                    \hline\hline
                    Transponse of $A$. & $A^T$ or \~A or $A'$ & Interchange rows and columns in $A$. \\

                    Complex conjugate of $A$. & $\bar{A}$ or $A^*$ & The complex conjugate of each element in $A$. \\

                    Transponse conjugate,\newline Hermitian conjugate,\newline adjoint,\newline Hermitian adjoint.\newline & 
                    $A^\dagger$ & The complex conjugate of each element in $A^T$.\\

                    Inverse of $A$ & $A^{-1}$ & See Par.\eqref{par:Inverse of a Matrix}\\

                \end{tabular}
                \label{tab:special matrices}
            \end{center}
            Also the following are special properties, 
            \begin{center}
                \begin{tabular}{l l}
                    \hline
                    A matrix is & if it satisfies\\
                    \hline \\
                    real           & $A = \bar{A}$ \\
                    symmetric      & $A = A^T$, $A$ real \\
                    skew-symmetric or antisymmetric & $A = -A^T$, $A$ real \\
                    orthogonal     & $A^{-1} = A^T$, $A$ real \\
                    pure imaginary & $A = - \bar{A}$ \\ 
                    Hermitian      & $A = A^\dagger$ \\
                    anti-Hermitian & $A = - A^\dagger$ \\
                    unitary        & $A^{-1} = A^\dagger$ \\
                    normal         & $A A^\dagger = A^\dagger A$\\
                    \hline
                \end{tabular}
                \label{tab:matrix special properties}
            \end{center}

            \paragraph{Kronecker $\sigma$} is defined  as 
            \coloredeq{eq:Kronecker}{
                \sigma_{ij} = \begin{cases}
                    1, &\text{ if }i=j\\
                    0, &\text{ if }i \neq j
                \end{cases}
            }
            We can define the unit matrix as one whose elements $I = \sigma_{ij}$. We use Eq.\eqref{eq:matrix multiplication in indeces} to prove
            matrix associative law for matrix, that is 
            $$ A(BC) = (AB)C = ABC $$
            First we write, $(BC)_{kj} = \sum_l B_{kl} C_{li}$ then we have
            \begin{align*}
                [A(BC)]_{ij} &= \sum_k A_{ik} (BC)_{kj} = \sum_k A_{ik}\, \sum_l B_{kl} C_{lj} \\
                             &= \sum_k \sum_l A_{ik} B_{kl} C_{lj} = (ABC)_{ij}
            \end{align*}
            which is the index notation of $A(BC)= ABC$.

            Silimilarly, we can fins the transpose of the product ot two matrices. Recall that $A_{ik}^T = A_{ki}$.
            \begin{align*}
                (AB)_{ik}^T &= (AB)_{ki} = \sum_j A_{kj} B_{ji} = \sum_j A_{jk}^T B_{ij}^T \\
                            &=\sum_j B_{ij}^T A_{jk}^T = (B^T A^T)_{ik} \\
                    (AB)^T &= B^T A^T
            \end{align*}
            In general, 
            \coloredeq{eq:associative trnsponse}{
                \begin{aligned}
                    \parbox[t]{0.6\linewidth}{The transpose of a product of matrices is equal to the product of the transposes in \underline{reverse order}.}
                \end{aligned} &&
                (ABCD)^T = D^T D^T B^T A^T
            }
            A similar theorem foe the inverse
            \coloredeq{eq:associative inverse}{
                \begin{aligned}
                    \parbox[t]{0.5\linewidth}{The inverse of a product of matrices is equal to the product of the inverses in \underline{reverse order}.}
                \end{aligned} &&
                (ABCD)^{-1} = D^{-1} D^{-1} B^{-1} A^{-1}
            }

            \paragraph{Trace of a Matrix} % (fold)
            \label{par:Trace of a Matrix}
            The \textbf{trace} (or \textbf{spur}) of a matrix $A$ is the sum of the elements in the main diagonal.\\
            \fbox{The trace of a product of matrices is not changed by arrange them in cyclic order}. Meaning that,
            \begin{align*}
                \tr{ABC} = \tr{BCA} = \tr{CAB}
            \end{align*}
            proof,
            \begin{align*}
                \tr{ABC} = \sum_i (ABC)_{ii} &= \sum_i \sum_j \sum_k A_{ij} B_{jk} C_{ki} \\
                                            &= \sum_i \sum_j \sum_k B_{jk} C_{ki} A_{ij} = \tr{BCA} \\
                                            &= \sum_i \sum_j \sum_k C_{ki} A_{ij} B_{jk} = \tr{CAB}
            \end{align*}
            \textit{Heads up}: In general $\tr{ABC}$ is not equal to $\tr{ACB}$. To memorize it, take the last matrix an insetrt it before the first one, and on.
            % paragraph Trace of a Matrix (end)
        \section{Linear Vector Spaces}
            We dealt with vectors such that $\vec{r} = x\ihat + y\jhat + z\khat$, that represents a vector from the origin to the point $(x, y, z)$. The all possible 
            sets of vectors make up a 3-dimensional space $R_3$ ($R$ for real) or $V_3$ ($V$ for vector) or $E_3$ ($E$ for Euclidean). Generally, an ordered set of $n$
            points or vectors make up a $n$-dimensional space $V_n$. All the termonolgy are also aplies for $n$-D, such the distance and the orthogonality.

            \paragraph{Subspace, Span, Basis, Dimension} % (fold)
            \label{par:Subspace, Span, Basis, Dimension}

            \bulletpar Two vectors in 3-dimensional space represent a plane, or a vector space $V_2$. All thier possible combinations lie on that plane. Thye are also a 
            a part of $V_3$, we say that $V_2$ is a \textit{subspace} of $V_3$.

            \bulletpar A set of vectors \textbf{spans} a space if all the vectors in the space be written as linear combinations of the spanning set. A \textit{basis} vector is 
            a set of \textit{linearly independent} vectors which span a vector space.

            \bulletpar A \textit{dimension} of a vector space is equal to the number of the basis vectors. \\
            For example, the three unit vectors $\ihat, \jhat, \khat$ span the vector space $V_3$; they are linearly independent, thus a basis for the space $V_3$.
            % paragraph Subspace, Span, Basis, Dimension (end)

            \paragraph{Inner Product, Norm, Orthogonality} % (fold)
            \label{par:Inner Product, Norm, Orthogonality}
            Let us generalize the the dot product in Eq\eqref{eq:dot product} to $n$ dimensions
            \coloredeq{eq:dot product to nD}{\vec{A} \cdot \vec{B} = \sum_{i=1}^n A_i B_i}

            Similarly, to generalize the \textit{norm} of $\vec{A}$
            \coloredeq{eq:norm to nD}{A = |\vec{A}| = \sqrt{\vec{A} \cdot \vec{A}} = \sqrt{\sum_{i=1}^n A^2_i}}
            For orthogonality we have 
            \coloredeq{eq:orthogonality to nD}{\text{$\vec{A}$ and $\vec{B}$ are orthogonal if}\qquad  \sum_{i=0}^n A_i B_i=0}
            % paragraph Inner Product, Norm, Orthogonality (end)

            \paragraph{Schwarz Inequality} % (fold)
            \label{par:Schwarz Inequality}
            We can also use the formula $\vec{A} \cdot \vec{B} = AB \cos{\theta}$ to find the angle between two vectors in $n$-dimensions. The result must satisfy
            $|\cos{\theta}| \leq 1$, thus 
            \coloredeq{eq:Schwarz inequality}{|\vec{A} \cdot \vec{B}| \leq AB, \qquad or \quad 
            \left|\sum_{i=1}^n A_i B_i \right| \leq \sqrt{\sum_{i=1}^n A^2_i} \sqrt{\sum_{i=1}^n B^2_i}}
            \textit{Proof}: if $\vec{B} = \vec{0}$, Eq.\eqref{eq:Schwarz inequality} is stisfied. If $\vec{B} \neq \vec{0}$, consider the vector 
            $\vec{C} = B\vec{A} - (\vec{A} \cdot \vec{B})\vec{B}/B$. Then, $\vec{C} \cdot \vec{C} = \sum C_i^2 \geq 0 $
            \begin{align*}
                \vec{C} \cdot \vec{C} &= B^2 (\vec{A} \cdot \vec{A}) - 2B(\vec{A} \cdot \vec{B})(\vec{A} \cdot \vec{B})/B + (\vec{A} \cdot \vec{B})^2 
                (\vec{B} \cdot \vec{B})/B^2 \\
                &= A^2 B^2 -2(\vec{A} \cdot \vec{B})^2 + (\vec{A} \cdot \vec{B})^2 \\
                &= A^2 B^2 - (\vec{A} \cdot \vec{B})^2 = C^2 \geq 0
            \end{align*}
            Thus we can define the angle between two vectors as $\cos{\theta} = \vec{A} \cdot \vec{B} / AB$.
            % paragraph Schwarz Inequality (end)

            \paragraph{Orthonormal Basis; Gram-Schmidt Method} % (fold)
            \label{par:Orthonormal Basis; Gram-Schmidt Method}
            We call a set of vectors \textbf{orthonormal} if they are all mutually \textit{orthogonal} (perpendicular), and each vector is \textit{normal}ized;
            $\ihat$, $\jhat$, and $\khat$ as an example.

            \bulletpar Consider we have three basis vectors $\vec{A}$, $\vec{B}$, and $\vec{C}$. First, normalize $\vec{A}$. Substract 
            from $\vec{B}$ its projection along $\vec{A}$, then normalize it. And so on.
            % paragraph Orthonormal Basis; Gram-Schmidt Method (end)

            \paragraph{Complex Euclidean Space} % (fold)
            \label{par:Complex Euclidean Space}
            It is allowed to have a vector with components of comlex numbers. In Eq.\eqref{eq:dot product to nD} we need to make what under the square root positive. So 
            we replace $A^2_i$ by $|A_i|^2 = A^*_i A_i$. likewise, in Eq.\eqref{eq:norm to nD}, Eq.\eqref{eq:orthogonality to nD}, and Eq.\eqref{eq:Schwarz inequality}.
            \begin{align*}
                \vec{A} \cdot \vec{B} = \sum_{i=1}^n A^*_i B_i \\
                |\vec{A}| = \sqrt{\sum_{i=1}^n A^*_i A_i}
            \end{align*}
            \bulletpar In matrix form, suppose that $A$ is a column matrix with elemnts $A_i$. Then its transpose conjugate $A^\dagger$ is a row matrix with elements 
            $A^*_i$. We can write, $\sum A^*_i B_i = A^\dagger B$.
            % paragraph Complex Euclidean Space (end)

        \section{Eigenvalues and Eigenvectors; Diagonalizing Matrices}
            In some special matrix transformation, there are some specific vectors that do not move off of their span; such vectors are called 
            \textbf{eigenvectors}. They satisfiy $\vec{R} = \lambda \vec{r}$, where $\lambda$ is called the \textbf{eigenvalue}.

            \bulletpar \textbf{Eigenvalues} Consider the transformation of $\vec{R} = M\vec{r}$,
            \begin{align}
                \label{eq:eigenvalue:1}
                \begin{bmatrix}
                    X \\ Y
                \end{bmatrix} = 
                \begin{bmatrix}
                    a & b \\ c & d
                \end{bmatrix}
                \begin{bmatrix}
                    x \\ y
                \end{bmatrix}
            \end{align}
            The eigenvector condition $\vec{R} = \lambda \vec{r}$
            \begin{align}
                \label{eq:eigenvalue:2}
                \begin{bmatrix}
                    X \\ Y
                \end{bmatrix} = 
                \begin{bmatrix}
                    a & b \\ c & d
                \end{bmatrix}
                \begin{bmatrix}
                    x \\ y
                \end{bmatrix} = 
                \lambda \begin{bmatrix}
                    x \\ y
                \end{bmatrix}
            \end{align}
            Thus we get
            \begin{align}
                \label{eq:eigenvalue:3}
                \begin{bmatrix}
                    a - \lambda & b \\ c & d - \lambda
                \end{bmatrix} = 0 \qquad or
                \begin{cases}
                    (a- \lambda)x + b y = 0 \\
                    c x + (d- \lambda)y = 0
                \end{cases}
            \end{align}
            These are homogeneous equations. Thus from \eqref{homogeneous equations:1}, a set of homogeneous equations has a solution other than $x=y=0$, iff the 
            determinant of coefficients is zero.

            \coloredeq{finding the characteristic equation}{
                \text{For the maatrix } M = \begin{bmatrix}
                    a & b \\ c & d
                \end{bmatrix} &&
                \text{the characteristic equation is} \begin{vmatrix}
                    a - \lambda & b \\ c & d - \lambda
                \end{vmatrix} = 0
            }
            \bulletpar \textbf{Eigenvectors} By finding the the eigenvalue, $\lambda$, and substituting it back into Eq.\eqref{eq:eigenvalue:3}, will result line
            equations. Any vector from the origin to a point along those lines is considered as an eigenvector, where they stay in thier span.

            \paragraph{Diagonalizing a Matrix} % (fold)
            \label{par:Diagonalizing a Matrix}
            Again substitute the eigenvalues into Eq.\eqref{eq:eigenvalue:3}, and distinguish the eigenvectors by subscripts, 
            \begin{align}
                \label{eq:diagonalizing a matrix:1}
                \begin{aligned}
                    a x_1 + b y_1 = \lambda_1 x_1 && a x_2 + b y_2 = \lambda_2 x_2 \\
                    c x_1 + d y_1 = \lambda_1 x_1 && c x_2 + d y_2 = \lambda_2 x_2
                \end{aligned}
            \end{align}
            then by writing these equation as one matrix equation, \fbox{$MC = CD$}
            \begin{align}
                \label{eq:diagonalizing a matrix:2}
                \begin{bmatrix}
                    a & b \\ c & d
                \end{bmatrix} \begin{bmatrix}
                    x_1 & x_2 \\
                    y_1 & y_2
                \end{bmatrix} = \begin{bmatrix}
                    x_1 & x_2 \\
                    y_1 & y_2
                \end{bmatrix} \begin{bmatrix}
                    \lambda_1 & 0  \\ 0 & \lambda_2
                \end{bmatrix}
            \end{align}
            where $(x_1, y_1)$ is a point along the found line obtained from $\lambda_1$; which is the eigenvector of the transformation. If $\det{C} \neq 0$, then $C$ 
            has an inverse $C^{-1}$, so 
            \coloredeq{eq:diagonal matrix}{
                \parbox[t]{0.65\linewidth}{The matrix $D$ has nonzero numbers only in its main diagonal, so it is called \textbf{diagonal matrix}.
                Also, the matrix $D$ is called \textbf{similar} to $M$. When we obtain $D$ given $M$, we \textit{diagonalized} $M$ by a 
                \textit{similarity tranformation}} &&
                C^{-1}MC = D
            }
            \bulletpar The diagonalization process simplify problems by a better choice of variables. For example, it is simpler to to describe some of deformations by 
            using axes along the eigenvectors. Thus geometrically, 
            \bulletpar A complete example, consider two coordinate systems, $(x, y)$ and $(x', y')$, and two vectors that are beign transform from $\vec{r}=(x, y)$ into
            $\vec{R}=(X, Y)$. The $\vec{r}$ and $\vec{r'}$ in the two systems are related to each other by Eq.\eqref{eq:axes rotation matrix}, to find $(z, y)$,
            \begin{align}
                \begin{aligned}
                    \label{eq:diagonalization example:1}
                    x = x' \cos{\theta} - y' \sin{\theta} \\
                    y = x' \sin{\theta} + y' \cos{\theta}
                \end{aligned} \qquad \Leftrightarrow \quad 
                \begin{aligned}
                    r = C r', && C = \begin{bmatrix}
                        \cos{\theta} & -\sin{\theta} \\ \sin{\theta} & \cos{\theta}
                    \end{bmatrix}
                \end{aligned}
            \end{align}
            Similarly we have, $R = C R'$. Consider a transformation matrix $M$ in $(x, y)$ system, as
            \begin{align}
                \label{eq:diagonalization example:2}
                R = M r
            \end{align}
            Now, we can get the transformation equation in $(x', y')$ system by substituting in Eq.\eqref{eq:diagonalization example:2}, so
            \begin{align}
                \label{eq:diagonalization example:3}
                R' = C^{-1} M C r' \Leftrightarrow R' = D r'
            \end{align}
            Thus we conclude that
            \coloredeq{diagonal matrix meainig}{
                \parbox[t]{0.8\linewidth}{$D = C^{-1}MC$ is the transformation matrix described in the $(x', y')$ system, where $M$ describes the tranformation in the
                $(x, y)$ system.}
            }
            \starpar If $C$ is chosen to make $D$ a diagonal matrix, then the new axes $(x', y')$ are along the eigenvectors of $M$.
            If the eigenvectors are perpendicular, then the new axess $(x', y')$ are a set of perpendicular axes rotated by $\theta$ from 
            axes $(x, y)$.

            \bulletpar However, if $C$ is not an orthogonal matrix, then the axes $(x', y')$ are not perpendicular and $|\vec{r}| \neq |\vec{r'}|$.

            \bulletpar Thus, if the unit eigenvectors in matrix $C$; are perpendicular, then $C$ is \textit{orthogonal}. This is only if the matrix $M$ is 
            \textit{symmetric}.
            % paragraph Diagonalizing a Matrix (end)

            \paragraph{Degeneracy} % (fold)
            \label{par:Degeneracy}
            means two (or more) independent eigenvectors correspond to the same eigenvalue.\\
            For a \textit{symmetric} matrix, we saw that the eigenvectors corresponding to different eigenvalues are \textit{orthogonal}.
            % paragraph Degeneracy (end)

            \paragraph{Diagonalizing Hermitian Matrices} % (fold)
            \label{par:Diagonalizing Hermitian Matrices}
            We saw how to diagonalize symmetric matrices by \textit{orthogonal similarity transformation}. Consider thid analogy
            \begin{center}
                \begin{tabular}{c | c}
                    \hline Real & Complex \\ \hline \\
                    Symmetric $S^T = S$     & Hermitian $H^\dagger = H$ \\
                    Orthogonal $O^T = O^{-1}$ & Unitary $U^\dagger = U^{-1}$ \\
                    \hline
                \end{tabular}
            \end{center}
            \starpar The eigenvalues of a Hermitian matrix are always \textit{real}. \\
            Consider $H$ to be a Hermitian matrix, and $\vec{r}$ the column matrix of a non-zero
            eigenvector of $H$ corresponding to the eigenvalue $\lambda$. Thus the eigenvector condition is 
            \begin{align*}
                Hr = \lambda r
            \end{align*}
            Taking the transpose conjugate ($\dagger$) of this this equation, we get
            \begin{align*}
                (Hr)^\dagger = r^\dagger H^\dagger = r^\dagger H \qquad
                \text{and} \quad \lambda r \Rightarrow \lambda^* r^\dagger
            \end{align*}
            For a Hermitian matrix we have $H^\dagger = H$, $\lambda$ is just a number. Hence, 
            \begin{align*}
                Hr = \lambda r \qquad \text{and} \qquad r^\dagger H = \lambda^* r^\dagger
            \end{align*}
            Then multiply the left by the row matrix $r^\dagger$, and the right by the column matrix $r$,
            \begin{align*}
                r^\dagger Hr = \lambda r^\dagger r \qquad \text{and} \qquad r^\dagger H r = \lambda^* r^\dagger r
            \end{align*}
            Substracting the two equations we get  $(\lambda - \lambda^*) r^\dagger r = 0$. But we assumed $r \neq 0$, so $\Lambda^* = \lambda$, hence the eigenvalue 
            $\lambda$ is always \textit{real}.

            \bulletpar For a Hermitian matrix, the eigenvectors corresponding to two different eigenvalues are \textit{orthogonal}. Consider  this,
            \begin{align*}
                H r_1 = \lambda_1 r_1 \qquad H r_2 = \lambda_2 r_2
            \end{align*}
            \begin{align*}
                r_1^\dagger H r_2 = \lambda_1 r_1^\dagger r_2 = \lambda_2 r_1^\dagger r_2 \qquad \text{or} \quad (\lambda_1 - \lambda_2) r_1^\dagger r_2 = 0
            \end{align*}
            If $\lambda_1 \neq \lambda_2$, then $r_1 \cdot r_2 = 0$, hence, they are \textit{orthogonal}.

            \starpar If a matrix $M$ has a real eigenvalues and can be diagonalize by a \textbf{unitary similarity transformation}, then $M$ is \textit{Hermitian}.
            We state \fbox{MU = UD}
            \begin{align*}
                U^{-1} M U &= D \\
                (U^{-1} M U)^\dagger &= U^{-1} M^\dagger U = D^\dagger = D \\
                \text{thus}\quad U^{-1} M U &= U^{-1} M^\dagger U
            \end{align*}
            We get, $M = M^\dagger$, hence, $M$ is Hermitian.

            \coloredeq{Hermitian theorem}{\parbox[t]{0.8\linewidth}{
                A matrix has real eigenvalues and can be diagonalize by a \textit{unitary similarity transformation} iff it is \textbf{Hermitian}.
            }}
            But a \underline{real} Hermitian matrix is a symmetric matrix, and \underline{real} unitary is a orthogonal matrix,

            \coloredeq{Symmetric theorem}{\parbox[t]{0.8\linewidth}{
                A matrix has real eigenvalues and can be diagonalize by a \textit{orthogonality similarity transformation} iff it is \textbf{symmetric}.
            }}
            The normal matrices include, symmetric, Hermitian, orthogonal, and unitary matrices, thus

            \coloredeq{Normal theorem}{\parbox[t]{0.8\linewidth}{
                A matrix has real eigenvalues and can be diagonalize by a \textit{unitary similarity transformation} iff it is \textbf{normal}.
            }}
            % paragraph Diagonalizing Hermitian Matrices (end)

            \paragraph{Powers and Functions of Matrices}
            For the matrix $M$ raised to the power $n$, we can use the diagonalization,
            \begin{align*}
                M^n = C D^n C^{-1} \qquad \text{where} \qquad D = C^{-1} M C
            \end{align*}
        \section{An Introduction to Groups}
            Consider $\pm 1, \pm i$. No matter what the products and powers of them we compute, we never get numbers other than these foure. This  property of a set 
            of elements with a law of combinations is called \textbf{closure}. We are interested in groups of matrices, that is, in matrix representaions os of groups.

            \paragraph{A Group} % (fold)
            \label{par:A Group}
            is a set ${A, B, C, \dots}$ of elements --can be numbers, matrices, or operations-- togather with a law of combinations of two elements,  having four 
            conditions 
            \begin{enumerate}
                \item \textit{Closure}: The combination of any two elements is an element of the group.
                \item \textit{Associative}: The of combination satisfies the associative law; $(AB)C = A(BC)$.
                \item \textit{Unit element}: There is a unit element $I$ with the property that $AI = IA = A$ for every element of the group.
                \item \textit{Inverse}: Every element of the group has an inverse in the group; like, for $A$, there is an element $B$ such that $AB = BA I$.
            \end{enumerate}

            \bulletpar Thus, the set $\pm 1, \pm i$ is valid under these conditions. The \textbf{order of a finite group} is the number of elements in the group. If the elements 
            of a group of order $n$ are of the form $A, A^2, A^3, \dots, A^n =1$, it is called a \textbf{cyclic group}.

            \bulletpar Hence $\pm 1, \pm i$ is a cyclic group of order 4. A \textbf{subgroup} is a subset which is itself a group. The whole group or the unit element, 
            are called \textbf{trivial subgroup}, any other subgroup is called \textbf{proper subgroup}.Here $\pm 1, \pm i$, has the proper subgroup $\pm 1$.
            % paragraph A Group (end)

            \bulletpar A \textit{group multiplication} refers to the law of combination for  the group.

            \bulletpar Two gourps are called \textbf{isomorphic} if their multiplication tables are identical except for the names we attach to the elements.

            \bulletpar If evrey two group elements \textit{commute}, the group is called \textbf{Abelian}.

            \bulletpar Two group elements $A$ and $B$ are called \textbf{conjugate} elements, if there is a a group element $C$ such that $C^{-1}AC = B$. By making $C$
            be successively one \textit{group element} after another, we can find all the \textit{group conjugate} of $A$. This set is called a \textbf{class}.

### 2022-08-04 latex/PHYS213/chapters/chapter1.tex
add LaTeX files

--- 2022-08-04 latex/summary.tex
@@
-\documentclass[a4paper]{report}
-
-\usepackage{needed}
-
-\title{PHYS213}
-\author{Alfaifi, Ammar}
-\date{\today}
-
-\begin{document}
-    \maketitle
 \chapter{The Quantum Theory of Light}
@@
         \label{fig:Bragg scattering}
+        \includegraphics[width=0.9\linewidth]{figures/x-ray spectrum.png}
+        \caption{The x-ray spectrum of a metal target.}
+        \label{fig:x-ray spectrum}
     \end{wrapfigure}
@@
     \coloredeq{eq:Bragg equation}{\Delta{L} = 2d\sin{\theta} = n\lambda}
+    {\tiny \begin{itemize}
+        \item $n$ is the order of the intensity maximum, 
+        \item $\lambda$ is the x-ray wavelength, 
+        \item $d$ is the spacing between planes, 
+        \item $\theta$ is the angle of the intensity maximum measured from plane $A$.
+    \end{itemize}}
+    If $\lambda$ is known $d$ of crystal can be found. If $d$ is also known, the x-ray intensity vs. wavelength may be determined
+    as in Fig.\eqref{fig:x-ray spectrum}. This continuous x-ray spectrum results from glancing or indirect scattering 
+    of electrons from metal atoms. In such collisions only part of the electron’s energy is converted to electromagnetic radiation.
+    This radiation is called \textbf{bremsstrahlung} (German for braking radiation), which refers to the 
+    radiation given off by any charged particle when it is \textit{decelerated}.
 
-\end{document}
+    \paragraph{$\bullet$} The $\lambda_{min}$ is independent of the target composition and only depends on the tube voltage, $V$.
+    This might be explained by assuming it to the case of a \textit{head-on electron–atom collision} in which all of the 
+    incident electron’s kinetic energy is converted to electromagnetic energy in the form of a single x-ray photon. For this case we have
+    \begin{align}
+        \label{eq:electron energy into x-ray}
+        eV = hf = \frac{hc}{\lambda_{min}} \qquad or \quad \lambda_{min} = \frac{hc}{eV}
+    \end{align}
+
+    \begin{wrapfigure}{r}{0.3\linewidth}
+        \includegraphics[width=0.9\linewidth]{figures/compton effect.png}
+        \caption{Scattered x-ray intensity versus wavelength of Compton scattering.}
+        \label{fig:Scattered x-ray}
+        \includegraphics[width=0.9\linewidth]{figures/particle-like collision.png}
+        \caption{The quantum view}
+        \label{fig:scattered quantum x-ray}
+    \end{wrapfigure}
+
+    \paragraph{The Compton Effect} % (fold)
+    \label{par:The Compton Effect}
+    In 1922, Arthur Holly Compton confirmed that x-ray photons behave like particles with momentum $hf/c$.
+
+    \paragraph{$\bullet$} classical theory predicted that incident radiation of frequency $f_0$ should accelerate an electron 
+    in the direction of propagation of the incident radiation, and it should cause \textit{forced oscillations} of the electron 
+    and reradiation at frequency $f'$, where $f' \le f$.\footnote{This decrease in frequency of the reradiated wave is caused 
+    by a double Doppler shift, first because the electron is receding from the incident radiation, and second because the electron
+    is a moving radiator as viewed from the fixed lab frame} The frequency or wavelength of the scattered radiation should depend 
+    on the length of time the electron was exposed to the incident radiation as well as on the intensity of the incident radiation.
+
+    \paragraph{$\bullet$} Compton's result was that the wavelength of the x-ray is independent of the intensity of radiation and
+    exposure duration, and depends only on the scattering angle!.
+
+    \paragraph{\color{c3}$\bigstar$} Quantum model easily explained the lower scattered frequency $f'$, where the incident photon gives 
+    some its energy $hf$ and momentum to the recoiling electron.
+
+    \paragraph{$\bullet$}  From Fig.\eqref{fig:Scattered x-ray} the shifted peak at $\lambda'$ is caused by the scattering of 
+    x-rays from nearly free electrons. Assuming the photon exhibits particle-like behavior and collides elastically with a free
+    electron initially at rest, Fig.\eqref{fig:scattered quantum x-ray}. Because electron recoils at high speed, we treat the
+    collision relativistically. Thus from the energy conservation, 
+    \begin{align}
+        \label{eq:Compton derivation:1}
+        E + m_ec^2 = E' + E_e
+    \end{align}
+    where $E$ is incident photon energy, $m_ec^2$ is the rest mass-energy of the electron, $E'$ is the energy of scattered
+    photon, $E_e$ is the total relativistic energy of the electron after the collision. Likewise, by the momentum conservation, 
+    \begin{align}
+        \label{eq:Compton derivation:2}
+        p  = p' \cos{\theta} + p_e \cos{\phi} &&
+        p' \sin{\theta} = p_e \sin{\phi}
+    \end{align}
+    where $p$ is the momentum of the incident photon, $p'$ is the momentum of the scattered photon, and $p_e$ 
+    is the recoiling electron momentum. We observed the angle of the scattered photons $\theta$, so we shall solve
+    Eq.\eqref{eq:Compton derivation:1} and Eq.\eqref{eq:Compton derivation:2} simultaneously to eliminate $\phi$,
+    \begin{align}
+        \label{eq:Compton derivation:3}
+        p_e^2 = p'^2 + p^2 - 2p p' \cos{\theta}
+    \end{align}
+    Paradoxically, we need to use the wave nature of light to explain the particle-like behavior 
+    of photons. Recall that the energy of a photon and the frequency of the associated light wave are related 
+    by $E=hf$. If we assume that a photon obeys the relativistic express $E^2 = p^2c^2 + m^2c c^4$, and the photon mass is zero, 
+    \begin{align}
+        \label{eq:Compton derivation:4}
+        p_{photon} = \frac{E}{c} = \frac{hf}{c} = \frac{h}{\lambda}
+    \end{align}
+    A paradoxical situation; a particle property, the photon momentum, is given in terms of a wave property, $\lambda$, of 
+    an associated light wave. If the relations $E = hf$ and $p = hf/c$ are substituted into 
+    Eq.\eqref{eq:Compton derivation:1} and Eq.\eqref{eq:Compton derivation:3}, these become respectively
+    \begin{align}
+        \label{eq:Compton derivation:5}
+        E_e = hf - hf' + m_e c^2, \qquad 
+        p_e^2 = \left(\frac{hf'}{c}\right)^2 + \left(\frac{hf}{c}\right)^2 - \frac{2 h^2 f f' }{c2} \cos{\theta}
+    \end{align}
+    Because the Compton measurements do not concern the total energy and momentum of the electron, 
+    we eliminate $E_e$ and $p_e$ by substituting Eq.\eqref{eq:Compton derivation:4} into the following expression for 
+    the electron’s relativistic energy,
+    \begin{align}
+        \label{eq:C}
+        E_e^2 = p_e^2c^2 + m_e^2 c^4
+    \end{align}
+    And we get the Compton's equation,
+    \coloredeq{eq:Compton equation}{\lambda - \lambda' = \frac{h}{m_e c}(1-\cos{\theta})}
+    {\tiny \begin{itemize}
+        \item $\lambda_C = \frac{h}{m_e c}$ is Compton wavelength = \SI{0.0243}{\AA}
+    \end{itemize}}
+    % paragraph The Compton Effect (end)
+
+    \paragraph{Particle-Wave Complementarity} % (fold)
+    \label{par:Particle-Wave Complementarity}
+    Compton effect offers evidence that when light \textit{interacts} with matter it behaves as if it were composed 
+    of particles with energy $hf$ and momentum $h/\lambda$.
+    % paragraph Particle-Wave Complementarity (end)

### 2022-08-04 latex/PHYS213/chapters/chapter2.tex
add LaTeX files

\chapter{The Particle Nature of Matter}
    \section{The Atomic Nature of Matter}
        French chemist Lavoisier and his wife, who established the conservation of matter in 
        many careful chemical experiments; \\Dalton, who perceived the atomicity of nature in the law of multiple 
        proportions of compounds; \\Avogadro, who in a most obscure and little-appreciated paper, postulated that 
        all pure gases at the same temperature and pressure have the same number of molecules per unit volume; \\
        and Maxwell, who showed with his molecular-kinetic theory of gases how macroscopic quantities, such as 
        pressure and temperature, could be derived from averages over distributions of molecular properties.\\
        Jean Perrin and the ubiquitous Albert Einstein, who carried on very important theoretical and experimental 
        work concerning Brownian motion, the zigzag movement of small suspended particles caused by molecular impacts.
        Their work produced additional confirmation of the atomic-molecular hypothesis and resulted in improved values of 
        Avogadro’s number as late as the early 1900s.

    \section{The Composition of Atoms}
        There were four major succecful works:
        \begin{itemize}
            \item The discovery of the law of electrolysis in 1833 by Michael Faraday.
            \item The identification of cathode rays as electrons and the measurement of thec harge-to-mass 
            ratio $(e/m_e$) of these particles by Joseph John ( J. J.) Thomson in 1897.
            \item The precise measurement of the electronic charge ($e$) by Robert Millikan in 1909.
            \item The establishment of the nuclear model of the atom by Ernest Rutherford and coworkers Hans Geiger and 
            Ernest Marsden in 1913.
        \end{itemize}

        \begin{wrapfigure}{r}{0.2\linewidth}
            \includegraphics[width=0.9\linewidth]{figures/Electrolysis of molten NaCl.png}
            \caption{Electrolysis of molten NaCl.}
            \label{fig:Electrolysis of molten NaCl}
        \end{wrapfigure}

        \paragraph{Faraday’s Experiment}
        of the electrolysis of molten common salt (NaCl). 
        Faraday found that if \SI{96500}{C} of charge (1 faraday) is passed through such a molten solution, 23.0 g of Na will deposit 
        on the cathode and 35.5 g of chlorine gas will bubble off the anode, see Fig.\eqref{fig:Electrolysis of molten NaCl}. 
        In this case, exactly 1 gram atomic weight 
        or mole of each element is released because both are monovalent. For divalent and trivalent elements, exactly 21 and 13 of 
        a mole, respectively, would be released.

        \paragraph{$\bullet$} Doubling the quantity of charge passed doubles the mass of neutral element freed. Thus 
        \textbf{Faraday’s law of electrolysis}, 
        \coloredeq{eq:Faraday’s law of electrolysis}{m= \frac{q M}{(\num{96500})(valence)}} 
        {\tiny \begin{itemize}
            \item $m$ the mass of the released substance in grams 
            \item $M$ the molar mass in grams
            \item $q$ the total charge passed in coulombs 
        \end{itemize}}

        \paragraph{$\bullet$} Faraday’s law of electrolysis is explained in terms of atoms picture, 
        see Fig.\eqref{fig:Electrolysis of molten NaCl}. Charges passes through the molten solution in form of ions, which they left
        an \textit{excess} or \textit{deficiency} of one or more electrons. By the electric field of the battery, these ions move to
        the anode (where they lose $e^-$) or the cathode (where they gain $e^-$) and are released as neutral elements.

        \starpar Although it was not clear in 1833, Faraday’s law of electrolysis confirmed three important ideas:
        {\small \begin{enumerate}
            \item Matter consists of molecules and that molecules consist of atoms.
            \item Charge is quantized, because only integral numbers of charges are transferred at the electrodes.
            \item The subatomic parts of atoms are positive and negative charges, although they remained unknown.
        \end{enumerate}}

        \paragraph{Joseph John (J. J.)} In 1897, Thomson discovered that the “rays” seen in low-pressure 
        gas discharges were actually caused by negative particles (electrons) ended a debate dating back nearly 30 years:
        Were cathode rays material particles or waves?

        \bulletpar In Thomson's $e/m_e$ experiment, electrons are accelerated from the cathode to the anode, collimated by slits in 
        the anodes, and then allowed to drift into a region of perpendicular electric and magnetic fields.
        The $\vec{E}$ and $\vec{B}$ first are applied to produce an undefflected beam. If the $\vec{B}$ field is then turned off, 
        the beam deflected on the phosphorescent screen. From the deffelctiona and measured values of $\vec{E}$ and $\vec{B}$ the 
        charge-to-mass ratio, $e/m_e$ can be determined.

        \begin{wrapfigure}{r}{0.3\linewidth}
            \includegraphics[width=0.9\linewidth]{figures/thomson experiment.png}
            \caption{Defflection of negitve particle}
            \label{fig:defflection path}
        \end{wrapfigure}

        \bulletpar In Thomson experiment, see Fig.\eqref{fig:defflection path}, consider when only $\vec{E}$ exists between the plates. 
        $v_x$ remains the constant. $y_x$ is
        constant except between the plates, where the $e^-$ undergoes a constant upward acceleration due to electric force 
        following a parabolic path.

        To solve for the deflection angle, $\lambda$, we must solve for $v_x$ and $v_y$. Because $v_y$ initially is zero, $e^-$ leaves
        the plates with $y$ component velocity given by 
        \begin{align}
            \label{eq:Thomson e/m:1}
            v_y = a_yt
        \end{align}
        We know that $a=F_E/m_e = Ee/m_e = Ve/m_e d$, and $t = \ell/v_x$, where $V$ is the applied voltage. Thus 
        \begin{align}
            \label{eq:Thomson e/m:2}
            v_y = \frac{V e \ell}{m_e d v_x}
        \end{align}
        From Fig.\eqref{fig:defflection path}, $\tan{\theta}= v_y/v_x$ and assuming $\tan{\theta}\approx\theta$, 
        \begin{align}
            \label{eq:Thomson e/m:3}
            \theta \approx \frac{V \ell}{d v_x^2} \left(\frac{e}{m_e}\right)
        \end{align}
        To find $v_x$, Thomson apply a $\vec{B}$ field with exixting $\vec{E}$ field to balance the defflection, then thier forces must
        be equal like
        \begin{align}
            \label{eq:Thomson e/m:4}
            F_E = F_B \Leftrightarrow qE = qv_xB \qquad or \quad v_x = \frac{E}{B} = \frac{V}{Bd}
        \end{align}
        Substituting Eq.\eqref{eq:Thomson e/m:4} into Eq.\eqref{eq:Thomson e/m:3}, 
        \begin{empheq}[box=\fbox]{align}
            \frac{e}{m_e} = \frac{V\theta}{B^2 \ell d}
        \end{empheq}

        \starpar Thomson’s original value was only about \SI{1e11}{C/kg} (accepted value \SI{1.758803e11}{C/kg}), prior experiments 
        on the electrolysis of hydrogen ions had given $q/m$ values for hydrogen of about \SI{e8}{C/kg}. Thus it was clear 
        that Thomson had discovered a particle with a mass about 1000 times smaller than the smallest atom.

        \starpar Thomson noted that the $e/m_e$ ratio was independent of the discharge gas and the cathode metal. Also, 
        the particles emitted when electrical discharges were passed through different gases were found to be the same as 
        those observed in the photoelectric effect. Based on these observations, Thomson concluded that these particles must 
        be a universal constituent of all matter.

        \paragraph{Milikan's Value of the Elementary Charge} % (fold)
        \label{par:Milikan's Value of the Elementary Charge}
        In 1897, Thomson had been unable to determine e or me separately. About two years later, he was close  to the value of $e$ (\SI{1.602e-19}{C}) 
        with values of \SI{2.3e-19}{C} for charges emitted from zinc illuminated by ultraviolet light and \SI{1.1e-19}{C} for charges produced 
        by ionizing x rays and radium emissions.

        \bulletpar He concluded that “$e$ is the same in magnitude as the charge carried by the 
        hydrogen atom in the electrolysis of solutions.” The technique used by Thomson and his students to measure $e$ is especially 
        interesting because it represents the first use of the \texttt{cloud chamber technique} in physics and also formed the starting point 
        for the famous Millikan oil-drop experiment.

        \bulletpar Charles Wilson, one of Thomson’s students, had discovered that ions act as 
        condensation centers for water droplets when damp air is cooled by a rapid expansion. Thomson used this idea to form charged 
        clouds by using the apparatus.

        $Q$ is the measured total charge of the cloud, $W$ is the measured weight of the cloud, and $v$ is the rate of fall or terminal 
        speed. Thomson assumed that the cloud was composed of spherical droplets having a constant mass (no evaporation) and that 
        the magnitude of the drag force $D$ on a single falling droplet was given by \textbf{Stokes’s law},
        \begin{align}
            \label{eq:Stokes's law}
            D = 6 \pi a \eta v
        \end{align}
        {\tiny \begin{itemize}
            \item $a$ the radius of the droplet
            \item $\eta$ the viscosity of the air
            \item $v$ the terminal speed of the droplet
        \end{itemize}}
        The following procedure was used to find $a$ and $w$, the weight of a single drop. Because $v$ is constant, the droplet is in 
        equilibrium under its weight, $w$, and the drag force, $D$. Hence, we require that $w = D$, or
        \begin{align}
            \label{eq:Stokes's law:1}
            w = D \quad \equiv \quad \frac{4}{3} \pi a^3  \rho g = 6 \pi a \eta v 
            \qquad so \quad 
            a = \sqrt{\frac{9 \eta v}{2 \rho g}}
        \end{align}
        where $\rho$ is the mass density of the droplet. We can find $w$, the weight. 
        Then the number of drops $n$ (or number of ions) is given by $W/w$ and the electronic charge $e$ is equal to $Q/n$, 
        assuming that each droplet carries only \textit{one} electronic charge. This method is inaccurate because the 
        theory applies only to a \textit{single} particle and the particles are all assumed to be \textit{identical} in order to compare 
        the theory to experiments performed on a cloud.

        \bulletpar By observing single droplets Millikan eliminated the problems of assuming all particles to be identical and of making uncertain 
        measurements on a cloud. Millikan’s basic idea was to measure the rate of fall of a single drop acted on by gravity and drag forces, 
        apply Stokes’s law to determine the drop radius and mass, then to measure its upward velocity in an opposing electric field, and 
        hence determine the total charge on an individual drop\footnote{This idea was first applied to charged clouds of water vapor by 
        H. A. Wilson in 1903. Millikan switched from water to oil to avoid problems of a changing droplet mass and radius caused by water 
        evaporation.}.

        \bulletpar Millikan’s experiment is to pass oil droplet charged by an atomizer into a small hole in the upper plate of 
        parallel-capacitor plates. These droplets can be seen against the darkness when they are illuminated. If an $\vec{E}$ field is 
        applied to the capacitor plates, the drop may move slowly upward. A single droplet with constant mass and radius may be followed 
        for hours, alternately \emph{rising} and \emph{falling}, by simply turning the $\vec{E}$ field on and off. 
        The charge is found after a long series of measurements of constant upward velocities one observes a discontinuous change or 
        jump to a different upward velocity (higher or lower). \emph{This discontinuous change is caused by the attraction of an ion to the 
        charged droplet and a consequent change in droplet charge}. Such changes become more frequent when a source of ionizing radiation 
        is placed between the plates.

        \bulletpar The quantitative analysis of the Millikan experiment starts with Newton’s second law applied to the oil drop, 
        $\sigma F_y = ma_y$. Because the drag force $D$ is large, a constant velocity of fall is quickly achieved, and all measurements 
        are made for the case $a_y =0$, or $\sigma F_y =0$. If we assume that the magnitude of the drag force is proportional to the speed 
        ($D = Cv$), and refer to Figure 4.9, we find
        \begin{align*}
            Cv - mg = 0 \qquad \vec{E}\text{ off} \\
            q_1E - mg - C v'_1 = 0 \qquad \vec{E}\text{ on}
        \end{align*}
        Eliminating $C$, 
        \begin{align}
            \label{eq:Millikan's experiment:1}
            q_1 = \frac{mg}{E} \left(\frac{v + v'_1}{v}\right)
        \end{align}
        When the droplet undergoes change in its \emph{upward velocity} from $v'_1$ to $v'_2$ (other remain constant),
        \begin{align}
            \label{eq:Millikan's experiment:2}
            q_2 = \frac{mg}{E} \left(\frac{v + v'_2}{v}\right)
        \end{align}
        Dividing Eq.\eqref{eq:Millikan's experiment:1} by Eq.\eqref{eq:Millikan's experiment:2}
        \begin{empheq}[box=\fbox]{align}
            \frac{q_1}{q_2} = \frac{v+ v'_1}{v + v'_2}
            \label{eq:Millikan's experiment:3}
        \end{empheq}

        \starpar In fact Eq.\eqref{eq:Millikan's experiment:2} is a proof of the quantization of charge, because if successive speed ratios 
        are ratios of \emph{whole} numbers, successive charges on the drop must be multiples of the same elementary charge.

        \bulletpar  To determine the actual value of $e$, we need to find the mass. We can use the droplet radius $a$, from Stokes’s law in
        Eq.\eqref{eq:Stokes's law:1}. Thus the mass is 
        \begin{align*}
            m = \rho \cdot \text{volume} = \rho \frac{4}{3} \pi a^3
        \end{align*}

        \bulletpar Stokes’s law, as Millikan was aware, is only approximately correct for tiny spheres moving through a gas. 
        The expression $D=6 \pi a \eta v$ holds quite accurately for a $\SI{0.1}{cm}$ radius sphere moving through a liquid or for any 
        case where the moving-object radius, $a$, is large compared with the mean free path, $L$, of the surrounding molecules. 
        (The mean free path is essentially the average distance between molecules.) In the Millikan experiment, however, $a$ is of the 
        same order of magnitude as the mean free path of air at STP. Consequently, Stokes’s law overestimates the drag force, because the 
        droplet actually moves for appreciable times through a frictionless “vacuum.” Millikan corrected Stokes’s law by using a drag force 
        whose magnitude is
        \begin{align}
            \label{eq:Corrected Stokes's law}
            D \frac{6 \pi a \eta v}{1 + \alpha (l/a)}
        \end{align}
        and found that $\alpha=\num{0.81}$ gave the most consistent values of $e$ for drops of different radii. Further corrections to Stokes’s 
        law were made by Perrin and Roux.
        % paragraph Milikan's Value of the Elementary Charge (end)

        \paragraph{Rutherford's Model of the Atom} % (fold)
        \label{par:Rutherford's Model of the Atom}
        Ernest Rutherford, and his students Ernest Marsden and Hans Geiger conducted a series of experiments 1909-1914. By noticing a beam of $\alpha$ particles broadened
        on a passing through a metal foil easily penetrated the thin film of metal. Rutherford's apparatus is about: a collimated beam of $\alpha$ particles emitted with 
        speed \SI{2e7}{m/s} hit a thin gold foil. Most of the $\alpha$'s passed straight through the foil, but some were \textit{scattered} at angle $\phi$. The number of 
        scattered $\alpha$'s at each angle per unit detector area and per unit time was measured by aid of a microscope.

        \bulletpar Rutherford's insight was that, the mass and kinetic energy of the $\alpha$'s are large, even for a head-on collision with a particle of mass of a hydrogen
        atom would only deflect the $\alpha$ particle with angle of $1^\circ$!.

        \bulletpar However the repulsion force on a $\alpha$ particle would much greater the positive charge is assumed to be concentrated at a single point.

        \bulletpar Rutherford assumed that the large-angle scattering is produced by a single nuclear collision and the repulsion dorce is given by Coulomb's law
        \begin{align}
            \label{eq:Coulomb between alpha and nucleus}
            F = k \frac{(2e)(Ze)}{r^2}
        \end{align}
        where $r$ is thr distance between $\alpha$ \& the nucleus, $+2e$ is the charge of $\alpha$, and $Ze$ is the nucleus charge. Rutherford found that the number of 
        $\alpha$ particles entering the detector per unit time, $\Delta{n}$ at an angle $\phi$,
        \begin{align}
            \label{eq:Rutherford number of alpha}
            \Delta{n} = \frac{k^2 Z^2 e^4 N n A}{4R^2 (\frac{1}{2}m_\alpha v_\alpha^2)^2 \sin{\phi/2}^2}
        \end{align}
        {\tiny \begin{itemize}
            \item $R$ distance of the detector and the foil
            \item $\phi$ scattering angle
            \item $N$ number of of nuclei per unit area of the foil
            \item $n$ total number of $\alpha$ particles
            \item $A$ area of the detector
        \end{itemize}}
        Geiger confirmed the dependence of the scattering angle, the foil thinkness, and the $\alpha$ speed.

        \bulletpar Rutherford relized that Eq.\eqref{eq:Coulomb between alpha and nucleus} hold iff the $\alpha$ prticle does not have enough energy to deform of penetrate
        the scattering nucleus, he looked for the threshold where energy $\alpha$ should only just penetrating the nuclear redius at closest approach, so
        \begin{align}
            \label{eq:Rutherford alpha approach}
            \frac{1}{2} m_\alpha v_\alpha^2 = k \frac{(2e)(Ze)}{d_{min}}
        \end{align}
        At high kinetic energy this equation begins to fail. Rutherford confirmed for failure for Eq.\eqref{eq:Rutherford number of alpha} and 
        Eq.\eqref{eq:Rutherford alpha approach} for heavy metals.

        \starpar As result Rutherford and his students had shown that all the mass and charge of $Ze$ were concentrated in a minute nucleus, and that the electron must 
        circle this nucleus.

        However there new three questions raised: 
        \begin{enumerate}
            \item If in the nucleus only the protons, weher are the other half of the nuclear mass?
            \item What is the force that hold protons in such small distance of \SI{10e-14}{m}?
            \item How do thr electrons move around the nucleus?
        \end{enumerate}
        Rutherford's answer for the first one is that there are additional groupings od neutral particles of electron-proton pair. For the second one, that protons are held
        by an intense electrical force. However, James Chadwick descover it is a very strong force as well as the neutron. The last question was done by Niels Bohr.
        % paragraph Rutherford's Model of the Atom (end)

        \paragraph{The Bohr Atom} % (fold)
        \label{par:The Bohr Atom}
        Recall that glowing solids and liquids (even gases at high densities) emit a continuous distribution of wavelengths. This exhinits  the blackbody radiation curve of
        intensity vs. wavelength.

        \bulletpar In contrast, the discrete line spectrum emitted by low-pressure gas subject to an electric discharge. When light from such setup is examined through
        spectroscope, it is found to consist of a few bright lines.

        \bulletpar Also these lines are a characteristic of a particular element. In 1859, Kirchhoff contributed to spectroscopy, wehre there were the foundation of 
        \textbf{absorption spectroscopy}. Where this explained Joseph Fraunhofer's dark D-lines in the solar spectrum.

        \bulletpar Not all the emission lines are exist in absorption lines. They depend on the temperature and of the absorbing vapor.

        \bulletpar Anders \AA ngstr\"om, meausred the four visible lines of hydrogen atom, while in 1885, Johann J. Balmer found a formula that predict the wavelengths
        of these visible lines: $H_\alpha$ red, $H_\beta$ green, $H_\gamma$ blue, and $H_\sigma$ violet. Balmer's formula is
        \begin{align}
            \label{eq:Balmer formula:1}
            \lambda(cm) = C_2 \left( \frac{n^2}{n^2 - 2^2} \right), \qquad n=3, 4, 5, \dots 
        \end{align}
        where $\lambda$ is the wavelength emitted in $cm$ and $C_2=$\SI{3645.6e-8}{cm}, called the \textbf{convergence limit} because it gaves the wavelength of the line 
        with largest $n$. Today Balmer's formulas is in one single equation 
        \coloredeq{eq:Balmer wavelength formula}{\frac{1}{\lambda} = R \left( \frac{1}{n_f^2} - \frac{1}{n_i^2} \right)}
        where $n_f$ and $n_i$ are integers, and $R$ Rydberg wavelength.
        \begin{table}
            \centering
            \begin{tabular}{c c}
                \hline
                Lyman Seris (uv) & $n_f=1$ \\
                Balmer Seris (vis-uv) & $n_f=2$ \\
                Paschen Seris (IR) & $n_f=3$ \\
                Brackett Seris (IR) & $n_f=4$ \\
                Pfund Seris (IR) & $n_f=5$ \\
                \hline
            \end{tabular}
            \caption{\label{tab:Special Series for the Hydrogen Atom} Special Series for the Hydrogen Atom}
        \end{table}

        \bulletpar Electron revoloving idea around the nucleus has led to a disaster. According Maxwell's theroy, accelerated charges revoloving with orbital frequency $f$
        should radiate light with frequency $f$. As an electron radiate, its oribit radius decreases and its frequency increases, falling into the nucleus. Bohr postulated
        that classical radiation theroy, which had been confirmed by Hertz's experiment, did not hold to atomic-sized systems. He used Plank and Einstein work, where he 
        postulated that: electrons in atoms are generally confined to certain stable, nonradiating energy levels and orbits known as \textbf{stationary states}. He also 
        used Einstein's consept of the photon.

        \paragraph{Assumptions of the Bohr Theory} For an atom of hydrogen:
        \begin{itemize}
            \item The electron moves in circular orbits around the protons unser the influence of the Coulomb force.
            \item Only certain orbits are stabel, where there electrons are not radiate, Hence the energy is fixed and classical mechanics is valid.
            \item Radiation is emitted by the atom when the electron 'jumps' from a higher energatic state to a lower one. This jump cannot be visualized or treated
            \item classically. \textbf{The frequency $f$ of the emitted photon is independent of the electron's orbital motion}. 
            \coloredeq{eq:emitted photon energy}{E_i - E_f = hf}
            \item The size of allowed orbits is determined by an additional quantum condition imposed on the electron's orbital angular momentum; an integer multiple of 
            $\hbar$. \coloredeq{eq:Bohr orbital momentum}{m_e v r = n \hbar, \qquad n=1, 2, 3, \dots}
        \end{itemize}

        Now the total energy of the electeon-proton system is 
        \begin{align}
            \label{eq:Bohr energy:1}
            E = K + U = \frac{1}{2} m_e v^2 - k \frac{e^2}{r}
        \end{align}
        Applying Newton’s second law $F_r = mv^2/r$,
        \begin{align*}
            \frac{k e^2}{r^2} = \frac{m_e v^2}{r}
        \end{align*}
        Thus the kinetic energy 
        \begin{align}
            \label{eq:Bohr energy:2}
            K = \frac{m_e v^2}{2} = \frac{k e^2}{2r}
        \end{align}
        Substituting Eq.\eqref{eq:Bohr energy:2} in Eq.\eqref{eq:Bohr energy:1},
        \begin{align}
            \label{eq:Bohr energy:3}
            E = - \frac{k e^2}{2r}
        \end{align}
        The negative indicate that this is a \textit{bound} electron-proton system.
        We can find the orbits radius from Eq.\eqref{eq:Bohr orbital momentum} and Eq.\eqref{eq:Bohr energy:3},
        \coloredeq{eq:Bhor orbits radii}{r_n = \frac{\hbar^2}{m_e k e^2} n^2, \qquad n=1, 2, 3, \dots} The $r_1$ is called Bhor radius 
        \begin{align*}
            a_0 = \frac{\hbar^2}{m_e k e^2} = \SI{0.529}{\AA}
        \end{align*}

        The quantization of orbits resulted in the quantization of energy, substituting $r_n = n^2 a_0$ in Eq.\eqref{eq:Bohr energy:3},
        \coloredeq{eq:Orbits energy}{
            E_n = - \frac{k e^2}{2 a_0} \left( \frac{1}{n^2} \right)
        }
        $$ E_n = - \frac{\SI{13.6}{eV}}{n^2} $$

        $n$ corresponds to the discrete, or quantized vlues of the atom's energy, called \textbf{quantum numbers}.
        $n=1$ is called the \textbf{ground state}, and $n=2$ is called the \textbf{first excited state}.
        $n= \infty$ and $E = 0$, the electron is removed. The minimum required energy to remove an electron is called \textbf{ionization energy}.

        By Eq.\eqref{eq:Orbits energy} and Bohr's third postulate, we can find the wavelength of emitted photon when it jumps, 
        $$ f = \frac{E_i - E_f}{h} $$
        or \coloredeq{eq:Bohr-Balmer wavelength}{\frac{1}{\lambda} = \frac{f}{c} = \frac{k e^2}{2 a_0 h c} \left( \frac{1}{n_f^2} - \frac{1}{n_i^2} \right)}
        This \textit{theretical} result is identical to Balmer's \textit{empirical} formula in Eq.\eqref{eq:Balmer wavelength formula}.

        \bulletpar With this, Bohr explained the apectral series by the electrons trnsitions. He extended his model for hydrogen to other atoms; whose on electron is 
        removed (ionized). Tn general, a single electron orbiting a nucleus of charge $+Ze$, Bohr's theory gives, 
        \begin{align}
            \label{eq:Bohr model other atoms}
            r_n =  \frac{a_0}{Z} n^2, \qquad and \quad E_n = - \frac{ke^2}{2 a_0} \left( \frac{Z^2}{n^2} \right)
        \end{align}

        Beside all that Bohr explained more in his three-part papaer in 1913:
        \begin{itemize}
            \item He explained why fewer lines are seen unthe absorption hydrogen than  in the emission spectrum.
            \item He explained the emmision of x rays from atoms.
            \item He explained the nuclear origin of $\beta$ particles.
            \item He explained the chemical properties of atoms in terms of the electron shell model.
            \item He explained how atoms associate to form molecules.
        \end{itemize}
        \bulletpar To go deeper, Bohr explained the absorption as the reverse process of emission. Ordinarily, hydrogen are in the ground state ($n=1$), so only the Lyman series 
        will take plce. Balmer series is not seen because the average thermal energy of each atoms is not sufficient to make the electron in the first excited state. 
        That is, the number of electrons in $n=2$ is insufficient in low temperatures.

        \bulletpar Bohr, in his paper part II, attempted to find stabel electronic arrangemnets so that the total angular momentum of all the electrons is quantized and 
        the total energy is minimum. Although this is so difficult when more electrins are introdued in the system, Bohr successed in showing that the hydrogen atom can 
        take another electron to becone an ion $H^-$, and the helium was stable by having a closed inner shell with two electrons. He also, consider the lithium (Z=3), and 
        explained the tendency of lithium to loose the an electron.

        \starpar Bohr's predictions about the multielectron atoms:
        \begin{itemize}
            \item Electrons of elements with higher atomic number form stable concentric \textit{rings}, with definite numbers of electrons allowed for each ring.
            \item The number of electrons in the outter most ring determines the \textit{valency}.
        \end{itemize}
        % paragraph The Bohr Atom (end)

    \section{Bohr's Correspondence Principle; Quantization of Angular Momentum}
        Bohr's Correspondence Principle: \textbf{predictions of quantum theory must correspond to the predictions classical physics in the region of sizes where 
        classical theory is known to hold}.
        For the classical sizes of length, mass, and time, if the quantum number becomes large because if increase size of these quantities, we can say that
        $$ \lim_{n \to \infty}\, [quantum physics] = [classical physics]$$
        This principle becomes a useful tool to test a new quantum results, As Bohr himselfe did to arrive to the consept of the quantized orbital angular momentum.
        Bohr argued that his correspondence principle, the quantum condition for emission ($\Delta{E}=hf$) ans Maxwell's classical radiation theory (accelerated charges 
        with orbital frequency $f$ radiate light with the sane frequency) \textit{must simultaneously hold for the case of the extremely large electronic orbits}.

        \bulletpar Consider two large adjacent orbits $r_1$ and $r_2$,separated with energy $dE$, and orbital angular frequency $\omega$, where the last is approximately
        constant in transition between large orbits. To get know the allowed vslues of the angular momentum from the atom's change in energy, we need a connection 
        bewtween \textit{total energy} of the atom Eq.\eqref{eq:Bohr energy:3}, and the \textit{total momentum} of the atom ($L = m_e v r = m_e \omega r^2$). If the 
        electron is hild by Coulomb force, we can show that
        \begin{align}
            \label{eq:Bohr quantization of L:1}
            \frac{k e^2}{r^2} = \frac{m_e v^2}{r} \quad \Rightarrow\, \frac{1}{r} = \frac{m_e k e^2}{L^2}
            \quad \Rightarrow\, E = - \frac{m_e k^2 e^4}{2 L^2}
        \end{align}
        Taking the derivitive to get the connection bewteen the change in energy and the change in momentum,
        \begin{align}
            \label{eq:Bohr quantization of L:2}
            \frac{dE}{dL} = \frac{m_e k^2 e^4}{L^3}
        \end{align}
        Before let us find $L$ in terms of $\omega$ (electron orbital angular frequency), from Eq.\eqref{eq:Bohr quantization of L:1}
        \begin{align}
            \label{eq:Bohr quantization of L:3}
            L^2 = m_e k e^2 v r \quad \Rightarrow L^3 = (m_e v r) m_e k e^2 r \quad 
            \Rightarrow 
        \end{align}
        Again from Eq.\eqref{eq:Bohr quantization of L:1}
        \begin{align}
            \label{eq:Bohr quantization of L:4}
            ke^2 = m_e v^2 r \quad \Rightarrow\, m_e v r = \frac{k e^2}{v}
        \end{align}
        Substituting in Eq.\eqref{eq:Bohr quantization of L:2}
        \begin{align}
            \label{eq:Bohr quantization of L:5}
            \frac{dE}{dL} = \omega
        \end{align}
        Now consider the enission of a photon of energy $dE = \hbar \omega'$, trnsition from $r_1$ to $r_2$, 
        Eq.\eqref{eq:Bohr quantization of L:5} becomes
        \begin{align}
            \label{eq:Bohr quantization of L:6}
            dE = \omega\,  dL \qquad or \quad \hbar \omega' = \omega\, dL
        \end{align}
        where $\omega'$ is the photon angular frequency, and $\omega$ is the electron orbital frequency. Normally, they are not related, but according the correspondence
        principle for such large orbits, quantum theory must give the same frquency as the Maxwell's law of radiation. Thus, Maxwell's theory requires that 
        the orbital frequency equals to the light frequency, $\omega' = \omega$, hence 
        \begin{align}
            \label{eq:Bohr quantization of L:7}
            \hbar \omega = \omega\, dL \qquad or \quad dL = \hbar
        \end{align}
        Eq.\eqref{eq:Bohr quantization of L:7} shows that the change in the electronic angular momentum for transition between adjacent, large orbits, is \textit{always}
        $\hbar$. Thus, the magintude of total angular momentum of an electron in a orbit is an integer multiple of $\hbar$
        \begin{align}
            \label{eq:Bohr quantization of L:result}
            L = m_e v r = n \hbar
        \end{align}
        Bhor relized that although Eq.\eqref{eq:Bohr quantization of L:result} was derived for large orbits case, it was a universal quantum principle.

    \section{Direct Confirmation of Atomic Energy Levels}
        We saw before the idea of the quantized energy levels in atoms, by observing the the optical line apectra. Howeverm in 1914, a directed experiment was conducted by
        James Franck and Gustav Hertz.

        \begin{wrapfigure}{r}{0.25\linewidth}
            \includegraphics[width=0.9\linewidth]{figures/Franck-Herts apparatus.png}
            \label{fig:Franck-Herts apparatus}
            \caption{Franck-Herts apparatus}
        \end{wrapfigure}

        As in Fig.\eqref{fig:Franck-Herts apparatus}, Franck-Herts apparatus is about electrons emitted by the filament and accelerated throught long regoin 
        ($\approx 1 cm$)by a positive potential on the grid, $V$. The electrons can reach the cllector and be detected by elctrometer (a sensitive ammeter), if they 
        have enough energy to pass potential of 1.5V between the grid and the collector.

        \bulletpar At low electron energies or accelerating voltages, perfectly elastic collisions occur between the electrons and Hg atoms, in which the 
        \textit{sum} of the kinetic energies of electron-atom system is conserved. Since Hg atom is much more massive than the electron, the electron transfers 
        litle KE to the atom. Even with multiple collisions, the electron each the grid with KE of $e \cdot V$.

        \bulletpar As the accelerating voltage is increased, a threshold voltage where the inelastic collisions occur at the grid, with KE$=e \cdot$. In such collisions
        electrons can transfer almost all of thier KE to the atom, raising it to its first excited state. Thus these electrons are unable to overcome the retarding 
        voltage, hence the current $I$ is drop.

        \bulletpar For the second dip in the graph, where $V$ is increased enough so that an electron have two successive inelastic collisions.

        \bulletpar Franck and Hertz with ammeter and voltmeter measurements showed that atoms only accept discrete amounts of energ from electron beam.
        Also, they showed that the enrgy levels obtained from electron bombardment agreed with the spectroscopic results.

### 2022-08-04 latex/PHYS213/chapters/chapter3.tex
add LaTeX files

\chapter{Matter Wave} % (fold)
\label{cha:Matter Wave}
    \section{The Pilot Wave of De Broglie}
        By 1920s scientists reliazed that Bohr theory had some issues
        \begin{itemize}
            \item It failed to predict of the observed intensities of the spectral lines.
            \item It had limited success in predicting emission and apsorption wavelengths for multielectron atoms.
            \item It failed to provide an equation of motion governing the time development of atomic systems starting from some initial state.
            \item It overemphasized the particle nature of amtter and could not explain the newly dicovered wave-particle duality of light.
            \item It did not supply a general scheme for "quantizing" other systems, especially those without periodic motion.
        \end{itemize}

        In 1923, Louis Victor de Broglie postulated that \textit{because photons have wave and particle characteristics, perhaps all forms of matter have wave 
        as well as particle properties}. According to him, electrons had a dual wave-particle nature.

        He found the the wavelength of a \textit{matter wave}
        \coloredeq{eq:de Broglie wavelength}{\lambda = \frac{h}{p} \qquad f=\frac{E}{h}}
        {\tiny \begin{itemize}
            \item $h$ is Planck's constant, 
            \item $p$ is the relativistic momentum
            \item $E$ is the total relativistic energy
        \end{itemize}}

        \paragraph{De Broglie's Explanation of Quantization in Bohr Model} % (fold)
        \label{par:De Broglie's Explanation of Quantization in Bohr Model}
        De Broglie applied the ideas of familiar wave properties such as standing wave and interference. He postulated that the electron as an standing wave bent aorund
        the nucleus with radius of Bohr orbit $r$. Thus only disceret set of wavelengths is allowed for standing waves.
        Amazingly, Bohr theory for quantization of angular momentum can be esaily explained by de Broglie's hypothesis: \textbf{the allowed Bohr orbits arise because the 
        electron matter waves interfere constructively when an integral number of wavelengths exactly fits into the circumference of a circle orbit}. Thus
        \begin{align}
            \label{eq:de Broglie hypothesis}
            n \lambda = 2 \pi r
        \end{align}
        where $r$ is the radius of the orbit. From Eq.\eqref{eq:de Broglie wavelength}, we have $\lambda=h/m_e v$, substituting this into Eq.\eqref{eq:de Broglie hypothesis}
        and solving for $L$, $m_e v r$
        \coloredeq{eq:de Broglie quantization}{m_e v r = n \hbar}
        which is exactly Bohr condition for quantization of angular momentum.
        % paragraph De Broglie's Explanation of Quantization in Bohr Model (end)

        \paragraph{The Davisson-Germer Experiment} % (fold)
        \label{par:The Davisson-Germer Experiment}
        In 1924, de Broglie had sugested that a stream of electrons traversing a small aperture shouls undergo diffraction phenomena. In 1925, Einstein was led to the need
        of postulating matter waves from an analysis of fluctuations of molecular gas. Also, he noted that a molecular beam should show a small diffraction effects.
        In 1927, the proof of the wave nature eas obtained by the work of Davisson and Germer, and George P. Thomson.

        \bulletpar Davisson-Germen's experiment was to understand the arrangement of atoms on the surface of a nickle sample by elastically scattering a beam of low-speed
        electrons from a polycrystalline nickle traget. They had three parameters that can be changed-- \textit{electron energy}; \textit{nickle target orientation}, 
        $\alpha$; and \textit{scatterin angle}, $\phi$.

        \bulletpar Their experiment was going normally, with constant electron energy of \SI{100}{eV}, the scattered intensity rapidly decreased as $\phi$ increased.
        Accidentally, someone dropped a flask of liquid air on the glass vacvum system, tearing the vaccum and oxidizing the nickle target, that had been in high
        temperature. To remove the oxide, they reduce the sample by heating it carefully in a flowing stream of hydrogen. After this reassembling, they found very
        different results: \textit{Strong variations in the intensity of scattered electrons with angle}.

        \bulletpar This heating annealed the nickle target, causing large single-crystal regions to develop in the polycrystalline sample. These crystalline regions 
        provided the extended regualr lattice needed to observe electron diffraction. Davisson and Germen relized that it was the elastic scattering from \textit{single
        crystals} that produced such unusual results.

        \bulletpar The idea is that they calculated the eavelength of the electrons from simple diffraction formula and compared it with de Broglie's formula.
        Thus, for an electron we have 
        \begin{align*}
            \frac{1}{2}m_e v^2 = eV
        \end{align*}
        Substitute $v$ into de Broglie relation
        \begin{align*}
            \lambda = \frac{h}{m_e v} = \frac{h}{\sqrt{2Vem_e}}
        \end{align*}
        Also, the the wavelength csn be obtainde by considering the nickle atoms to be a reflaction diffraction grating. Only the atoms at the surface is considere, since
        x-rays cannot penetrate deeply into crystal. Thus to find the constructive interference we have 
        \begin{align}
            d \sin{\phi} = n \lambda
        \end{align}
        The diffraction lines from high-energy electrons are so sharp, and these electron can penetrate deeper in the crystal. As result, we need to use Bragg's law instead.

        \bulletpar Davisson and Germen concluded that this wavelength agrees very well with that from de Broglie's formula, confirming the electron matter wave. If de Broglie's 
        postulate is true for all matter, then any objects with mass $m$ has wavelike property.
        % paragraph The Davisson-Germer Experiment (end)

        \paragraph{The Electron Microdcope} % (fold)
        \label{par:The Electron Microdcope}
        The ability of controlling electron eavelength, even shorther than the visible light, provide us with a great tool investigate small objects.
        First \textbf{transmission electron microscope} (TEM), was built by Max Knoll and Ernst Ruska in 1931; it focuses electron beams accelerated by \SI{100}{kV}
        with magnetic lenses
        and creates a flta-looking two-dimensional shadow pattern on its screen, the resultof varying degrees of electron transmission through the object.

        \bulletpar magnificationAn optical microscope using ultraviolet has a maginification of 2000 and resolving power of \SI{100}{nm}. TEM can reach a magnification of \num{1000000}
        and resolving power of \SI{0.2}{nm}

        \bulletpar A second type of electron microscope with less resolution and magnification than TEM, but capable of producing three-dimensional images, is
        called \textbf{scanning electron microscope} (SEM).
        % paragraph The Electron Microdcope (end)

    \section{Wave Groups and Dispersion}
        The matter eave of a moving particle has a large probability to be found in a small region of space only at a specific time. Hence, the traveling 
        sinusoidal matter wave od finite extent and constant amplitude cannot represent a localized moving particle.

        \bulletpar Thus we need to introduce a pulse or a \textbf{wave group} of limited spatial extent. This pulse can be found by adding up sinusiodal wave with 
        different wavelengths. The resulting wave group will have a \textbf{group apeed} $v_g$ identical to the classical particle speed.

        \begin{wrapfigure}{r}{0.25\linewidth}
            \includegraphics[width=0.9\linewidth]{figures/group-wave.png}
            \label{fig:wave group}
            \caption{A wave group}
        \end{wrapfigure}

        \bulletpar In fact, all observed waves are limited to definite regions of space and are called \textit{pulses}, \textit{eave groups}, or \textit{eave 
        pachets} in matter waves. A plane wave with an exact wavelength and an infinite extension is an abtraction. Water waves, light waves, waves along a string,
        and sound wave all must be modeled by a group wave.

        \starpar A \textbf{wave group} consists of a superposition of waves with \textit{different wavelengths}, with the amplitude and phase of each component
        wave adjusted to make a constructive interference over a small region of space. A good example is the beats in sound waves.

        \bulletpar Consider a one=dimensional wave propagating in the positive $x$ direction with a \textbf{phase speed} $v_p$. This traveling wave with 
        wavelength $\lambda$ and frequenct $f$, and amplitude $A$ is written like 
        \begin{align}
            \label{eq:wave group:1}
            y = A \cos{\left( \frac{2 \pi}{\lambda} - 2 \pi ft \right)} \qquad \text{and} \quad v_p = \Lambda f
        \end{align}
        take $\omega = 2 \pi f$ (\textit{angular frequency}) and $k=2 \pi/\Lambda$ (\textit{wavenumber}), thus 
        \begin{align}
            \label{eq:wave group:2}
            y = A \cos{(k x - \omega t)}
        \end{align}
        and the the \textit{phase velocity}
        \coloredeq{eq:group velocity}{v_p = \frac{\omega}{k} }

        \bulletpar Let us form a superposition of two waves having the same amplitude and treavleing in the positive $x$, but different in wavelength, frequency, 
        and phase velocities. Hence,
        \begin{align}
            \label{eq:wave group:3}
            y = y_1 + y_2 = A \cos{(k_1 x - \omega_1 t)} + \cos{(k_2 x - \omega_2 t)}
        \end{align}
        Using the trigonometric identity: $\cos{a} + \cos{b} = 2\, \cos{\frac{1}{2}(a-b)} \cos{\frac{1}{2}(a+b)}$, we get 
        \begin{align}
            \label{eq:wave group:4}
            y = 2A \cos{\frac{1}{2}\left[ (k_2-k_1)x - (\omega_2-\omega_1)t \right]} \cos{\frac{1}{2} \left[(k_1+k_2)x - (\omega_1+\omega_2)t \right]}
        \end{align}
        For case of two waves with slightly different values of $k$ and $\omega$, we see that $\Delta{k} = k_2-k_1$ $\Delta{\omega} = \omega_2-\omega_1$ represent 
        small wave, but $(k_1+k_2)$ and $(\omega_1 + \omega_2)$ represent larger wave. So we can say that Eq.\eqref{eq:wave group:4} has a broad sinusoidal 
        \underline{envelope}, 
        \begin{align}
            \label{eq:wave group:5}
            2A \cos{\left( \frac{\Delta{k}}{2}x - \frac{\Delta{\omega}}{2}t \right)}
        \end{align}
        modulating a high-frequency wave within the envelope
        \begin{align}
            \label{eq:wave group:6}
            \cos{\left( \frac{k_1+k_2}{2}x - \frac{\omega_1+\omega_2}{2}t \right)}
        \end{align}
        This superposition is shown in Fig.\eqref{fig:wave group}. We can find the speed of the high-frequency wave and the envelope wave by dividing the 
        coefficient of $t$ by the one of $x$,
        \begin{align}
            \label{eq:wave group:7}
            v_p = \frac{(\omega_1+\omega_2)/2}{(k_1+k_2)/2} \approx \frac{\omega_1}{k_1} = v_1
        \end{align}
        and for the envelope velocity
        \begin{align}
            \label{eq:wave group:7}
            v_g = \frac{\Delta{\omega}}{\Delta{k}}
        \end{align}

        \bulletpar A general characteristic of wave groups is that both is a linited duration in time, $\Delta{t}$, and a limited extent in space, $\Delta{x}$.
        Thus the smaller the spatial width of the pulse, $\Delta{x}$, the larger the range od wavelengths or wavenumbers, $k$,
        \begin{align}
            \label{eq:wave group:8}
            \Delta{x} \Delta{k} \approx 1
        \end{align}
        Likwise, the smaller the temporal duration, $\Delta{t}$, the larger the frequencies, $\omega$
        \begin{align}
            \label{eq:wave group:9}
            \Delta{t} \Delta{\omega} \approx 1
        \end{align}
        In electronics, this condition is called \textit{response time-bandwidth formula}.

        \starpar \textbf{Both $\Delta{x}$ and $\Delta{k}$ cannot become arbitrarily small, but as one decreases the other must increase}. Our simple two-wave 
        model can show the general principle given by Eq.\eqref{eq:wave group:8} and Eq.\eqref{eq:wave group:9}. Let the spatial extent of our group be the 
        distance between two adjacent minima ($y=0$), labeled $\Delta{x}$ in Fig.\eqref{fig:wave group} we find from the envelope term: 
        $2A \cos{(\frac{1}{2}\Delta{kx})}$, then $\frac{1}{2} \Delta{k} \Delta{x}=\pi$
        \begin{align}
            \label{eq:wave group:10}
            \Delta{k} \Delta{x} = 2 \pi
        \end{align}
        where $\Delta{k} = k_2 - k_1$ is the present wavenumbers. Similarly, if we hold $x$ constant and changing $t$, we have 
        $\frac{1}{2}(\omega_2-\omega_1)\Delta{t} = \pi$, 
        \begin{align}
            \label{eq:wave group:11}
            \Delta{\omega} \Delta{t} = 2 \pi
        \end{align}
        as result we see that Eq.\eqref{eq:wave group:10} and Eq.\eqref{eq:wave group:11} agree with the \textit{general principles}.

        \bulletpar The addition of two waves with discrete frequencies is informative, but produces an infinite wave instead of a true pulse. Im general, many waves 
        having a continuous distribuation of wavelengths must be added to form a packet, finite over a limited range. In this case, the Eq.\eqref{eq:wave group:7},
        for \textbf{group velocity} becomes, 
        \coloredeq{eq:group velocity}{
            v_g = \frac{d\omega}{dk} \Big|_{k_0}
        }
        where $k_0$ is the centeral wavenumber of the many waves present. We can find a connection bewtween the group velocity and the phase velocity, if 
        $\omega = k v_p$ then,
        \begin{align}
            \label{eq:group and phase velovity}
            v_g = \frac{d\omega}{dk} \Big|_{k_0} = v_p \Big|_{k_o} + k \frac{dv_p}{dk} \Big|_{k_0}
        \end{align}

        \bulletpar \textbf{Dispersion} happens in materials in which the \textit{phase velocity} changes with wavelength.

% chapter Matter Wave (end)

### 2022-08-04 latex/PHYS213/mapping.tex
add LaTeX files

\documentclass[twocolumn]{article}

% \usepackage{../needed}
\usepackage{xcolor}
\usepackage{wrapfig}
\usepackage{empheq}
\usepackage{amsmath}
\usepackage[a4paper, margin=0.95in]{geometry}

\definecolor{c3}{RGB}{80, 90, 140}
\definecolor{c2}{RGB}{230, 240, 255}

\newcommand{\coloredeq}[2]{\begin{empheq}[box=\colorbox{c2}]{align}\label{#1}#2\end{empheq}}

\title{PHYS213 Paper}
\author{Alfaifi, Ammar}
\date{\today}

\begin{document}
    \maketitle
    \begin{abstract}
        Up to now, we only considered, in prevoius chapters, the quantum mechanics in one dimensional
        space. We need to extend the quantumn idea in more degree of freedom, i.e., for 3D space, as well
        as the centeral force for atoms. Further, we aim to distinguish the \emph{sharp} observables.
    \end{abstract}
    \section{\color{c3}\MakeUppercase{Particle in a Three-Dimensional Box}}
        \paragraph{} % (fold)
        \label{par:Particle in A Three-Dimensional Box}
        We consider a Cubic box whose length $L$. Inside this box a free particle collide
        with \emph{smooth} elastically with box' walls, presrving the magnitude of each
        momentum component, in addition to the total energy. This means that the particle occupies the 
        the region $0<x, y, z<L$.
        % paragraph Particle in A Three-Dimensional Box (end)

        \paragraph{}
        Since this is a three dimensional coordinate, the wavefunction $\Psi$ is a function of
        the position vector $\vec{r}=(x, y, z)$. The probability density
        $|\Psi(\vec{r})|^2$ is a \emph{probability per volume}. Thus, the \textbf{Schr\"odinger's
        equation} is given by \coloredeq{3D box equation}{
            -\frac{\hbar}{2m} \nabla^2\Psi + U(\vec{r}) = i\hbar \frac{\partial \Psi}{\partial t}
        }

    \section{\color{c3}\MakeUppercase{centeral forces and angular momentum}}
        \paragraph{}
        \begin{wrapfigure}{r}{0.5\linewidth}
            \includegraphics[width=0.9\linewidth]{figures/spherical-coordinates}
            \caption{Spherical coordinates for an electron.}
            \label{fig:spherical coordinates}
        \end{wrapfigure}

        In contrast of the Cartesain coordinates, we turn our attention to the spherical coordinates,
        since it is more convenient while dealing with \textbf{centeral forces}. In atoms situation,
        the centeral force is the Columb attractive force originating from the nucleus, affecting on
        electrons. Now the vector position becomes $\vec{r} = (r, \theta, \phi)$, as in Figure~\ref{fig:spherical coordinates}. It is important to
        note that in orbiting electron, its angular momentum $\vec{L}$ is only \emph{shart} along 
        a chosen axis while the others components being "fuzzy". Thus, \textbf{it is impossible to
        specify simultaneously any two components of angular momentum}, so we take $L_z$ as the sharp
        component.

        \paragraph{} Using separation of variables so that the spatial wavefunction becomes 
        $\psi(\vec{r}) = \psi(r, \theta, \phi) = R(r)\Theta(\theta)\Phi(\phi)$ so we can find an 
        easy solution for the wavefunction. Then we can find that the sharp observable $|\vec{L}|$, $L_z$, 
        and $E$ for centeral force. We have
        \begin{align*}
            |\vec{L}| = \sqrt{\ell(\ell+1)} \hbar && \ell = 0, 1,\dots, (n-1) \\
            L_z = m_\ell \hbar && m_\ell = 0, \pm 1, \dots, \pm \ell
        \end{align*}

    \section{\color{c3}\MakeUppercase{space quantization}}
        \paragraph{} Last time we know that the orentaion of the angular momentum vector 
        $\vec{L}$ is specified by the numbers $\ell$ and $m_\ell$. The fact that the direction of
        $\vec{L}$ is quantized with respect to an axis is called \textbf{space quantization}. This
        naming is because of \textbf{this property does not originate from the law of force but 
        derives from the space structure itself}.

    \section{\color{c3}\MakeUppercase{atomic hydrogen}}
        Now we want to consider the easiest real atom, that is hydrogen. Using the Columb's law to
        find the radial potential and the wavefunction for centeral force, we can 
        conclude the following result for \textbf{hydrogen-like ions}
        \coloredeq{hydrogen-like equation}{
            \Psi(r, \theta, \phi, t) = R(r)\, Y_\ell^{m_\ell}(\theta, \phi)\, e^{-i \omega t}
        }
        The number $n$ is called the \textbf{priciple quantum number}. It specifies the \textbf{shell}, 
        and for each number $n=1, 2, 3, \dots$ has a corresponding name $K,\, L,\, M,\, \dots$ so on.
        Likewise, the \textbf{orbital quantum number} $\ell = 0, 1, 2, \dots$ represents the subshells, 
        named with $s,\, p,\, d,\, f,\, \dots$ so on. In \textbf{optical transations}, photons carry
        off corresponding energy difference bewtween the states. But not all transations can occur.
        To conserve the \emph{total} angular momentum, the angular momentum of the electron must differ by
        exactly one unit, that is 
        \begin{align*}
            \Delta{\ell}= \pm 1 \quad and \quad \Delta{m_\ell} = 0, \pm 1
        \end{align*}
\end{document}

### 2022-08-04 latex/PHYS300/PHYS210.tex
add LaTeX files

\documentclass{article}

\usepackage{../needed}

\title{PHYS210 - HW2}
\author{Alfaifi, Ammar}
\date{\today}

\includeonly{chapters/chapter4}

\begin{document}
    \maketitle

    \paragraph{}
    This is a solution for the problem 2-22.
    We want to find the equation of motion of a charged particle in an  
    electromagnetic field using \textbf{Lorentz equation}.
    If the electric field vector $\vec{E}$ and the magnetic field vector $\vec{B}$,
    the force on a particle of mass $m$ that carries a charge $q$ and has a velocity
    $\vec{v}$ is given
    \begin{equation} 
        \label{eq:lorentz}   
        \vec{F} = q \vec{E} + q \vec{v} \times \vec{B}
    \end{equation}

    where we assume $v \ll c$.
    \paragraph{(a)} If there is no electeic field and the particle enters the
    the magnetic field in direction perpendicular to the magnetic field. The 
    particle follows a uniform circular motion with acceleration of $a_c=m \frac{v^2}{r}$
    where $r$ is the radius of the path of the circular motion.
    Thus, by Newton's second law we have
    $$
        F = ma_c \Rightarrow qvB sin(90^\circ) = m  \frac{v^2}{r}
    $$
    solving for $r$
    $$ r = \frac{mv}{qB} = \frac{v}{\omega_c} $$
    where $\omega_c \equiv qB/m$, is the \textit{cyclotron frequency}.

    \paragraph{(b)} Choose the z-axis to lie in the direction of $\vec{B}$ and let
    the plane containing $\vec{E}$ and $\vec{B}$ be the yz-plane. Thus,
    $$
        \vec{B} = B \hat{z}, \qquad \vec{E} = E_y \hat{y} + E_z \hat{z}
    $$
    Show that the z component of the motion is given by 
    $$
        z(t) = z_0 + \dot{z_0}t + \frac{qE_z}{2m} t^2
    $$
    where $z(0) \equiv z_0$, and $ \dot{z}(0) \equiv \dot{z_0}$.

    Deviding both sides Eq.1, by $q$, we get
    \begin{equation}
        \frac{\vec{F}}{q} = (E_y \hat{y} + E_z \hat{z}) + (v_x \hat{x}+v_y \hat{y}+v_z \hat{z}) \times B \hat{z}
    \end{equation}

\end{document}

### 2022-08-04 latex/PHYS300/chapters/chapter1.tex
add LaTeX files

\chapter{Series} \label{ch:series}
        \section{Infinite Series}

            \begin{itemize}
                \item A \textit{sequence} is simply a set of quantities, one for each n, represented with $a_n$. 
                \item A \textit{series} is an indicated sum of such represented with $'\sum'$ symbol.
                \item In general \textbf{infinite series} can be written as $ \Rightarrow  a_1 + a_2 + a_3 + \dots + a_n + \dots $
                \item The geomtric series are written as $ \Rightarrow a+ar+ar^2+\dots+ar^{n-1} + \dots$
                \item or in short notation: $ \Rightarrow \sum_{n=1}^{\infty} a_n $
                \item The sum first of first: \textit{n} term is \colorbox{c1}{$ S_n=\frac{a(1-r^n)}{1-r} $}
                \begin{quote}
                    $\bigstar$ The sum of an infinite series is the limit of the sum of $n$ terms as $n\to\infty$.
                    The \textit{sum of series} is \colorbox{c1}{$ S = \lim_{n \to \infty} S_n=\frac{a}{1-r}$}
                \end{quote}

                \begin{quote}
                    $\bigstar$ The geomtric series have a finite sum if and only if $|r|<1$.\\
                    Then the series is called \textit{convergent}, Otherwise it is called \textit{divergent}.
                \end{quote}
                \item the \textit{remainder} (or the remainder after $n$ terms) is \colorbox{c1}{$R_n=S-S_n$}.\\
                Thus, \colorbox{c1}{$\lim_{n\to\infty}\, R_n = 0$.}
                % \coloredeq{R_n=S-S_n \\ \lim_{n\to\infty}\, R_n = 0}.

            \end{itemize}

        \section{Series Tests}

            \begin{itemize}
                \item If the terms of an infinite series do not tend to zero
                (that is, if a $\lim_{n\to\infty}\,$), the series diverges. 
                If $\lim_{n\to\infty}\, a_n=0$, we must test further (Except the alternating series).
                \item An alternating series is an example of a \textit{conditional series}.
                \item Its positive or negative terms (alone) \textit{diverges}, thus we can control 
                to which number (the sum, S) such series approaches.
                \item Physically, We cannot stop at some point and say that the rest of the series is 
                negligible as we could in the bouncing ball problem in Section 1.
                \item But if we specify the order in which the charges are to be placed, then the sum 
                $S$ of the series is determined ($S$ is probably different from $F$ in (8.1) unless 
                the charges are placed alternately).
            \end{itemize}

            \paragraph{$\bigstar$} 
                The convergence or divergence of a series is not affected by multiplying 
                every term of the series by the same nonzero constant. Neither is it affected by 
                changing a finite number of terms (for example, omitting the first few terms).

            \paragraph{$\bigstar$}
                Two \textit{convergent} series $\sum_{n=1}^{\infty}\, a_n$ and $\sum_{n=1}^{\infty}\, b_n$
                may be added (or subtracted) term by term ($a_n+b_n$).
                The resulting series is \textit{convergent}, and its sum is obtained
                by adding (subtracting) the sums of the two given series.

            \paragraph{$\bigstar$}
                The terms of an \textit{absolutely convergent series} may be rearranged in 
                any order without affecting either the \textit{convergence} or the \textit{sum}. This is not 
                true of \textit{conditionally convergent series} as we have seen in Section 8.

        \section{Power Series}
            It is distiguished by having a varible $x^n$ in its terms multiply by a constant.\\
            By definition: \coloredeq{eq:power}{\sum_{n=0}^{\infty}\, a_n(x-a)=a_0+a_1(x-a)^1+a_2(x-a)^2+\dots}.
            \begin{itemize}
                \item The \textit{radius of convergence} $R$ of Eq:\ref{eq:power} depends on $x$ values .
                \item We find $R$ by the \textit{ratio test} so that $L<1$.
            \end{itemize}

        \paragraph{$\bigstar$}
            We see then that a power series (within its interval of convergence) defines a 
            function of $x$, namely $S(x)$.

        \paragraph{Theorems}
            \begin{enumerate}
                \item A power series may be differentiated or integrated term by term; 
                the resulting series converges to the derivative or integral of the function 
                represented by the original series \textit{within} the same interval of convergence as the 
                original series (that is, not necessarily at the endpoints of the interval).
                \item Two power series may be added, subtracted, or multiplied; the resultant series 
                converges at least in the common interval of convergence. You may divide two series if 
                the denominator series is not zero at $x = 0$, or if it is and the zero is canceled by 
                the numerator [as, for example, in $(\sin{x})/x$; see (13.1)]. The resulting series will 
                have \textit{some} interval of convergence (which can be found by the ratio test or more simply 
                by complex variable theory—see Chapter 2, Section 7).
                \item One series may be substituted in another provided that the values of the substituted series 
                are in the interval of convergence of the other series.
                \item The power series of a function is unique, that is, there is just one power series of the 
                form $\sum_{n=0}^{\infty}\, a_nx^n$ which converges to a given function.
            \end{enumerate}

            \paragraph{$\bigstar$}
                Expanding functions can be done using \textit{Taylor series}
                \coloredeq{eq:taylor}{f(x)=\sum_{n=0}^{\infty}\, \frac{f^{(n)}(a)}{n!}(x-a)^n}

            \begin{itemize}
                \item \textit{Maclaurin series} for $f(x)$ is a special case of Eq\eqref{eq:taylor} where $a=0$.
            \end{itemize}

### 2022-08-04 latex/PHYS300/chapters/chapter4.tex
add LaTeX files

\chapter{Vector Analusis}
    \section{Theorems}
        \paragraph{Line Integral} The line integral for $f(x, y)$; $x=x(t)$ \& $y=y(t)$
            \begin{align*}
                \int_C f(x, y)\, ds = \int_a^b f(x(t), y(t))L\, dt
            \end{align*}
            The integral of '$f(x, y)$' along '$C$' with respect to '$x$'. Or,
            \begin{align*}
                \int_C f(x, y)\, dx\, dy = \int_a^b f(x(t), y(t)) x'(t)\, dt
            \end{align*}
            where $x'(t) dt = dx$.

        \starpar $\vec{r}(t) = (1-t)\vec{r_1} + \vec{r_2}$

        \paragraph{Line integral of vector fields}
            If $\vec{F} = P\ihat + Q\jhat + R\khat$, where $Q$, $P$, and $R$ are scalra functions of $(x, y, z)$.
            \begin{align*}
                \int_C \vec{F} \cdot d\vec{r} = \int_a^b \vec{F}(\vec{r}(t)) \cdot \vec{r'}(t) dt = \int_C \vec{F} \cdot \vec{T}\, ds\\
                \Rightarrow \int_C P\, dx + Q\, dy + R\, dz
            \end{align*}
            where $\vec{T}$ is the unit tangent vector.

        \paragraph{Fundamental theorem for line integral}
            \[ \int_a^b F'(x)\, dx = F(b) - F(a) \]

            \[
                \int_C \nabla{f} \cdot d\vec{r} = \int_C df = \vec{r}(b) - \vec{r}(a)
            \]
            which says that the line integral of $\nabla{f}$ is the \underline{net} change in $f$.
            Note that $\oint \vec{F} \cdot d\vec{r} = 0$

        \paragraph{Green's Theorem}
            \[
                \oint_C \vec{F} \cdot d\vec{r} = \oint_R P\, dx + Q\, dy 
                = \int_R \left( \frac{\partial Q}{\partial x} - \frac{\partial P}{\partial y}  \right) \, dA
                \]
            let $ \frac{\partial Q}{\partial x} - \frac{\partial P}{\partial y} = 1$  to get the area $A$.

            In vector form 
            \[
                \oint_R \vec{F} \cdot d\vec{r} = \int_D \nabla{\vec{F}} \cdot \khat\, dA
                \]
            As an special case of Stoke's theorem
            The integral of the tangential component of $\vec{F}$ \underline{along} $C$ is the the double integral of the vertical component of curl$\vec{F}$
            \underline{over} the region $D$ enclosed by $C$. Also we have
            \[
                \oint_R \vec{F} \cdot \hat{n} \, ds = \int_D \nabla \cdot \vec{F}\, dA 
                \]
            The line integral of the normal component of $\vec{F}$ \underline{along} $C$ is the double integral of the divergence of $\vec{F}$ 
            \underline{over} region $D$ enclosed by $C$.

            \starpar tangent $d\vec{r} = dx \ihat + dy \jhat + dz \khat$
            \starpar normal  $ \hat{n} ds = dy \ihat - dx \jhat$

        \paragraph{Stoke's Theorem} 
        \[ 
            \oint_C \vec{F} \cdot d\vec{r} = \int_S \nabla \times \vec{F} \cdot d\vec{S}
            \]
        The line integral \underline{around} the boundry curve of $S$ of the tangential component of $\vec{F}$ is the the surface integral \underline{over}
        $S$ of the normal component of curl$\vec{F}$.

        \paragraph{Divergence Theoprem}
        \[
            \oint_A \vec{F} \cdot d\vec{s} = \int_E \nabla \cdot \vec{F} dV 
            \]
        The flux of $F$ \underline{across} the boundry surface of $E$ is the triple integral of the divergence of $F$ \underline{over} $E$.

### 2022-08-04 latex/PHYS306/Guided Wave.tex
add LaTeX files

\documentclass{article}

\usepackage{amsmath}
\usepackage{amssymb}

\title{PHYS306 - Summary Homework}
\author{Alfaifi, Ammar S.}
\date{\today}

\begin{document}
    \maketitle
    \begin{abstract}
        So far we have dealt with infinite-extent plane waves; now we need
        to consider electromaganetic waves confined to the interior of a hollow
        pipe, called \textbf{wave guide}
    \end{abstract}
    \paragraph{Boundary conditoins} % (fold)
    \label{par:Boundary conditoins}
        We assume the wave guide is perfect conductor, so that $\vec{E}=\vec{0}$
        and $\vec{B}=\vec{0}$ inside the the material itself, thus the boundary
        conditoins at the inner wall are
        \begin{align}
            \label{eq:boundary conditions}
            \begin{cases}
                \vec{E}^{||} = \vec{0} \\
                \vec{B}^\perp = 0
            \end{cases}
        \end{align}
    % paragraph Boundary conditoins (end)

    \paragraph{Wave Equations} % (fold)
    \label{par:Wave Equations}
        Notice that free charges and currents will be induced at the surface
        to enforce these constraints. We are intreseted in monochromatic waves
        that propagate through the tube. Thus, $\vec{E}$ and $\vec{B}$ are in form
        \begin{align}
            \label{eq:fields form}
            \begin{cases}
                \vec{\tilde{E}}(x, y, z, t) = \vec{\tilde{E_0}}(x, y) e^{i(kz-\omega t)}, \\
                \vec{\tilde{B}}(x, y, z, t) = \vec{\tilde{B_0}}(x, y) e^{i(kz-\omega t)}, \\
            \end{cases}
        \end{align}
        Also the electric magnetic field must satisfy Maxwell's equations in the
        interior of the wave guide:
        \begin{align}
            \label{eq:Maxwell's equation}
            \begin{cases}
                \vec{\nabla} \cdot  \vec{E} = 0 \qquad \vec{\nabla} \times \vec{E} = - \frac{\partial \vec{B}}{\partial t} \\
                \vec{\nabla} \cdot  \vec{B} = 0 \qquad \vec{\nabla} \times \vec{B} = \frac{1}{c^2} \frac{\partial \vec{E}}{\partial t} \\
            \end{cases}
        \end{align}
        Now we need just to find those amplitudes, $\vec{\tilde{E_0}}$ and $\vec{\tilde{B_0}}$,
        such that the fields Eq.\eqref{eq:fields form} obey the differential equations 
        Eq.\eqref{eq:Maxwell's equation}, subject to boundary conditions Eq.\eqref{eq:boundary conditions}.
    % paragraph Wave Equations (end)

    \paragraph{Longitudinla Component} % (fold)
    \label{par:Longitudinla Component}
        In general, \textit{confined} eaves are not transverse; in order to fit the boundary
        conditoins we shall have to include longitudinal components $E_z$ and $B_z$:
        \begin{align}
            \label{eq:Amplitudes in componen form}
            \vec{\tilde{E_0}} = E_x \hat{x} + E_y \hat{y} + E_z \hat{z} \qquad 
            \vec{\tilde{B_0}} = B_x \hat{x} + B_y \hat{y} + B_z \hat{z} 
        \end{align}
        where each componen is a function of $x$ and $y$.
    % paragraph Longitudinla Component (end)

    Pluging Eq.9.180 into Maxwell-divergence equations yields:
    \begin{align}
        \label{eq:Z component equations}
        \begin{cases}
            \left[ \frac{\partial^2}{\partial x^2} + \frac{\partial^2}{\partial y^2} +(\omega/c)^2 + k^2 \right] E_z = 0 \\
            \left[ \frac{\partial^2}{\partial x^2} + \frac{\partial^2}{\partial y^2} +(\omega/c)^2 + k^2 \right] B_z = 0
        \end{cases}
    \end{align}

    If $E_z=0$, we call these \textbf{TE waves}; if $B_z=0$, they are called \textbf{TM waves};
    if both vanishes, we call it \textbf{TEM waves}. But, TEM situatoin can never happen in
    a hollow wave guide.

    \paragraph{TE Waves in a Recatngular Guide} % (fold)
    \label{par:TE Waves in a Recatngular Guide}
        Now we consider a wave guide of recatangular shape, with height $a$ and width $b$,
        we want to study the propagation of TE waves.
        Thus, the problem is to solve $B_z$ in Eq.\eqref{eq:Z component equations}, subject
        to the boundary condition of $B$ in Eq.\eqref{eq:boundary conditions}. By separation of
        varaiables,
        \begin{align}
            B_z(x, y) = X(x) Y(y)
        \end{align}

        so that 
        \begin{align}
            \frac{1}{X} \frac{d^2 X}{dx^2} +\frac{1}{Y} \frac{d^2 Y}{dy^2}
            + \left[ (\omega/c)^2 -k^2 ) \right] =0
        \end{align}

        We notice that the $x$- and $y$-dependent terms must be constant, So
        \begin{align}
            \frac{1}{X} \frac{d^2 X}{dx^2} = -k_x^2 \qquad \frac{1}{Y} \frac{d^2 Y}{dy^2} =-k_y^2
        \end{align}

        The general solution is
        \begin{align}
            X(x) = A \sin{(k_x x)} + B \cos{(k_x x)}
        \end{align}
        But the boundary condition for $B_x$ must be zero, that is, $dX/dx=0$. This gives 
        \begin{align}
            k_x=m\pi / a, \quad k_y= n\pi/ b \qquad m, n=0, 1, 2, \dots
        \end{align}
        Finaly, the solution of $TE_{mn}$ mode is
        \begin{align}
            B_z = B_0 \cos{(m\pi / ax)} \cos{(n\pi /bx)}
        \end{align}
        The wave number is given by
        \begin{align}
            k= \sqrt{(\omega/c)^2 -\pi^2 [(m/a)^2 + (n/b)^2]}
        \end{align}
        If the vlaue under the root is negative, there will be no propagating wave, but an
        exponentialy attenuated fileds. Hence we degine the \textbf{cutoff frequency} $\omega_{mn}$
        as, 
        \begin{align}
            \text{lowest possible} < \omega_{mn} \equiv c\pi\sqrt{(m/a)^2 + (n/b)^2} \le \omega
        \end{align}
        The velocities are
        \begin{align}
            v=\frac{\omega}{k}, \qquad v_g= \left( \frac{dk}{d\omega} \right)^{-1}
        \end{align}
    % paragraph TE Waves in a Recatngular Guide (end)
\end{document}

### 2022-08-04 latex/PHYS310/The Casimir Effect.tex
add LaTeX files

\documentclass{article}

\usepackage{amsmath}
\usepackage{graphicx}

\usepackage{biblatex}
\addbibresource{ref.bib}

\title{PHYS310 - The Casimir Effect}
\author{Alfaifi, Ammar}
\date{\today}

\begin{document}
    \maketitle
    \section{Introduction}
        \paragraph{} 
        Physics is the field that never get uninteresting. Uncountable number of phenomena in our
        nature need to have explained. Physicist, most of them, like to divide the field of physics into four
        subfields. That are, classical physics, the study of objects if low speed of motion; special 
        relativity, the study of objects in near-light-speed motion; quantum mechanics, the study of
        particles in scale of atomic sizes; and the quantum field theory, the study of both particles and 
        in high speed. The latter two are the concern in this article. The Casimir, after the Dutch 
        physicists Hendrik Casimir effect is one of the consequences of quantum field theory.

    \section{Quantum Fluctuation}
        \paragraph{}
        What appears to us that is vacuum, it turns out not. In a point in space there appears a phenomenon
        called quantum fluctuation, which is a temporary random change in the amount of energy due to
        Heisenberg's uncertainty principle.

        \paragraph{}
        Vacuum fluctuations appear as virtual particles, which are always created in 
        particle-antiparticle pairs. \cite*{virtual-sa} Moreover, they are created spontaneously without a source of 
        energy, vacuum fluctuations and virtual particles are said to violate the conservation of energy.
        This is theoretically can be allowed because the particles annihilate each other within a time limit 
        determined by the uncertainty principle, so they are not directly observable. Recall that the 
        uncertainty principle in energy and times is given by
        \begin{align}
            \Delta{E} \Delta{t} \ge \hbar / 2
        \end{align}
        where the value of $\hbar / 2$ is almost $5.27286 \times 10^{-35}\, Js$. 
        It means that pairs of virtual particles with energy $\Delta{E}$ and lifetime less than 
        $\Delta{t}$ are created and vanish in empty space. Such particles cannot be detected in direct 
        manner, but their consequences can. As can be seen in the Figure~\ref{Fluctuation}, a picture of simulation 
        illustrate the typical four-dimensional structure of gluon-field configurations averaged over in describing the vacuum properties of QCD.

        \begin{figure}
            \includegraphics[width=0.5\linewidth]{fluc.png}
            \centering
            \caption{
                The volume of the box is 2.4 by 2.4 by 3.6 fm, big enough to hold a couple of protons. 
                Contrary to the concept of an empty vacuum, QCD induces chromo-electric and
                chromo-magnetic fields throughout space-time in its lowest energy state.
                [Center for the Subatomic Structure of Matter (CSSM) and Department of Physics, 
                University of Adelaide, 5005 Australia]\cite*{sim}}
            \label{Fluctuation}
        \end{figure}

    \section{The Casimir Effect}
        \begin{figure}
            \centering
            \includegraphics[width=0.5\linewidth]{plates.png}
            \caption{Two close parallel uncharged conducting plates. The green waves
            represent the quantum fluctuation in electromagnetic field.}
            \label{Plates}
        \end{figure}

        \paragraph{}
        The Casimir–Polder force. After a discussion with Niels Bohr, who suggested it had something to do with zero-point energy, Casimir alone formulated the theory predicting a force between neutral conducting plates in 1948, which is called the Casimir effect in the narrow sense.

        \paragraph{}
        The Casimir effect is force that arises from the quantum fluctuation of electromagnetic field in space, e.g., between two close parallel uncharged conducting plates, as in Figure~\ref{Plates}.
        To understand this effect consider those two close parallel conducting plates, they will trap 
        waves of fluctuating electromagnetic field as discussed in the previous section. But notice that 
        the longer-wavelength waves between the plates are vanished. Meaning that as the distance between
        the plates shrink the long wavelength waves cannot be fit. \cite*{casimir-sa} As result, the total energy exist between the 
        plates will be less, compared to elsewhere in the vacuum.

        \paragraph{}
        Due to this process the plates will attract to each other, as if they were connected to spring.
        This force can be given quantitatively by the formula \cite{byjus}
        \begin{align}
            F = \frac{\pi h c}{480} \frac{A}{L^4}
        \end{align}
        where $h$ is the Planck's constant, $A$ is the area of the plate, and $L$ is the spacing between the two plates. In specific, for two-one-meter square plates and with spacing between them, $10,000$ nanometer, they will exhibit a force order of $10^-7$ Newton.

    \section{Measuring the Casimir Effect}
        \paragraph{}
        Marcus Sparnaay, in 1958, conducted one of the first experimental tests in a delicate and challenging experiment with parallel plates, obtaining results not in contradiction with the Casimir theory but with significant experimental errors. A more accurate measurement of the Casimir effect was conducted by Steve K. Lamoreaux of Los Alamos National Laboratory and Umar Mohideen and Anushree Roy of California, Riverside. In practice, we use one flat plate and another plate that is a part of a sphere with a large radius to measure the Casimir effect because using two parallel plates would require accurate alignment to ensure they are parallel. Finally, in 2001, using microresonators, a group at the University of Padua succeeded in measuring the Casimir force between parallel plates.

    \section{Conclusion}
    In this article we leaned about the fascinating phenomenon that happens in vacuum, i.e., the quantum fluctuations of electromagnetic field. Moreover, we saw one of its consequences, the Casimir effect. In short, it is a force that appears between two uncharged conducting plates with very small spacing. This force tends to 
    push the plates into each other. This phenomenon has applications in the field nanotechnologies as well as affects in manufacturing the semiconductors' chips.

    \printbibliography
\end{document}

### 2022-12-07 latex/PHYS410/presentation/main.tex
add slides

\documentclass{loyola-beamer}
\renewcommand{\titlelogo}{logo/luc-rev.svg}
\renewcommand{\slidelogo}{logo/luc.svg}
\usepackage{physics}

\title{Coherent States}
% \subtitle{Made with Beamer}
\author{Alfaifi, Ammar}
\institute{KFUPM}
\renewcommand{\slidefoot}{Department of Physics - PHYS410}

\begin{document}

\begin{titleframe}{}
    \maketitle
\end{titleframe}

\begin{frame}{Contents}
    \tableofcontents
\end{frame}

% Start of the slides

\begin{titleframe}{Comparison}
    Stationary States vs. Coherent States
\end{titleframe}

\section{Stationary States}

\begin{frame}{Stationary States}
    Stationary states are eigenstates of the Hamiltonian $\hat{H}$

    \vspace{\baselineskip}
    For quantum harmonic oscillator the Hamiltonian is

    \begin{align*}
        \hat{H} = \hbar \omega \left(  a^\dagger a + \frac{1}{2} \right)
        \label{eq:Hamiltonian}
    \end{align*}

    \begin{align*}
        \hat{H} |n \rangle = E_n |n \rangle
    \end{align*}

    \begin{align*}
        \hat{a} |n \rangle = \sqrt{n} |{n-1} \rangle
    \end{align*}

\end{frame}

% New slide
\section{Coherent States}

\begin{frame}{Coherent States}
    Mathematically, a coherent state is defined to be the (unique)
    eigenstates of the annihilation operator â with corresponding eigenvalue $\alpha$.

    \vspace{\baselineskip}
\end{frame}

\begin{frame}{What are Coherent States?}

    Coherent states are defined as eigenstates of the annihilation operator. Specifically, they are defined as:
    \begin{equation}
        \hat{a} \left| \alpha \right\rangle = \alpha \left| \alpha \right\rangle
    \end{equation}
    since $\hat{a}$ is not hermitian, $\alpha$ is a complex number and $\left|\alpha\right\rangle$ is the coherent state.

\end{frame}

\begin{frame}{Derive Coherent States}

    We write the coherent states in the basis of the Hamiltonian.

    \begin{align*}
        | \alpha \rangle  = \sum_{n=0}^{\infty} \, c_n(\alpha) |\alpha \rangle;
        \qquad c_n(\alpha) = \langle \alpha | \alpha \rangle
    \end{align*}

    \begin{align*}
        \hat{a} \ket{\alpha} & = \sum_{n=0}^\infty c_n(\alpha)\, \hat{a} \ket{n}
        = \sum_{n=1}^\infty c_n(\alpha)\sqrt{n}\, \ket{n-1}                            \\
                             & = \sum_{n=0}^\infty c_{n+1}(\alpha)\sqrt{n+1}\, \ket{n}
    \end{align*}

    \begin{align*}
        \hat{a} \ket{\alpha} = \alpha \ket{\alpha} \Rightarrow
        \sum_{n=0}^\infty c_{n+1}(\alpha)\sqrt{n+1}\, \ket{n}
        = \alpha \sum_{n=0}^\infty c_{n}(\alpha)\, \ket{n}
    \end{align*}

\end{frame}

\begin{frame}{Derive Coherent States}
    \begin{align*}
        c_n(\alpha) = \frac{\alpha^n}{\sqrt{n!}}\, c_0
    \end{align*}

    The coherent states are
    \begin{align*}
        \ket{\alpha} = \sum_{n=0}^\infty \frac{\alpha^n}{\sqrt{n!}} c_0(\alpha) \, \ket{n}
    \end{align*}

    Normalization
    \begin{align*}
        1= \bra{\alpha}\ket{\alpha} & = | c_0(\alpha) |^2
        \overbrace{\sum_{n=0}^\infty \frac{|\alpha|^{2n}}{n!}}^{e^{|\alpha|^2}} \\
        | c_0(\alpha) |^2           & = e^{-|\alpha|^2}
        \Rightarrow c_0(\alpha) = e^{-|\alpha|^2/2}
    \end{align*}

\end{frame}

\begin{frame}{Derive Coherent States}
    Finally, the coherent state in the energy basis is
    \begin{align*}
        \ket{\alpha} = e^{-|\alpha|^2/2}\,
        \sum_{n=0}^\infty \frac{\alpha^n}{\sqrt{n!}} \, \ket{n}
    \end{align*}

\end{frame}

\begin{frame}{Properties of Coherent States}

    Coherent states have several important properties which make them useful for many applications.

    \begin{itemize}
        \item Coherent states are minimum-uncertainty states: they have the minimum possible uncertainty in both position and momentum.
        \item Coherent states have a Gaussian wavefunction, which makes them easy to work with.
        \item Coherent states are stable: they remain unchanged over time.
        \item Coherent states can be used to approximate classical behavior.
    \end{itemize}

\end{frame}

\begin{frame}{Applications of Coherent States}

    \begin{itemize}
        \item coherent states are used in quantum optics to describe light.
        \item coherent states can be used to simulate the behavior of classical systems.
        \item coherent states can be used to encode quantum information.
        \item coherent states are used in quantum computing to implement quantum algorithms.
    \end{itemize}

\end{frame}

\section{Time Evolution}

\begin{frame}{Time Evolution}
    \begin{align*}
        \ket{\psi(0)} = \ket{\alpha_0} = e^{-|\alpha_0|^2/2}\,
        \sum_{n=0}^\infty \frac{\alpha_0^n}{\sqrt{n!}} \, \ket{n}
    \end{align*}

    \begin{block}{Harmonic oscillator}
        $\hat{H}$ is time independent: conservative system.
    \end{block}

    \begin{align*}
        \ket{\psi(t)} = e^{-|\alpha_0|^2/2}\,
        \sum_{n=0}^\infty \frac{\alpha_0^n}{\sqrt{n!}} \,  e^{-iE_nt\hbar} \ket{n}
    \end{align*}
\end{frame}

\begin{frame}{Time Evolution}
    \begin{align*}
        \ket{\psi(t)} e^{-i\omega t/2} e^{-|\alpha_0|^2 /2}
        \sum_{n=0}^\infty \frac{(\alpha_0 e^{-i\omega t})^n}{\sqrt{n!}}\, \ket{n}
    \end{align*}
    let $\alpha \equiv \alpha_0 e^{-i\omega t}$

    \begin{align*}
        \ket{\psi(0)} = \ket{\alpha_0} \rightarrow \ket{\psi(t)} = e^{-i\omega t/2} \ket{\alpha};
    \end{align*}

    \begin{block}{Conclusion}
        A coherent state stays coherent at all times.
    \end{block}
\end{frame}

\section{Creation Operator}

\begin{frame}{About Creation Opreator}
    \begin{alertblock}{Note}
        Creation operator has no eigenstates.
    \end{alertblock}
    \begin{align*}
        \hat{a}^\dagger \ket{\lambda} = \lambda \ket{\lambda}
    \end{align*}
    \begin{align*}
        \sum_{n=0}^\infty c_n \sqrt{n+1}\, \ket{n+1}
        = \lambda \sum_{n=0}^\infty c_n\, \ket{n}
    \end{align*}
    \begin{align*}
        \ket{0} & : \quad 0 = \lambda c_0 \Rightarrow c_0 =0               \\
        \ket{1} & : \quad c_0 = \lambda c_0 \Rightarrow c_1 =0             \\
                & \vdots                                                   \\
        \ket{n} & : \quad c_{n-1} \sqrt{n}=\lambda c_n \Rightarrow c_n = 0
    \end{align*}

\end{frame}

\begin{frame}{Conclusion}

    In conclusion, coherent states are special types of quantum states which have several
    useful properties and applications. They are widely used in quantum physics, quantum optics,
    and quantum information.

\end{frame}

\begin{titleframe}{Thank you for listening!}

\end{titleframe}

\end{document}

### 2022-12-11 latex/PHYS308/report.tex
dunno ;

\documentclass[a4paper, 11pt]{report}

\usepackage{graphicx}
\usepackage{amsmath}

\title{Mobile Phone Detector based on Op Amp}
\author{Alfaifi, Ammar}
\date{\today}

\begin{document}

\maketitle

\begin{abstract}
	This report presents the design and implementation of a mobile phone detector based on an operational amplifier (op amp). The detector uses the op amp to amplify the electromagnetic (EM) radiation emitted by a mobile phone and trigger an alarm when the radiation reaches a certain level. The design and implementation process, as well as the results and limitations of the detector, are discussed in detail.
\end{abstract}

\section{Introduction}

Mobile phones have become an indispensable tool for communication, information access, and entertainment. However, their use can be disruptive in certain settings such as classrooms, libraries, and theaters. In such settings, it is important to ensure that mobile phones are turned off or put on silent mode to avoid disturbing others.

To enforce this rule, many institutions and organizations use signs and announcements to remind users to turn off their phones. However, these methods are not always effective and can be ignored by some users. Therefore, there is a need for a more effective and automated solution to detect and alert users about the use of mobile phones in restricted areas.

In this project, I present a simple and effective mobile phone detector that uses an operational amplifier (op amp) to sense the electromagnetic field (EMF) generated by a mobile phone. The detector can be used to alert users to turn off their phones in restricted areas.

\section{Design and Implementation}

The mobile phone detector consists of an op amp circuit that senses the EMF generated by a mobile phone. The circuit is shown in Figure 1.

\begin{figure}[ht]
	\centering
	\includegraphics[width=0.5\textwidth]{detector-circuit.png}
	\caption{Circuit diagram of the mobile phone detector}
	\label{fig:circuit-diagram}
\end{figure}

The circuit diagram of the mobile phone detector is shown in Figure \ref{fig:circuit-diagram}.
The op amp used in this circuit is the LM358, which is a low-power, dual-channel op amp with high gain and a wide bandwidth. The EMF generated by a mobile phone is picked up by an antenna and fed into the non-inverting input of the op amp. The inverting input is connected to a reference voltage, which is generated using a voltage divider circuit.

The output of the op amp is fed into a comparator, which compares the output voltage with a threshold voltage. If the output voltage exceeds the threshold voltage, the comparator outputs a high signal, which can be used to trigger an alert such as a buzzer or LED.

The gain of the op amp circuit can be adjusted using the feedback resistor, which is connected between the output and the inverting input. A higher value of the feedback resistor results in a higher gain and a more sensitive detector.

\begin{figure}[ht]
	\centering
	\includegraphics[width=0.5\textwidth]{breadboard.jpeg}
	\caption{My first testing for the circuit in lab.}
	\label{fig:breadboard}
\end{figure}

Figure \ref{fig:breadboard} shows my first testing for the circuit in lab. The first, it was not woking as intended.
At later time, I re build the circuit again but with op amp LM358P chip and it started responding. Then I begun
tuning the circuit to give best results, by changing the feedback resistor from $2M\Omega$ up to $3\Omega$ and
using different values for the capacitor in series with the antenna.

\begin{figure}[ht]
	\centering
	\includegraphics[width=0.5\textwidth]{Mobile-Phone-Detector-PCB.png}
	\caption{PCB design for the mofile phone detect circuit found in internet}
	\label{fig:PCB}
\end{figure}

From the Figure \ref{fig:PCB} we can contact online company to print this circuit in two-layer printed board.
This wil give us an idea how the circuit would look like in real applicaitons, as well as testing it further.

\section{Results and Limitations}

The mobile phone detector was tested with a variety of mobile phones, including both switched on and switched off phones. The results showed that the detector was able to accurately detect the EM radiation emitted by mobile phones, regardless of whether they were switched on or off.

However, the detector has some limitations. First, it is sensitive to other sources of EM radiation, such as fluorescent lights and other electronic devices. This can cause false alarms and reduce the overall accuracy of the detector. Second, the detection range of the detector is limited, and it is only effective within a few feet of the mobile phone.

\section{Conclusion}

In conclusion, the mobile phone detector based on an op amp is an effective and low-cost project, to deletct
the existing of a mobile phone around via EM radiation emitted by the GSM signals. However, the circuit require
more optimzation to tune the sensitivity, and might adding a varialbe capacitor help catching mltiple
other frequencies. Moreover, this circuit can be brought from prototype into real world by printing it as
PCB. This will reduce it sized dramaticaly, and make it easy to carry.

\end{document}

### 2022-12-11 latex/PHYS441/paper.tex
dunno ;

\documentclass{article}
\usepackage{amsmath}
\usepackage{graphicx}

\title{Light-Light Scattering: Feynman Diagrams}
\author{Alfaifi, Ammar \and Al-Ali, Muhammad}
\date{June 6, 2022}

\begin{document}

\maketitle

\begin{abstract}
	Light-light scattering, also known as photon-photon scattering, is a quantum electrodynamic (QED) phenomenon in which two photons interact and scatter off each other. In this paper, we discuss the theory of light-light scattering and present the Feynman diagrams for the process. We also discuss the implications of light-light scattering and its experimental verification.
\end{abstract}

\section{Introduction}

Light-light scattering, also known as photon-photon scattering, is a fundamental quantum electrodynamic (QED) process in which two photons interact and scatter off each other. The phenomenon was first predicted by Einstein in 1917 \cite{einstein1917theory}, who showed that photons have momentum and can exert a force on each other. However, it was not until the development of quantum field theory in the 1950s that a complete theory of light-light scattering was developed.

Light-light scattering is a purely quantum mechanical process and cannot be explained using classical physics. It is a rare phenomenon, as the probability of two photons interacting is very small. However, the process has important implications for the study of the fundamental nature of light and its interactions with matter.

In this paper, we discuss the theory of light-light scattering and present the Feynman diagrams for the process. We also discuss the implications of light-light scattering and its experimental verification.

\section{Theory of Light-Light Scattering}

The theory of light-by-light scattering was first proposed by physicist Richard Feynman in 1948. Feynman used a mathematical technique called Feynman diagrams to represent the interaction between particles. In a Feynman diagram, photons are represented by lines, and their interactions are represented by vertices where the lines meet.

Feynman diagrams show that light-by-light scattering is a very rare event because photons do not have any electric charge, and therefore, do not experience the electromagnetic force. This means that photons do not interact with each other through the exchange of virtual particles such as gluons or W/Z bosons.

However, photons can still interact with each other through the weak force, which is one of the four fundamental forces of nature. The weak force is responsible for the radioactive decay of atoms, but it is much weaker than the electromagnetic force. This is why light-by-light scattering is a very rare phenomenon.

In 2015, a team of scientists at the Large Hadron Collider (LHC) at CERN observed light-by-light scattering for the first time. The team used the LHC to collide two high-energy photons, which resulted in the scattering of the photons. This observation confirmed the theoretical predictions of light-by-light scattering and provided new insights into the fundamental nature of light.

Light-by-light scattering has many potential applications in the fields of physics, astronomy, and engineering. For example, it could be used to study the properties of dark matter and the structure of the universe. It could also be used to develop new technologies such as laser-based particle accelerators and high-precision sensors.

The theory of light-light scattering is based on quantum electrodynamics (QED), which is the quantum mechanical theory of the electromagnetic field and its interactions with charged particles. In QED, photons are treated as particles that can interact with each other through the exchange of virtual particles such as electron-positron pairs.

The Feynman diagrams for light-light scattering are shown in Figure 1. The process can be thought of as the interaction of two photons, labeled as 1 and 2, which scatter off each other and produce two outgoing photons, labeled as 3 and 4. The process is mediated by the exchange of a virtual electron-positron pair, which is created and annihilated in the process.

\begin{figure}[h]
	\centering
	\includegraphics[width=0.5\linewidth]{Feynman-diagram.png}
	\caption{Feynman diagrams for light-light scattering.}
	\label{fig:light-light-scattering-feynman}
\end{figure}

The probability of light-light scattering can be calculated using the Feynman diagrams and the rules of quantum electrodynamics. The result is a very small probability, which makes the phenomenon difficult to observe experimentally.

\section{Implications and Experimental Verification}

The phenomenon of light-light scattering has important implications for the study of the fundamental nature of light and its interactions with matter. It is a fundamental QED process that cannot be explained using classical physics and provides evidence for the quantum nature of light.

Light-light scattering has also been used to test the validity of QED. The predicted probability of the process has been verified experimentally to a high degree of precision, providing strong evidence for

\section{Conclusion}

In conclusion, light-by-light scattering is a fascinating and complex phenomenon that is still not fully understood. With the help of advanced technologies such as the LHC and Feynman diagrams, we can continue to study and learn more about this elusive interaction.

\end{document}

### 2022-12-13 latex/PHYS441/paper.tex
add articles

@@
 \usepackage{graphicx}
+\usepackage{biblatex}
 
@@
 	\centering
-	\includegraphics[width=0.5\linewidth]{Feynman-diagram.png}
+	\includegraphics[width=0.5\linewidth]{figures/Feynman-diagram.png}
 	\caption{Feynman diagrams for light-light scattering.}
@@
 
+\subsection{ATLAS Observation}
+Light-by-light scattering is a very rare phenomenon in which two photons – particles of light – interact,
+producing another pair of photons. This process was among the earliest predictions of quantum electrodynamics (QED),
+the quantum theory of electromagnetism, and is forbidden by classical physics theories (such as
+Maxwell's theory of electrodynamics).
+
+Direct evidence for light-by-light scattering at high energy had proven elusive for decades, until the Large Hadron
+Collider (LHC) began its second data-taking period (Run 2). Collisions of lead ions in the LHC
+provide a uniquely clean environment to study light-by-light scattering. Bunches of lead ions that are accelerated to
+very high energy are surrounded by an enormous flux of photons. Indeed, the coherent action from the large number of
+82 protons in a lead atom with all the electrons stripped off (as is the case for the lead ions in the LHC)
+give rise to an electromagnetic field of up to 1025 Volt per metre. When two lead ions pass close by each
+other at the centre of the ATLAS detector, but at a distance greater than twice the lead ion radius, those
+photons can still interact and scatter off one another without any further interaction between the lead ions, as the reach of
+the (much stronger) strong force is bound to the radius of a single proton. These interactions are
+known as ultra-peripheral collisions.
+
+`The ATLAS Collaboration has reported the observation of light-by-light scattering with a significance
+beyond 8 standard deviations.`
+
+\begin{figure}[ht]
+	\centering
+	\includegraphics[width=0.5\linewidth]{figures/LbyL-fig2.png}
+	\caption{Invariant mass distribution of the measured final state photon pairs (markers), compared
+		to the expected light-by-light scattering signal (red line) and expected background contributions
+		(shaded areas). (Image: ATLAS Collaboration/CERN)}
+	\label{fig:Atlas data}
+\end{figure}
+
+In a result published in Nature Physics in 2017, the ATLAS Collaboration found thirteen candidate events for light-by-light scattering in lead-lead collision data recorded in 2015, for 2.6 events expected from background processes.
+The corresponding significance of this result was 4.4 standard deviations – making it the first direct evidence of high-energy light-by-light scattering.
+
+\begin{figure}[ht]
+	\centering
+	\includegraphics[width=0.7\linewidth]{figures/EventDisplay_LbyL.png}
+	\caption{ATLAS event display showing the energy deposits of two photons in the electromagnetic
+		calorimeter (green) on opposite sides and no other activity in the detector, which is the clean signature
+		of light-by-light scattering. The Feynman diagram of this process is shown in the lower right corner.
+		(Image: ATLAS Collaboration/CERN)}
+	\label{fig:Atlas_event}
+\end{figure}
+
+\subsection{Light-by-Light Scattering at Low Energies}%
+
+\subsection{Positron Production in Multiphoton Light-by-Light Scattering}%
+\label{subsec:positron}
+
+\label{subsec:light-light-low-energy}
+A signal of 106±14 positrons above background has been observed in collisions of a low-emittance 46.6 GeV electron beam with terawatt pulses from a Nd:glass laser at 527 nm wavelength in an experiment at the Final Focus Test Beam at SLAC.
+The positrons are interpreted as arising from a two-step process in which laser photons are backscattered to GeV energies by the electron beam followed by a collision between the high-energy photon and several laser photons to produce an electron-positron pair.
+These results are the first laboratory evidence for inelastic light-by-light scattering involving only real photons.
+
 \section{Conclusion}
@@
 
+    \printbibliography
 \end{document}

### 2022-12-14 latex/PHYS441/paper.tex
add figures

@@
 \usepackage{biblatex}
+\addbibresource{ref.bib}
 
@@
 Light-light scattering, also known as photon-photon scattering, is a fundamental quantum electrodynamic (QED) process in which two photons interact and scatter off each other.
-The phenomenon was first predicted by Einstein in 1917 \cite{einstein1917theory}, who showed that photons have momentum and can exert a force on each other.
+The phenomenon was first predicted by Einstein in 1917 \cite{einstein}, who showed that photons have momentum and can exert a force on each other.
 However, it was not until the development of quantum field theory in the 1950s that a complete theory of light-light scattering was developed.
@@
 
-The Feynman diagrams for light-light scattering are shown in Figure 1.
+The Feynman diagrams for light-light scattering are shown in Figure \ref{fig:light-light-scattering-feynman}.
 The process can be thought of as the interaction of two photons, labeled as 1 and 2, which scatter off each other and produce two outgoing photons, labeled as 3 and 4.
@@
 	\centering
-	\includegraphics[width=0.7\linewidth]{figures/EventDisplay_LbyL.png}
+	\includegraphics[width=0.5\linewidth]{figures/EventDisplay_LbyL.png}
 	\caption{ATLAS event display showing the energy deposits of two photons in the electromagnetic
@@
 \subsection{Light-by-Light Scattering at Low Energies}%
+\begin{figure}[!th]
+	\centering
+	\includegraphics[width=0.5\linewidth]{figures/photon-photon-scattering.jpg}
+	\caption{The cross-section for photon-photon scattering as a function of photon frequency (from Lifshitz et al. - 1982) }
+	\label{fig:corss-secction}
+\end{figure}
+\begin{figure}[!th]
+	\centering
+	\includegraphics[width=0.6\linewidth]{figures/photons.jpg}
+	\caption{Three different types of interactions between photons.
+In the left picture an incoming photon interacts with e.g.
+the strong external field from an atomic nucleus, resulting in an outgoing photon with new characteristics (Delbrück scattering); in the middle picture a single photon interacts with e.g.
+a strong external magnetic field as to produce two low-frequency photons (photon splitting); finally, in picture to the right two photons interacts directly via the quantum vacuum, producing two scattered photons.
+(Courtesy: Mattias Marklund)}
+	\label{fig:photon-types}
+\end{figure}
+photon–photon scattering is a non-classical effect arising in quantum electrodynamics (QED) due to virtual electron–positron pairs in vacuum.
+Under every-day circumstances, the effect is very weak (see FIG.
+1).
+However, under the right conditions the interaction between photons and these virtual pairs will result in what is known as photon–photon collisions.
+Close relatives to this effect is Delbrück scattering and photon splitting (see FIG.
+2).
+Formulated as an effective field theory, using the Heisenberg–Euler Lagrangian (Heisenberg \& Euler (1936); Schwinger (1951)), such scattering results in nonlinear corrections to Maxwell’s vacuum equations, similar in form to what is known as Kerr nonlinearities in nonlinear optics.
+The effective self-interaction term in Maxwell’s equations is small (proportional to the fine structure constant squared), which means that the field strengths need to reach appreciable values until such effects become pronounced (Marklund \& Shukla (2006); Mourou et al.
+(2006)).
+However, the smallness of the photon-photon scattering cross section has been argued to be a possible window to new physics (Anoniadis (1998); Arkani-hamed et al.
+(1998); Cheung (1999); Davoudiasl (1999)), such as weak scale quantum gravity.
+Physical implications, as well as possible detection techniques, of the effects of photon–photon scattering have attracted interest since the 1930’s (for surveys, see Refs.
+Marklund \& Shukla (2006) and Mourou et al.
+(2006) and references therein), and the topic is hotter than ever.
 
@@
 The positrons are interpreted as arising from a two-step process in which laser photons are backscattered to GeV energies by the electron beam followed by a collision between the high-energy photon and several laser photons to produce an electron-positron pair.
-These results are the first laboratory evidence for inelastic light-by-light scattering involving only real photons.
+These results are the first laboratory evidence for inelastic light-by-light scattering involving only real photons.\cite{PhysRevLett.79.1626}
 
 \section{Conclusion}
-
 In conclusion, light-by-light scattering is a fascinating and complex phenomenon that is still not fully understood.
@@
 
+\newpage
 \printbibliography
+
 \end{document}

### 2022-12-15 latex/PHYS441/paper.tex
almost done

@@
 Light-light scattering, also known as photon-photon scattering, is a fundamental quantum electrodynamic (QED) process in which two photons interact and scatter off each other.
-The phenomenon was first predicted by Einstein in 1917 \cite{einstein}, who showed that photons have momentum and can exert a force on each other.
+The phenomenon was first predicted by Einstein in 1917~\cite{einstein}, who showed that photons have momentum and can exert a force on each other.
 However, it was not until the development of quantum field theory in the 1950s that a complete theory of light-light scattering was developed.
@@
 
-The Feynman diagrams for light-light scattering are shown in Figure \ref{fig:light-light-scattering-feynman}.
-The process can be thought of as the interaction of two photons, labeled as 1 and 2, which scatter off each other and produce two outgoing photons, labeled as 3 and 4.
-The process is mediated by the exchange of a virtual electron-positron pair, which is created and annihilated in the process.
-
 \begin{figure}[h]
@@
 \end{figure}
+The Feynman diagrams for light-light scattering are shown in Figure~\ref{fig:light-light-scattering-feynman}.
+The process can be thought of as the interaction of two photons, labeled as 1 and 2, which scatter off each other and produce two outgoing photons, labeled as 3 and 4.
+The process is mediated by the exchange of a virtual electron-positron pair, which is created and annihilated in the process.
 
@@
 \subsection{ATLAS Observation}
+
+\begin{figure}[ht]
+	\centering
+	\includegraphics[width=0.5\linewidth]{figures/LbyL-fig2.png}
+	\caption{Invariant mass distribution of the measured final state photon pairs (markers), compared
+		to the expected light-by-light scattering signal (red line) and expected background contributions
+		(shaded areas). (Image: ATLAS Collaboration/CERN)}
+	\label{fig:Atlas data}
+\end{figure}
 Light-by-light scattering is a very rare phenomenon in which two photons – particles of light – interact,
@@
 
+\begin{figure}[ht]
+	\centering
+	\includegraphics[width=0.5\linewidth]{figures/EventDisplay_LbyL.png}
+	\caption{ATLAS event display showing the energy deposits of two photons in the electromagnetic
+		calorimeter (green) on opposite sides and no other activity in the detector, which is the clean signature
+		of light-by-light scattering. The Feynman diagram of this process is shown in the lower right corner.
+		(Image: ATLAS Collaboration/CERN)}
+	\label{fig:Atlas_event}
+\end{figure}
 Direct evidence for light-by-light scattering at high energy had proven elusive for decades, until the Large Hadron
@@
 
-\begin{figure}[ht]
-	\centering
-	\includegraphics[width=0.5\linewidth]{figures/LbyL-fig2.png}
-	\caption{Invariant mass distribution of the measured final state photon pairs (markers), compared
-		to the expected light-by-light scattering signal (red line) and expected background contributions
-		(shaded areas). (Image: ATLAS Collaboration/CERN)}
-	\label{fig:Atlas data}
-\end{figure}
-
 In a result published in Nature Physics in 2017, the ATLAS Collaboration found thirteen candidate events for light-by-light scattering in lead-lead collision data recorded in 2015, for 2.6 events expected from background processes.
@@
 
-\begin{figure}[ht]
-	\centering
-	\includegraphics[width=0.5\linewidth]{figures/EventDisplay_LbyL.png}
-	\caption{ATLAS event display showing the energy deposits of two photons in the electromagnetic
-		calorimeter (green) on opposite sides and no other activity in the detector, which is the clean signature
-		of light-by-light scattering. The Feynman diagram of this process is shown in the lower right corner.
-		(Image: ATLAS Collaboration/CERN)}
-	\label{fig:Atlas_event}
-\end{figure}
+\subsection{Light-by-Light Scattering at Low Energies}%
 
-\subsection{Light-by-Light Scattering at Low Energies}%
 \begin{figure}[!th]
@@
 \end{figure}
+
 \begin{figure}[!th]
@@
 photon–photon scattering is a non-classical effect arising in quantum electrodynamics (QED) due to virtual electron–positron pairs in vacuum.
-Under every-day circumstances, the effect is very weak (see FIG.
-1).
+Under every-day circumstances, the effect is very weak, see Figure~\ref{fig:corss-secction}.
 However, under the right conditions the interaction between photons and these virtual pairs will result in what is known as photon–photon collisions.
-Close relatives to this effect is Delbrück scattering and photon splitting (see FIG.
-2).
+Close relatives to this effect is Delbrück scattering and photon splitting, see Figure~\ref{fig:photon-types}.
 Formulated as an effective field theory, using the Heisenberg–Euler Lagrangian (Heisenberg \& Euler (1936); Schwinger (1951)), such scattering results in nonlinear corrections to Maxwell’s vacuum equations, similar in form to what is known as Kerr nonlinearities in nonlinear optics.
@@
 The positrons are interpreted as arising from a two-step process in which laser photons are backscattered to GeV energies by the electron beam followed by a collision between the high-energy photon and several laser photons to produce an electron-positron pair.
-These results are the first laboratory evidence for inelastic light-by-light scattering involving only real photons.\cite{PhysRevLett.79.1626}
+These results are the first laboratory evidence for inelastic light-by-light scattering involving only real photons.\cite{physrevlett}
+
+\section{Delbruck scattering}
+
+\begin{figure}[!th]
+	\centering
+	\includegraphics[width=0.5\linewidth]{figures/photon-photon-scattering.jpg}
+	\caption{vacuum polarization feynman diagram.}
+	\label{fig:vacuum-polarization}
+\end{figure}
+
+Delbruck scattering to understand it first we need to understand a process called vacuum polarization.
+Vacuum polarization is a random process by which a photon produces a pair of virtual fermion and anti-fermion and then annihilates repeatedly, the Feynman diagram of this process can be seen from the graph Figure~\ref{fig:vacuum-polarization}.
+Since this process involves charged particles (electron and anti-electron) this pair will create a virtual dipole and then would interact with other charged particles especially if the process occurred in nuclei.
+As a consequence of these interactions on the nucleus, the virtual electron will be attracted, and the virtual positron will be repelled creating a dipole that can cause shielding and reduce the charge if you were to observe from far enough.
+
+\begin{figure}[!th]
+	\centering
+	\includegraphics[width=0.5\linewidth]{figures/lamb-shift.png}
+	\caption{the lamb shift compared to the prediction of Bohr and Dirac}
+	\label{fig:lamb-shift}
+\end{figure}
+This shielding due to the vacuum polarization results in a deviation of the experimental Landé g-factor of the electron from the theoretical value and a difference in energy between two energy levels that was later called “Lamb shift” Figure~\ref{fig:lamb-shift}, which contradicts the prediction of the Dirac equation.
+These theoretical deviations were calculated by the genius Hans Bethe, and what is funny is that he did the calculation on his return train ride when he was returning to Cornell after attending a select workshop in the summer of 1947 in Shelter Island.
+Overall, Vacuum polarization was first discussed in papers by Paul Dirac and Werner Heisenberg in 1934, and the effects of it were calculated to first order in the coupling constant by Robert Serber and Edwin Uehling in 1935.
+But the effect of vacuum polarization of the electron and anti-electron was first observed in the 1940s, and for the quark and anti-quark, it was in the early 1970s.
+Due to this phenomenon, the notion of vacuum changed, this is because the vacuum is not empty anymore but will have some fluctuation and the presence of this virtual particle will disturb the electromagnetic field, however, the fluctuation and the presence of these charged particles will follow the energy-time Heisenberg uncertainty principle so that they will annihilate quickly.
+To sum up, vacuum polarization is the virtual fermion-antifermion produced from vacuum-polarized disturbed electromagnetic field.
+
+\begin{figure}[!th]
+	\centering
+	\includegraphics[width=0.5\linewidth]{figures/fyenman-diagram-Delbruck.png}
+	\caption{Feynman diagrams for Delbruck scattering to lowest order, where x denotes Columb fields.}
+	\label{fig:Feynman-Delbruck}
+\end{figure}
+Now, Delbruck scattering is the elastic scattering of photons in an electromagnetic field of nuclei (mostly heavy nuclei) as a consequence of the vacuum polarization, in which the two-photon scatters via electron-positron pairs as in the vacuum polarization Figure~\ref{fig:Feynman-Delbruck}.
+This scattering depends on the energy of the photons and the angle, for each case whether high energy and small angles or intermediate energy and moderately large angles or at low energies but in the whole angular range, we will not go deep into the mathematics of each case since it’s extremely advanced, but the theoretical part is approached by some method called quasi-classical for the description of quantum electrodynamical processes in the Coulomb field at high energies.
+Delbruck scattering is considered one of the nonlinear processes in quantum electrodynamics, along with photon splitting (in which a highly energetic photon turns in the electric field of an atom into two photons sharing its energy, and the observation of photon splitting is extremely hard) such that the interaction between the electromagnetic fields is not linear due to the polarizability of the vacuum, but what makes it the most interesting nonlinear processes that it can be precisely tested by experiment, also that it is an extremely interesting and successful experimental tool in nuclear physics.
+
+It was first theoretically discovered in 1933 by Max Delbrück to account for the differences between what was observed in a Compton scattering experiment on heavier atoms (done by Meitner and Kösters) and what was predicted.
+Then it was confirmed 20 years later by Hans Bethe who named it “Delbrück scattering”.
+It was observed for the first time at DESY in 1973 for high-energy and small-angle photon scattering.
+
+The cross-section of this process will have $Z_\alpha$ as a parameters or $Z|e|$ after setting $\hbar=c=1$, and since this specific photon-photon scattering contain external fields mediating the interaction between photons, the cross-section will be larger than for pure photon-photon scattering.
+
+\section{Breit–Wheeler process}
+
+\begin{figure}[!th]
+	\centering
+	\includegraphics[width=0.5\linewidth]{figures/Breit-Wheeler.png}
+	\caption{Feynman diagram of Breit–Wheeler pair process and the photon-photon scattering.}
+	\label{fig:Feynman-diagram-of-Breit–Wheeler}
+\end{figure}
+
+\begin{figure}[!th]
+	\centering
+	\includegraphics[width=0.5\linewidth]{figures/Feynman-Breit.png}
+	\caption{Feynman diagram of Breit–Wheeler pair production.}
+	\label{fig:Breit–Wheeler-pair}
+\end{figure}
+
+This process is another photo-photon interaction, and it resembles photon-photon scattering in which they both are scattering of two photons in the initial state, Figure~\ref{fig:Feynman-diagram-of-Breit–Wheeler}.
+Breit–Wheeler process is the collision of two photons then producing electron-positron pair see Figure~\ref{fig:Breit–Wheeler-pair}, it defined as
+ $\gamma + \gamma' \rightarrow e^+ + e^-$, it was predicted first by Gregory Breit and John Wheeler in 1934 from the work of Dirac.
+The energies of the photons must reach the threshold of the rest mass of the electron-positron pair, since the threshold for the electron-positron is much smaller than muon-antimuon, tau-antitau, quark-antiquark, it’s more likely to happen, however, the possibility of such process is very small in general.
+The pair production is nearly impossible to occur just between two photons generated randomly, this triggers the idea of using two ionized gold atoms to create the highly energetic photons which will have larger cross-sections, using this idea the pair production was directly observed at “Relativistic Heavy Ion Collider” in 2021.
+As the type of photons pair differ the process also will be different and each can be identified very easily in a collider, for example, the virtual-virtual photon collision is different from the virtual-real photon collision.
+The Breit–Wheeler is from the type of real-real photon collision it occurs at large angles.
+Also, there is another similar process in which a photon interacts with a laser field known as the multiphoton Breit–Wheeler process or nonlinear Breit–Wheeler.
+It is strong evidence of Einstein’s equation $E=mc^2$ as it shows the conversion from energy to matter.
 
 \section{Conclusion}
-In conclusion, light-by-light scattering is a fascinating and complex phenomenon that is still not fully understood.
 With the help of advanced technologies such as the LHC and Feynman diagrams, we can continue to study and learn more about this elusive interaction.
+Light by light scattering is a phenomenon in which two photons interact and produce a new photon.
+This process is extremely rare and difficult to observe, but it has important implications for our understanding of the universe.
+
+One of the most significant applications of light by light scattering is in the study of the behavior of particles at very high energies.
+By studying the scattering of photons, scientists can learn more about the properties of fundamental particles such as electrons and quarks.
+This knowledge can help us better understand the fundamental forces that govern the universe, such as electromagnetism and the strong and weak nuclear forces.
+
+Additionally, light by light scattering can be used to study the properties of materials.
+By shining a laser at a material and analyzing the scattered light, scientists can determine the composition, structure, and other properties of the material.
+This can be useful for a variety of applications, including materials science and medical imaging.
+
+In conclusion, light by light scattering is a fascinating phenomenon that has important applications in a variety of fields.
+By studying the scattering of photons, we can learn more about the fundamental forces that govern the universe and the properties of materials.
+This knowledge can help us advance our understanding of the world around us and develop new technologies.
 
 \newpage
+\nocite{*}
 \printbibliography

### 2022-12-15 latex/PHYS441/paper.tex
Al done

@@
-\documentclass{article}
+\documentclass{paper}
 \usepackage{amsmath}
@@
 
-\title{Light-Light Scattering: Feynman Diagrams}
+\title{Light-by-Light Scattering: Description \& Applications}
 \author{Alfaifi, Ammar \and Al-Ali, Muhammad}
-\date{June 6, 2022}
+\date{\today}
 
@@
 
-Feynman diagrams show that light-by-light scattering is a very rare event because photons do not have any electric charge, and therefore, do not experience the electromagnetic force.
+Feynman diagrams show that light-by-light scattering is a very rare event because photons do not have any electric charge, and therefore, do not experience electromagnetic force.
 This means that photons do not interact with each other through the exchange of virtual particles such as gluons or W/Z bosons.
 
-However, photons can still interact with each other through the weak force, which is one of the four fundamental forces of nature.
+However, photons can still interact with each other through weak force, which is one of the four fundamental forces of nature.
 The weak force is responsible for the radioactive decay of atoms, but it is much weaker than the electromagnetic force.
@@
 The Feynman diagrams for light-light scattering are shown in Figure~\ref{fig:light-light-scattering-feynman}.
-The process can be thought of as the interaction of two photons, labeled as 1 and 2, which scatter off each other and produce two outgoing photons, labeled as 3 and 4.
+The process can be thought of as the interaction of two photons, labelled as 1 and 2, which scatter off each other and produce two outgoing photons, labelled as 3 and 4.
 The process is mediated by the exchange of a virtual electron-positron pair, which is created and annihilated in the process.
@@
 	\includegraphics[width=0.5\linewidth]{figures/EventDisplay_LbyL.png}
-	\caption{ATLAS event display showing the energy deposits of two photons in the electromagnetic
-		calorimeter (green) on opposite sides and no other activity in the detector, which is the clean signature
+	\caption{ATLAS event display shows the energy deposits of two photons in the electromagnetic
+		calorimeter (green) on opposite sides and no other activity in the detector, which is the clear signature
 		of light-by-light scattering. The Feynman diagram of this process is shown in the lower right corner.
@@
 \end{figure}
-Direct evidence for light-by-light scattering at high energy had proven elusive for decades, until the Large Hadron
+Direct evidence for light-by-light scattering at high energy had proven elusive for decades until the Large Hadron
 Collider (LHC) began its second data-taking period (Run 2). Collisions of lead ions in the LHC
@@
 	\centering
-	\includegraphics[width=0.6\linewidth]{figures/photons.jpg}
+	\includegraphics[width=\linewidth]{figures/photons.jpg}
 	\caption{Three different types of interactions between photons.
-In the left picture an incoming photon interacts with e.g.
+In the left picture, an incoming photon interacts with e.g.
 the strong external field from an atomic nucleus, resulting in an outgoing photon with new characteristics (Delbrück scattering); in the middle picture a single photon interacts with e.g.
-a strong external magnetic field as to produce two low-frequency photons (photon splitting); finally, in picture to the right two photons interacts directly via the quantum vacuum, producing two scattered photons.
+a strong external magnetic field as produce two low-frequency photons (photon splitting); finally, in the picture to the right two photons interacts directly via the quantum vacuum, producing two scattered photons.
 (Courtesy: Mattias Marklund)}
@@
 \end{figure}
-photon–photon scattering is a non-classical effect arising in quantum electrodynamics (QED) due to virtual electron–positron pairs in vacuum.
-Under every-day circumstances, the effect is very weak, see Figure~\ref{fig:corss-secction}.
-However, under the right conditions the interaction between photons and these virtual pairs will result in what is known as photon–photon collisions.
-Close relatives to this effect is Delbrück scattering and photon splitting, see Figure~\ref{fig:photon-types}.
+photon–photon scattering is a non-classical effect arising in quantum electrodynamics (QED) due to virtual electron–positron pairs in a vacuum.
+Under everyday circumstances, the effect is very weak, see Figure~\ref{fig:corss-secction}.
+However, under the right conditions, the interaction between photons and these virtual pairs will result in what is known as photon–photon collisions.
+Close relatives to this effect are Delbrück scattering and photon splitting, see Figure~\ref{fig:photon-types}.
 Formulated as an effective field theory, using the Heisenberg–Euler Lagrangian (Heisenberg \& Euler (1936); Schwinger (1951)), such scattering results in nonlinear corrections to Maxwell’s vacuum equations, similar in form to what is known as Kerr nonlinearities in nonlinear optics.
-The effective self-interaction term in Maxwell’s equations is small (proportional to the fine structure constant squared), which means that the field strengths need to reach appreciable values until such effects become pronounced (Marklund \& Shukla (2006); Mourou et al.
+The effective self-interaction term in Maxwell’s equations is small (proportional to the fine structure constant squared), which means that the field strengths need to reach appreciable values until such e/Ndffects become pronounced (Marklund \& Shukla (2006); Mourou et al.
 (2006)).
-However, the smallness of the photon-photon scattering cross section has been argued to be a possible window to new physics (Anoniadis (1998); Arkani-hamed et al.
+However, the smallness of the photon-photon scattering cross section has been argued to be a possible window to new physics (Anoniadis (1998); Arkani-Hamed et al.
 (1998); Cheung (1999); Davoudiasl (1999)), such as weak scale quantum gravity.
-Physical implications, as well as possible detection techniques, of the effects of photon–photon scattering have attracted interest since the 1930’s (for surveys, see Refs.
+Physical implications, as well as possible detection techniques, of the effects of photon–photon scattering have attracted interest since the 1930s (for surveys, see Refs.
 Marklund \& Shukla (2006) and Mourou et al.
@@
 \label{subsec:light-light-low-energy}
-A signal of 106±14 positrons above background has been observed in collisions of a low-emittance 46.6 GeV electron beam with terawatt pulses from a Nd:glass laser at 527 nm wavelength in an experiment at the Final Focus Test Beam at SLAC.
+A signal of 106±14 positrons above the background has been observed in collisions of a low-emittance 46.6 GeV electron beam with terawatt pulses from a Nd:glass laser at 527 nm wavelength in an experiment at the Final Focus Test Beam at SLAC.
 The positrons are interpreted as arising from a two-step process in which laser photons are backscattered to GeV energies by the electron beam followed by a collision between the high-energy photon and several laser photons to produce an electron-positron pair.
@@
 
-\begin{figure}[!th]
-	\centering
-	\includegraphics[width=0.5\linewidth]{figures/photon-photon-scattering.jpg}
-	\caption{vacuum polarization feynman diagram.}
-	\label{fig:vacuum-polarization}
-\end{figure}
-
 Delbruck scattering to understand it first we need to understand a process called vacuum polarization.
-Vacuum polarization is a random process by which a photon produces a pair of virtual fermion and anti-fermion and then annihilates repeatedly, the Feynman diagram of this process can be seen from the graph Figure~\ref{fig:vacuum-polarization}.
+Vacuum polarization is a random process by which a photon produces a pair of virtual fermion and anti-fermion and then annihilates repeatedly, the Feynman diagram of this process can be seen in the graph Figure~\ref{fig:corss-secction}.
 Since this process involves charged particles (electron and anti-electron) this pair will create a virtual dipole and then would interact with other charged particles especially if the process occurred in nuclei.
@@
 \end{figure}
-This shielding due to the vacuum polarization results in a deviation of the experimental Landé g-factor of the electron from the theoretical value and a difference in energy between two energy levels that was later called “Lamb shift” Figure~\ref{fig:lamb-shift}, which contradicts the prediction of the Dirac equation.
+This shielding due to the vacuum polarization results in a deviation of the experimental Landé g-factor of the electron from the theoretical value and a difference in energy between two energy levels that was later called the “Lamb shift” Figure~\ref{fig:lamb-shift}, which contradicts the prediction of the Dirac equation.
 These theoretical deviations were calculated by the genius Hans Bethe, and what is funny is that he did the calculation on his return train ride when he was returning to Cornell after attending a select workshop in the summer of 1947 in Shelter Island.
@@
 Due to this phenomenon, the notion of vacuum changed, this is because the vacuum is not empty anymore but will have some fluctuation and the presence of this virtual particle will disturb the electromagnetic field, however, the fluctuation and the presence of these charged particles will follow the energy-time Heisenberg uncertainty principle so that they will annihilate quickly.
-To sum up, vacuum polarization is the virtual fermion-antifermion produced from vacuum-polarized disturbed electromagnetic field.
+To sum up, vacuum polarization is the virtual fermion-antifermion produced from a vacuum-polarized disturbing electromagnetic field.
 
@@
 
-The cross-section of this process will have $Z_\alpha$ as a parameters or $Z|e|$ after setting $\hbar=c=1$, and since this specific photon-photon scattering contain external fields mediating the interaction between photons, the cross-section will be larger than for pure photon-photon scattering.
+The cross-section of this process will have $Z_\alpha$ as parameters or $Z|e|$ after setting $\hbar=c=1$, and since this specific photon-photon scattering contain external fields mediating the interaction between photons, the cross-section will be larger than for pure photon-photon scattering.
 
@@
 
-This process is another photo-photon interaction, and it resembles photon-photon scattering in which they both are scattering of two photons in the initial state, Figure~\ref{fig:Feynman-diagram-of-Breit–Wheeler}.
+This process is another photo-photon interaction, and it resembles photon-photon scattering in which they both are scattering of two photons in the initial state, see Figure~\ref{fig:Feynman-diagram-of-Breit–Wheeler}.
 Breit–Wheeler process is the collision of two photons then producing electron-positron pair see Figure~\ref{fig:Breit–Wheeler-pair}, it defined as
@@
 
-One of the most significant applications of light by light scattering is in the study of the behavior of particles at very high energies.
+One of the most significant applications of light-by-light scattering is in the study of the behaviour of particles at very high energies.
 By studying the scattering of photons, scientists can learn more about the properties of fundamental particles such as electrons and quarks.
@@
 
-Additionally, light by light scattering can be used to study the properties of materials.
+Additionally, light-by-light scattering can be used to study the properties of materials.
 By shining a laser at a material and analyzing the scattered light, scientists can determine the composition, structure, and other properties of the material.
@@
 
-In conclusion, light by light scattering is a fascinating phenomenon that has important applications in a variety of fields.
+In conclusion, light-by-light scattering is a fascinating phenomenon that has important applications in a variety of fields.
 By studying the scattering of photons, we can learn more about the fundamental forces that govern the universe and the properties of materials.

### 2022-12-20 latex/ICS343/ICS343.tex
Add ICS343 summary

\documentclass{article}

\usepackage[a4paper, lmargin=10mm, rmargin=70mm]{geometry}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{graphicx}

\title{ICS343 Subjective Part Summary}
\author{Alfaifi, Ammar}
\date{}

\begin{document}

\maketitle

\section{TCP/IP Protocol Suit}%
\label{sec:TCP}
\begin{itemize}

  \item Transmission Control Protocol/Internet Protocol (TCP/IP) is a protocol suit, meaning a set of functionalities
        provided between each layer.
  \item It is built in layer structure where the each layer takes services from the one bellow it,
        and provides services to the above it.
  \item Application, Transport, Network, Data link, Physical
        \begin{itemize}
          \item \textbf{Physical}: carries individual bits in a frame. communication between two devices
                in the this layer is still logical. under this layer is the transmission medium; carring
                electrical or optical signals. Logical unit is \textit{bit}
          \item \textbf{Data link}: Any protocol that takes datagram and carries it in the link is suffice
                for this layer. This layer takes datagram encapsulate it into a \textit{frame}. hop-to-hop communication.
          \item \textbf{Network}: is responsible for creating a connection between the source and destination computers. It's
                host-to-host communication and routing the packet to best path. It includes the main protocol, Internet Protocol (IP),
                which defines a packet called \textit{datagram}.
          \item IP is connectionless, that provides no flow control, no error control, and no congestion control services.
                It provides unicast  (one-to-one), and multicast (one-to-many) routing protocols.
                This layer has auxiliary protocols: ICMP helps IP to report some problems when routing a packet,
                DHCP helps IP to get the network-layer address's for a host, ARP help IP to find the data-link layer address
                of a host or a router given the network-layer address.
          \item \textbf{Transport}: the lgical connection is end-to-end. The packet is called a segment or user datagram.
                It has Transmission Control Protocol (TCP), which is a connection-oriented that established a logical connection
                before sending data. User Datagram Protocol (UDP), is a connectionless, which is simple and fast.
          \item \textbf{Application}: Logical connection is end-to-end. They exchange \textit{messages}.
                The communications is between two \textit{processes}. Examples: HTTP, SMTP, FTP, SSH, DNS is used by other
                protocols to find the network-layer address of a host.

          \item well-known ports : 0 - 1023, controlled by ICANN.
          \item registered ports: 1024 - 49151, not controlled by ICANN, but can be registered.
          \item dynamic ports: 49152 - 65535, as temporary or private ports.
        \end{itemize}

\end{itemize}

\section{Transport-Layer Protocol}%
\label{sec:TLP}

\subsection{Stop-and-Wait Protocol}%
\label{subsec:stop-and-wait protocols}

\begin{itemize}
  \item Both sender and receiver use sliding window of size 1.
  \item sends one packet at a time, and waits for an ACK.
  \item every sent packet, a timer starts.
  \item ACK value is the sequence \# of next packet.
  \item the sequence numbers are: 0, 1, 0, 1, $\dots$
\end{itemize}

\subsection{Go-Back-N Protocol (GBN)}%
\label{subsec:GBN}

\begin{itemize}
  \item it tries to send many packets while waiting for an acknowledgment.
  \item ACK \# is \textit{cumulative}; the sequence \# of next packet.
  \item window size up to $2^m -1$.
  \item $S_f$ first outstanding, $S_{n}$, $S_{size}$
  \item from $S_{f}$ to $S_{n}-1$: packets sent but not acknowledged.
  \item from $S_{n}$: packets to be sent
  \item window slides when an error-free ACK, $S_{f} \leq \text{ackNo} \leq S_{n}$, received.
  \item receive window is always of size 1.
  \item receive window, $R_{n}$, next packet expected.
  \item there is only one timer, for the first outstanding packet.
  \item at the timer expiration, \textit{all} outstanding packets are resent.
  \item window is full if $S_{n} = (S_{f} + S_{size}) \mod 2^{m}$
  \marginpar{
    \includegraphics[width=0.5\textwidth]{figures/wind.jpeg}
        }
  \item in receiver side, with error-free ACK, $S_{f}=\text{ackNo}$. if $\text{ackNo}=S_{n}$ stop timer.
\end{itemize}

\subsection{Selective-Repeat Protocol}%
\label{subsec:SRP}
GBN is inefficient if network loses a lot of packets.
\begin{itemize}
  \item sender window size is $2^{m-1}$, the receive window should equal the send window
  \item In reliable protocols, receiver never deliver packets out of order to the application layer.
  \item ackNo is the sequence \# of the received packet; ackNo isn't cumulative.
        It does \textbf{not} mean all packets with sequence \# less than ackNo are received.
\end{itemize}

\subsection{Bidirectional Protocols: Piggybacking}%
\label{subsec:bidre}

The idea is that, we make the packets that carries data, to carry also the ackNo at
the same time. So that we don't overwhelm the network with packets carring just ackNo.

\section{HyperText Transfer Protocol}%
\label{sec:HTTP}

\begin{itemize}
  \item The server uses the port 80, and 443 for secured one (HTTPS).
  \item form v1.1 persistent mode is the default.
  \item In the nonpersistent connection: the client opens a TCP connection and sends a request.
        The server responds and closes the connection. The client reads and colses the connection.
  \item Methods in the header message:
        \begin{itemize}
          \item GET: requests a doc.
          \item HEAD: requests info about a doc.
          \item POST: sends info to the server.
          \item PUT: sends a doc to the server.
        \end{itemize}
  \item Some request headers:
        \begin{itemize}
          \item user-agent: the client program.
          \item accept: media formats client understands.
          \item host: host and port number of client.
          \item if-modified-since: last update of file; this tell the server
                to send the data only if it was modified since this data.
        \end{itemize}
  \item some response headers:
        \begin{itemize}
          \item server: info about the server.
          \item last-modified: the date of last change.
          \item content-type: media type.
        \end{itemize}
\end{itemize}

\section{Electronic Mail (Email)}%
\label{sec:Email}

\begin{itemize}
  \item US: user agent, MTA: message transfer agent, MAA: message access agent.
  \item US, using MTA client, sends a msg to the MTA server then to a spool (a queue)
        then to MTA client in the server then to the MTA server of reciever then to
        the receiver UA mail box them to MAA server then to MAA client.
  \item it needs 2 UAs, 2 pairs of MTAs (slient \& server), and a pair of MAAs.

\subsection{Message Transfer Agent: SMTP}%
        \label{subsec:SMTP}

        Some of Simple Mail Transfer Protocol (SMTP) commands:
        \begin{itemize}
          \item HELO: sender's host name
          \item MAIL FROM: sender of the msg
          \item RCPT TO: the recipient.
          \item DATA: message body.
          \item QUIT: ends the message
        \end{itemize}

\subsection{Message Access Agent}%
\label{subsec:MAA}
\begin{itemize}
  \item Post Office Protocol (POP) is very simple to get the mails form the server mail boxes.
  \item Internet Mail Access Protocol (IMAP) it can check the header of a msg, search, or partially
        download a msg.
\end{itemize}
  \item Multipurpose Internet Mail Extension (MIME) is to help sending non-textual data.
  \item Base64 is one of the methods to encode data into ASCII characters.
        Base64 means $\log_{2}64=6$ bits, converts each 6 bits into their ASCII chars.
  \item Note that 0 corresponds to \textbf{A}, so \textbf{a} is 26 and so on.
  \item the sequence: \textbf{A-Z}, \textbf{a-z}, \textbf{0-9}
\end{itemize}

\end{document}
